### 2020-02-19 conditional_independence_notes/paper.tex
Entailment for conditional independence: version 1

\documentclass{article}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\usepackage[utf8]{inputenc}
\usepackage{fullpage}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{xfrac}

\theoremstyle{definition}
\newtheorem{definition}{Definition}

\newcommand{\maxNumNodes}{\mathcal{M}_{\mathcal{N}}}
\newcommand{\maxNumClauses}{\mathcal{M}_{\mathcal{C}}}
\newcommand{\maybe}{\sfrac{1}{2}}

\author{Paulius Dilkas}
\title{A Constraint for Conditional Independence}

\begin{document}
\maketitle
% TODO: predicates should always be the first dimension, indices the second

Goal: conditional independence of two predicates conditioned on a
conjunction/disjunction of any number of predicates or a negation of a
predicate.

Central question: does this clause contain this predicate (filtering out
the possibility that the predicate is mentioned in the conditioning part)? We
can answer this question by iterating over each position and answering two other
questions:
\begin{itemize}
\item Does this position contain this predicate? Yes/No/Maybe.
\item Is this position part of the structure we are conditioning on? Yes/No/Maybe.
\end{itemize}

For each clause, we have two position x predicate matrices. One of them says (in
a yes/no/maybe fashion) whether this predicate can be in this position (that's
A). The other says whether this predicate is masked in this position
(yes/no/maybe).

The propagation algorithm for independence is extended with masks that are
either potential or definite. Masking happens in two stages: first, we mask
expressions within formulas, and then predicates. Masking algorithm uses an
algorithm for perfect bipartite matching.

For any integer $n$, let $[n] = \{ 0, 1, \dots, n-1 \}$.

\begin{definition}[Conditional independence]
  When assigning edges to the graph, mask off all possible instances of the
  expression we're conditioning on. For example, if we're conditioning on $a
  \land b$, then $c \gets a \land a \land b$ should produce no edges.
\end{definition}

\begin{definition}
  A predicate is \emph{definitely masked} if all instances are definitely
  masked. A predicate is \emph{potentially masked} if all instances are
  definitely/potentially masked.
\end{definition}

\begin{definition}[Primitives]
  For each clause, let $\mathbf{P}$ be a $\maxNumNodes{} \times |\mathcal{P}
  \cup \{ \land, \lor, \neg \}|$ matrix defined as
  \[
    P_{p,i} = \begin{cases}
      1 & \text{if } \mathtt{values}[i] = \{ p \} \\
      0 & \text{otherwise.}
    \end{cases}
  \]
  This matrix shows what symbol can (or must) be in what position.

  Likewise, let $\mathbf{C}$ be a $\maxNumNodes{} \times \maxNumNodes{}$ matrix
  defined as
  \[
    C_{i,j} = \begin{cases}
      1 & \text{if } \mathtt{structure}[i] = \{ j \} \\
      0 & \text{otherwise.}
    \end{cases}
  \]
  The $(i,j)$-th element shows whether node $i$ can be a child of node $j$.
\end{definition}

\begin{table}
  \centering
  \caption{Combining $\mathbf{M}$ and $\mathbf{P}$. The $\#$ stands for `impossible'.}
  \begin{tabular}{lll}
    \toprule
    & $P = 1$ & $P = 0$ \\
    \midrule
    $M = 1$ & $0$ & $\#$ \\
    $M = 0$ & $1$ & $0$ \\
    \bottomrule
  \end{tabular}
\end{table}
% P and C could be functions instead

\begin{algorithm}
  \caption{Constructing the adjacency matrix $\mathbf{A}$}
  \SetKwFunction{constructMask}{constructMask}
  \SetKwFunction{upgradeTo}{upgradeTo}
  $\mathbf{A} \gets \mathbf{0}_{|\mathcal{P}|, |\mathcal{P}|}$\;
  \ForEach{clause $c$}{
    Construct $\mathbf{P}$ and $\mathbf{C}$\;
    $\mathbf{M} \gets \constructMask{$\odot$, $\mathcal{C}$} $\;
    Let $\mathbf{A'} \in \{ 0, \maybe{}, 1 \}^{|\mathcal{P}| \times
      \maxNumNodes{}}$ be such that $A'_{i,j} = \begin{cases}
      0 & \text{if } P_{i,j} = M_{i,j} \ne \maybe{} \\
      1 & \text{if } P_{i,j} = 1 \text{ and } M_{i,j} = 0 \\
      \maybe{} & \text{otherwise}
    \end{cases}$\;
    \For{$h \in \mathtt{head}[c]$}{
      \For{$p \in \mathcal{P}$}{
        \For{$i \in [\maxNumNodes{}]$}{
          \uIf{$|\mathtt{head}[c]| = 1$ {\bf and } $A'_{p,i} = 1$}{
            $A_{h \gets p}.\upgradeTo{$1$}$\;
          }
          \ElseIf{$|\mathtt{head}[c]| = 1$ {\bf and} $A'_{p,i} = \maybe{}$}{
            $A_{h \gets p}.\upgradeTo{$\maybe{}$}$\;
          }
        }
      }
    }
  }
\end{algorithm}

The adjacency matrix $\mathbf{A} \in \{ 0, \maybe{}, 1 \}^{|\mathcal{P}| \times
  |\mathcal{P}|}$ is defined as:
\[
  A[h \gets p] = \begin{cases}
    1 & \text{if } \exists c \in [\maxNumClauses{}]: \mathtt{head}[c] = \{ h
    \} \text{ and } \exists i \in [\maxNumNodes{}]: A_{c;p,i} = 1 \\
    0 & \text{otherwise}
  \end{cases}\;
\]

How does this propagate back to the variables? Each `maybe' edge should hold a
set of (variable, value) pairs that can be eliminated if the propagation
algorithm decides that it cannot exist. Problem: propagating through
$\mathbf{A'}$ would mean propagating through $\mathbf{M}$ and $\mathbf{P}$. If
they're both uncertain, then we can't do anything. If $\mathbf{P}$ is uncertain,
we can remove the predicate, but if $\mathbf{M}$ is uncertain, there might be
multiple ways to complete the mask.

Essentially, we have three layers, i.e., we build two layers of abstraction on
top of the CP model. We make a decision on the top layer, and propagate the
changes to the bottom layers. A weakness is that some high-level problems can
have multiple solutions in the bottom layers, in which case we can't commit to
any particular solution and so can't do anything.

\begin{algorithm}
  \caption{Construct the mask matrix $\mathbf{M}$}
  \SetKwProg{Fn}{Function}{:}{}
  \SetKwFunction{markIndices}{markIndices}
  \SetKwFunction{constructMask}{constructMask}
  \KwData{the expression we are conditioning on: $\bigodot_{\mathsf{p} \in
      \mathcal{C}} \mathsf{p}$}
  \Fn{\constructMask{$\odot$, $\mathcal{C}$}}{
    $\mathbf{M} = \mathbf{0}_{|\mathcal{P}|, \maxNumNodes{}}$\;
    \For(\tcp*[h]{for each possible root}){$r \in [\maxNumNodes{}]$ such that
      $P_{r,\odot} \ne 0$}{
      $E \gets \{ \{p, i\} \mid p \in \mathcal{P}, i \in [\maxNumNodes{}],
      C_{i,r} \ne 0, P_{i,p} \ne 0 \}$\;
      $\mathbf{M'} \gets \markIndices{$(\mathcal{P}, [\maxNumNodes{}], E)$,
        $\maybe{}$}$\tcp*[l]{potential positions}
      $\mathbf{M} \gets \max\{ \mathbf{M}, \mathbf{M'}\}$\;
      \If{$P_{r,\odot} = 1$ {\bf and} $\mathbf{M'} \ne \mathbf{0}$}{
        $E \gets \{ \{p, i\} \mid p \in \mathcal{P}, i \in [\maxNumNodes{}],
        C_{i,r} = 1, P_{i,p} = 1, M'_{p,i} \ne 0 \}$\;
        % Because this M is a subset of the previous M
        $\mathbf{M} \gets \max \{ \mathbf{M}, \markIndices{$(\mathcal{P},
          [\maxNumNodes{}], E)$, $1$} \}$\tcp*[l]{determined positions}
      }
    }
    \Return{$\mathbf{M}$}\;
  }
\end{algorithm}

\begin{algorithm}
  \caption{Predicate-perfect bipartite matching}
  \SetKwProg{Fn}{Function}{:}{}
  \SetKwFunction{markIndices}{markIndices}
  \KwData{a bipartite graph $G = (\mathcal{P}, \mathcal{I}, E)$
    (predicates and indices) and the value $v$ to use for the matrix ($1$ or
    $\maybe{}$)}
  \KwResult{a $|\mathcal{P}| \times |\mathcal{I}|$ mask matrix $\mathbf{M}$}
  \Fn{\markIndices{$G$, $v$}}{
    $\mathbf{M} \gets \mathbf{0}_{|\mathcal{P}|,\maxNumNodes{}}$\;
    \lIf{$|\mathcal{I}| < |\mathcal{P}|$}{\Return{$\mathbf{M}$}}
    \While{true}{
      Let $(p, i) \in E$ be such that $M_{p,i} = 0$\;
      \lIf{there is none}{break}
      Let $G'$ be $G$ without vertices $p$, $i$ and all the incident edges\;
      Use alternating paths to find a $\mathcal{P}$-perfect matching $P$ in $G'$\;
      \uIf{found}{
        \lForEach{edge $(p', i') \in P$}{$M_{p',i'} = v$}
      }
      \Else{
        remove $e$ from $G$\;
        \lIf{this leaves an isolated vertex}{break}
      }
    }
    \Return{$\mathbf{M}$}\;
  }
\end{algorithm}

% TODO: need a way to prove the correctness of the algorithm. I'm not sure how I
% would approach empirical testing, so maybe do a proof instead.

% \begin{algorithm}
%   \SetKwProg{Fn}{Function}{:}{}
%   \SetKwFunction{potentialRoots}{potentialRoots}
%   \SetKwFunction{getTreeValues}{getTreeValues}
%   \SetKwFunction{getTreeStructureDomainValues}{getTreeStructureDomainValues}
%   \SetKwData{clause}{clause}
%   \KwData{connective $c$, a set of predicates $P$}
%   \Fn{\potentialRoots{\clause, $i$}} {
%     $R \gets \emptyset$\;
%     $V \gets \clause.\getTreeValues{$i$}$\;
%     \For{$v \in V$}{
%       \uIf{$v = c$}{
%         $R \gets R \cup \{ (i, |V| = 1) \}$\;
%       }
%       \ElseIf{$v \in P$}{
%         $R' \gets \clause.\getTreeStructureDomainValues{$i$}$\;
%         $R \gets R \cup \{ (r, |R'| = 1) \mid r \in R' \}$\;
%       }
%     }
%     \Return{$R$}\;
%   }
%   \caption{Potential root nodes of the required expression, assuming that the
%     node at index $i$ is part of the expression}
% \end{algorithm}

\end{document}