### 2020-02-23 paper_extension/paper.tex
A description of the 'required formula' constraint

\documentclass{article}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\usepackage[utf8]{inputenc}
\usepackage{fullpage}
\usepackage{mathtools}
\usepackage[capitalise]{cleveref}

\newcommand{\maxNumNodes}{\mathcal{M}_{\mathcal{N}}}
\newcommand{\maxNumClauses}{\mathcal{M}_{\mathcal{C}}}

\Crefname{subformula}{Subformula}{Subformulae}
\creflabelformat{subformula}{#2(#1)#3}

\begin{document}

\section{A Required Subformula}
% also note that /\ = max, and \/ = min

We describe how to add constraints to the model in order to guarantee that a
particular subformula appears somewhere in the program. In this section, a
\emph{subformula} is a formula of the form
\begin{equation} \label[subformula]{eq:subformula}
  \bigodot_{i=1}^n \mathsf{p}_i,
\end{equation}
where $\odot \in \{ \land, \lor \}$, and $(\mathsf{p}_i)_{i=1}^n$ is a finite
sequence of predicates, arguments of which are immaterial. Let
\cref{eq:subformula} be precisely the required subformula.

We say that a clause has \cref{eq:subformula} if we can find indices
$i_0, i_1, \dots, i_n \in \{0, \dots, \maxNumNodes{} - 1 \}$ such that
\[
  \mathtt{values}[i_0] = \odot,
\]
and, for each $j = 1, \dots, n$,
\[
  \mathtt{structure}[i_j] = i_0,
\]
and
\[
  \mathtt{values}[i_j] = \mathsf{p}_j.
\]
We use this idea to define several data structures for each clause, gradually
building up to a Boolean indicator for whether the clause has
\cref{eq:subformula}. First, let $\mathbf{A}$ be an $\maxNumNodes{} \times
\maxNumNodes{}$ Boolean matrix defined as
\[
  A_{i,j} \coloneqq (\mathtt{values}[i] = \odot) \land (\mathtt{structure}[j] = i),
\]
i.e., the $(i,j)$-th element is true if node $i$ is the root node of the
subformula, and node $j$ is its child. Next, let $\mathbf{B}$ be an
$\maxNumNodes{} \times n$ Boolean matrix defined as
\[
  B_{i,j} \coloneqq (\mathtt{values}[i] = \mathsf{p}_j),
\]
where the $(i,j)$-th element indicates whether node $i$ contains predicate
$\mathsf{p}_j$. We can then use $\mathbf{A}$ and $\mathbf{B}$ to define an
$\maxNumNodes{} \times n$ Boolean matrix $\mathbf{C}$ where the $(i,j)$-th
element shows whether node $i$ is a parent of a node that contains predicate
$\mathsf{p}_j$. We then have that $\mathbf{C} \coloneqq \mathbf{A} \mathbf{B}$,
i.e.,
\[
  C_{i,j} \coloneqq \bigvee_k A_{i,k} \land B_{k,j}.
\]
The Boolean variable $f_i$ that indicates whether the $i$-th clause has the
required subformula (for $i = 0, \dots, \maxNumClauses{} - 1$) can then be
expressed as
\[
  f_i \coloneqq \bigvee_j \bigwedge_k C_{j,k},
\]
i.e., the clause has \cref{eq:subformula} if there is a suitable parent node
that has all $n$ predicates as its children. Note that this allows for the
parent node to have more than $n$ children, and the order of children is
inconsequential. Finally, we can require the program to have the subformula
simply by asking for
\[
  \bigvee_i f_i
\]
to be satisfied.

\end{document}