### 2020-05-11 text/paper2/bars.tex
Some new plots

% Created by tikzDevice version 0.12.3 on 2020-05-11 12:49:46
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (173.45,130.09);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (173.45,130.09);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( -0.00,  0.00) rectangle (173.45,130.09);
\end{scope}
\begin{scope}
\path[clip] ( 31.71, 30.69) rectangle ( 97.08,108.01);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.71, 30.69) rectangle ( 97.08,108.01);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.71, 36.29) --
	( 97.08, 36.29);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.71, 62.99) --
	( 97.08, 62.99);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.71, 89.69) --
	( 97.08, 89.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.71, 49.64) --
	( 97.08, 49.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.71, 76.34) --
	( 97.08, 76.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.71,103.05) --
	( 97.08,103.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 49.54, 30.69) --
	( 49.54,108.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 79.25, 30.69) --
	( 79.25,108.01);
\definecolor{drawColor}{RGB}{51,51,51}

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 71.97) -- ( 51.50, 75.90);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 75.90) -- ( 51.50, 71.97);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 71.07) -- ( 51.50, 74.99);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.99) -- ( 51.50, 71.07);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 80.47) -- ( 51.50, 84.39);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 84.39) -- ( 51.50, 80.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 75.71) -- ( 51.50, 79.64);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 79.64) -- ( 51.50, 75.71);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 73.51) -- ( 51.50, 77.44);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 77.44) -- ( 51.50, 73.51);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 71.96) -- ( 51.50, 75.88);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 75.88) -- ( 51.50, 71.96);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.26) -- ( 51.50, 78.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 78.19) -- ( 51.50, 74.26);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 70.50) -- ( 51.50, 74.42);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.42) -- ( 51.50, 70.50);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.45) -- ( 51.50, 78.37);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 78.37) -- ( 51.50, 74.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 71.60) -- ( 51.50, 75.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 75.53) -- ( 51.50, 71.60);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.07) -- ( 51.50, 77.99);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 77.99) -- ( 51.50, 74.07);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 72.86) -- ( 51.50, 76.78);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 76.78) -- ( 51.50, 72.86);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 71.26) -- ( 51.50, 75.18);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 75.18) -- ( 51.50, 71.26);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 71.60) -- ( 51.50, 75.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 75.53) -- ( 51.50, 71.60);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 71.53) -- ( 51.50, 75.46);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 75.46) -- ( 51.50, 71.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.17) -- ( 51.50, 78.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 78.09) -- ( 51.50, 74.17);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 72.05) -- ( 51.50, 75.98);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 75.98) -- ( 51.50, 72.05);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 70.59) -- ( 51.50, 74.51);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.51) -- ( 51.50, 70.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 71.60) -- ( 51.50, 75.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 75.53) -- ( 51.50, 71.60);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 71.73) -- ( 51.50, 75.65);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 75.65) -- ( 51.50, 71.73);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 71.27) -- ( 51.50, 75.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 75.19) -- ( 51.50, 71.27);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 70.86) -- ( 51.50, 74.79);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.79) -- ( 51.50, 70.86);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 72.78) -- ( 51.50, 76.70);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 76.70) -- ( 51.50, 72.78);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 73.08) -- ( 51.50, 77.01);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 77.01) -- ( 51.50, 73.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 70.27) -- ( 51.50, 74.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.19) -- ( 51.50, 70.27);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 75.09) -- ( 51.50, 79.01);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 79.01) -- ( 51.50, 75.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.26) -- ( 51.50, 78.18);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 78.18) -- ( 51.50, 74.26);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 71.92) -- ( 51.50, 75.84);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 75.84) -- ( 51.50, 71.92);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 70.89) -- ( 51.50, 74.82);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.82) -- ( 51.50, 70.89);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 70.81) -- ( 51.50, 74.73);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.73) -- ( 51.50, 70.81);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 71.32) -- ( 51.50, 75.24);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 75.24) -- ( 51.50, 71.32);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 73.51) -- ( 51.50, 77.43);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 77.43) -- ( 51.50, 73.51);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 72.68) -- ( 51.50, 76.60);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 76.60) -- ( 51.50, 72.68);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 75.68) -- ( 51.50, 79.61);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 79.61) -- ( 51.50, 75.68);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 70.08) -- ( 51.50, 74.01);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.01) -- ( 51.50, 70.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 72.32) -- ( 51.50, 76.24);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 76.24) -- ( 51.50, 72.32);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 70.44) -- ( 51.50, 74.37);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.37) -- ( 51.50, 70.44);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 70.61) -- ( 51.50, 74.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.54) -- ( 51.50, 70.61);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 69.97) -- ( 51.50, 73.89);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 73.89) -- ( 51.50, 69.97);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 78.26) -- ( 51.50, 82.18);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 82.18) -- ( 51.50, 78.26);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.73) -- ( 51.50, 78.66);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 78.66) -- ( 51.50, 74.73);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 71.06) -- ( 51.50, 74.99);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.99) -- ( 51.50, 71.06);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58,101.49) -- ( 51.50,105.41);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58,105.41) -- ( 51.50,101.49);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58,102.54) -- ( 51.50,106.46);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58,106.46) -- ( 51.50,102.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 77.11) -- ( 51.50, 81.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 81.03) -- ( 51.50, 77.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58,101.45) -- ( 51.50,105.38);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58,105.38) -- ( 51.50,101.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58,101.49) -- ( 51.50,105.41);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58,105.41) -- ( 51.50,101.49);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 98.71) -- ( 51.50,102.64);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58,102.64) -- ( 51.50, 98.71);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 85.70) -- ( 51.50, 89.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 89.62) -- ( 51.50, 85.70);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 70.59) -- ( 51.50, 74.52);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.52) -- ( 51.50, 70.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 69.91) -- ( 51.50, 73.84);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 73.84) -- ( 51.50, 69.91);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 70.65) -- ( 51.50, 74.58);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.58) -- ( 51.50, 70.65);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 70.71) -- ( 51.50, 74.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.63) -- ( 51.50, 70.71);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 70.41) -- ( 51.50, 74.33);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.33) -- ( 51.50, 70.41);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 77.48) -- ( 51.50, 81.41);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 81.41) -- ( 51.50, 77.48);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 80.61) -- ( 51.50, 84.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 84.53) -- ( 51.50, 80.61);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 87.23) -- ( 51.50, 91.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 91.16) -- ( 51.50, 87.23);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 73.91) -- ( 51.50, 77.83);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 77.83) -- ( 51.50, 73.91);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 49.54, 57.22) -- ( 49.54, 71.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 49.54, 47.48) -- ( 49.54, 34.20);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 38.40, 57.22) --
	( 38.40, 47.48) --
	( 60.68, 47.48) --
	( 60.68, 57.22) --
	( 38.40, 57.22) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 38.40, 52.13) -- ( 60.68, 52.13);
\definecolor{drawColor}{RGB}{51,51,51}

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.78) -- ( 81.21, 85.71);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.71) -- ( 81.21, 81.78);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.35) -- ( 81.21, 86.27);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.27) -- ( 81.21, 82.35);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.08) -- ( 81.21, 83.00);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.00) -- ( 81.21, 79.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.03) -- ( 81.21, 82.96);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.96) -- ( 81.21, 79.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.81) -- ( 81.21, 86.73);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.73) -- ( 81.21, 82.81);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.76) -- ( 81.21, 84.68);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.68) -- ( 81.21, 80.76);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.46) -- ( 81.21, 84.39);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.39) -- ( 81.21, 80.46);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.59) -- ( 81.21, 91.52);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 91.52) -- ( 81.21, 87.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.62) -- ( 81.21, 84.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.55) -- ( 81.21, 80.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.17) -- ( 81.21, 87.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.09) -- ( 81.21, 83.17);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.21) -- ( 81.21, 83.13);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.13) -- ( 81.21, 79.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.27) -- ( 81.21, 83.20);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.20) -- ( 81.21, 79.27);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.26) -- ( 81.21, 83.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.19) -- ( 81.21, 79.26);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.88) -- ( 81.21, 90.81);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.81) -- ( 81.21, 86.88);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 78.48) -- ( 81.21, 82.41);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.41) -- ( 81.21, 78.48);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.54) -- ( 81.21, 88.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.47) -- ( 81.21, 84.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.09) -- ( 81.21, 91.01);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 91.01) -- ( 81.21, 87.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.58) -- ( 81.21, 86.51);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.51) -- ( 81.21, 82.58);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.17) -- ( 81.21, 90.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.09) -- ( 81.21, 86.17);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.00) -- ( 81.21, 89.93);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.93) -- ( 81.21, 86.00);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.62) -- ( 81.21, 85.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.54) -- ( 81.21, 81.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.72) -- ( 81.21, 93.64);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 93.64) -- ( 81.21, 89.72);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.29) -- ( 81.21, 88.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.21) -- ( 81.21, 84.29);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.16) -- ( 81.21, 87.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.08) -- ( 81.21, 83.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.29) -- ( 81.21, 84.22);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.22) -- ( 81.21, 80.29);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 78.56) -- ( 81.21, 82.49);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.49) -- ( 81.21, 78.56);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.11) -- ( 81.21, 94.04);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 94.04) -- ( 81.21, 90.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.43) -- ( 81.21, 87.35);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.35) -- ( 81.21, 83.43);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.66) -- ( 81.21, 86.58);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.58) -- ( 81.21, 82.66);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.00) -- ( 81.21, 86.93);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.93) -- ( 81.21, 83.00);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.39) -- ( 81.21, 83.31);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.31) -- ( 81.21, 79.39);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.28) -- ( 81.21, 87.20);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.20) -- ( 81.21, 83.28);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.34) -- ( 81.21, 83.26);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.26) -- ( 81.21, 79.34);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.21) -- ( 81.21, 91.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 91.14) -- ( 81.21, 87.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 78.88) -- ( 81.21, 82.80);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.80) -- ( 81.21, 78.88);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.19) -- ( 81.21, 86.12);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.12) -- ( 81.21, 82.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 78.96) -- ( 81.21, 82.89);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.89) -- ( 81.21, 78.96);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.84) -- ( 81.21, 85.77);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.77) -- ( 81.21, 81.84);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.45) -- ( 81.21, 84.37);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.37) -- ( 81.21, 80.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.23) -- ( 81.21, 83.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.16) -- ( 81.21, 79.23);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.89) -- ( 81.21, 85.82);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.82) -- ( 81.21, 81.89);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 78.83) -- ( 81.21, 82.75);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.75) -- ( 81.21, 78.83);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.26) -- ( 81.21, 86.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.19) -- ( 81.21, 82.26);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.31) -- ( 81.21, 89.24);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.24) -- ( 81.21, 85.31);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 78.51) -- ( 81.21, 82.43);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.43) -- ( 81.21, 78.51);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.59) -- ( 81.21, 85.52);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.52) -- ( 81.21, 81.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.56) -- ( 81.21, 85.49);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.49) -- ( 81.21, 81.56);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.08) -- ( 81.21, 84.01);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.01) -- ( 81.21, 80.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.36) -- ( 81.21, 85.29);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.29) -- ( 81.21, 81.36);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.54) -- ( 81.21, 85.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.47) -- ( 81.21, 81.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.19) -- ( 81.21, 83.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.11) -- ( 81.21, 79.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.87) -- ( 81.21, 89.79);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.79) -- ( 81.21, 85.87);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 78.50) -- ( 81.21, 82.42);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.42) -- ( 81.21, 78.50);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.66) -- ( 81.21, 83.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.59) -- ( 81.21, 79.66);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.79) -- ( 81.21, 84.71);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.71) -- ( 81.21, 80.79);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.63) -- ( 81.21, 87.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.55) -- ( 81.21, 83.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.11) -- ( 81.21, 85.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.03) -- ( 81.21, 81.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.05) -- ( 81.21, 88.97);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.97) -- ( 81.21, 85.05);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.98) -- ( 81.21, 89.90);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.90) -- ( 81.21, 85.98);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.03) -- ( 81.21, 83.96);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.96) -- ( 81.21, 80.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 78.65) -- ( 81.21, 82.58);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.58) -- ( 81.21, 78.65);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.50) -- ( 81.21, 87.43);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.43) -- ( 81.21, 83.50);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 78.41) -- ( 81.21, 82.33);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.33) -- ( 81.21, 78.41);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.91) -- ( 81.21, 87.84);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.84) -- ( 81.21, 83.91);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.23) -- ( 81.21, 87.15);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.15) -- ( 81.21, 83.23);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.91) -- ( 81.21, 89.83);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.83) -- ( 81.21, 85.91);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.55) -- ( 81.21, 85.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.47) -- ( 81.21, 81.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.78) -- ( 81.21, 83.71);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.71) -- ( 81.21, 79.78);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.03) -- ( 81.21, 82.95);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.95) -- ( 81.21, 79.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 78.87) -- ( 81.21, 82.80);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.80) -- ( 81.21, 78.87);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.19) -- ( 81.21, 84.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.11) -- ( 81.21, 80.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.94) -- ( 81.21, 83.86);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.86) -- ( 81.21, 79.94);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.19) -- ( 81.21, 85.12);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.12) -- ( 81.21, 81.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.21) -- ( 81.21, 85.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.14) -- ( 81.21, 81.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.32) -- ( 81.21, 85.24);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.24) -- ( 81.21, 81.32);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.16) -- ( 81.21, 84.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.08) -- ( 81.21, 80.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.04) -- ( 81.21, 86.96);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.96) -- ( 81.21, 83.04);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.20) -- ( 81.21, 85.13);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.13) -- ( 81.21, 81.20);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.88) -- ( 81.21, 87.80);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.80) -- ( 81.21, 83.88);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.26) -- ( 81.21, 85.18);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.18) -- ( 81.21, 81.26);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 78.69) -- ( 81.21, 82.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.62) -- ( 81.21, 78.69);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.78) -- ( 81.21, 85.71);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.71) -- ( 81.21, 81.78);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.05) -- ( 81.21, 86.97);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.97) -- ( 81.21, 83.05);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 78.98) -- ( 81.21, 82.91);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.91) -- ( 81.21, 78.98);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.90) -- ( 81.21, 85.83);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.83) -- ( 81.21, 81.90);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.97) -- ( 81.21, 83.90);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.90) -- ( 81.21, 79.97);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.54) -- ( 81.21, 85.46);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.46) -- ( 81.21, 81.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.33) -- ( 81.21, 90.25);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.25) -- ( 81.21, 86.33);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.99) -- ( 81.21, 86.91);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.91) -- ( 81.21, 82.99);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.08) -- ( 81.21, 83.00);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.00) -- ( 81.21, 79.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.06) -- ( 81.21, 91.99);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 91.99) -- ( 81.21, 88.06);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.88) -- ( 81.21, 84.81);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.81) -- ( 81.21, 80.88);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.96) -- ( 81.21, 89.89);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.89) -- ( 81.21, 85.96);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.65) -- ( 81.21, 84.57);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.57) -- ( 81.21, 80.65);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.79) -- ( 81.21, 88.72);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.72) -- ( 81.21, 84.79);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.85) -- ( 81.21, 83.78);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.78) -- ( 81.21, 79.85);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.99) -- ( 81.21, 86.91);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.91) -- ( 81.21, 82.99);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.51) -- ( 81.21, 86.44);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.44) -- ( 81.21, 82.51);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.82) -- ( 81.21, 86.74);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.74) -- ( 81.21, 82.82);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.26) -- ( 81.21, 84.18);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.18) -- ( 81.21, 80.26);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.22) -- ( 81.21, 90.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.14) -- ( 81.21, 86.22);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.27) -- ( 81.21, 83.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.19) -- ( 81.21, 79.27);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.16) -- ( 81.21, 83.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.08) -- ( 81.21, 79.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.77) -- ( 81.21, 88.69);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.69) -- ( 81.21, 84.77);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.23) -- ( 81.21, 84.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.16) -- ( 81.21, 80.23);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.04) -- ( 81.21, 82.96);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.96) -- ( 81.21, 79.04);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.14) -- ( 81.21, 91.07);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 91.07) -- ( 81.21, 87.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 78.88) -- ( 81.21, 82.80);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.80) -- ( 81.21, 78.88);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.54) -- ( 81.21, 86.46);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.46) -- ( 81.21, 82.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.43) -- ( 81.21, 88.36);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.36) -- ( 81.21, 84.43);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.01) -- ( 81.21, 85.93);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.93) -- ( 81.21, 82.01);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.55) -- ( 81.21, 87.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.47) -- ( 81.21, 83.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.19) -- ( 81.21, 85.12);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.12) -- ( 81.21, 81.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 97.06) -- ( 81.21,100.99);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29,100.99) -- ( 81.21, 97.06);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.13) -- ( 81.21, 90.06);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.06) -- ( 81.21, 86.13);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.28) -- ( 81.21, 86.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.21) -- ( 81.21, 82.28);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.75) -- ( 81.21, 88.68);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.68) -- ( 81.21, 84.75);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.14) -- ( 81.21, 90.06);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.06) -- ( 81.21, 86.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.28) -- ( 81.21, 83.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.21) -- ( 81.21, 79.28);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.35) -- ( 81.21, 85.28);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.28) -- ( 81.21, 81.35);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.95) -- ( 81.21, 83.87);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.87) -- ( 81.21, 79.95);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.73) -- ( 81.21, 84.66);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.66) -- ( 81.21, 80.73);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 92.83) -- ( 81.21, 96.76);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 96.76) -- ( 81.21, 92.83);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.59) -- ( 81.21, 84.52);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.52) -- ( 81.21, 80.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.31) -- ( 81.21, 92.24);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 92.24) -- ( 81.21, 88.31);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.07) -- ( 81.21, 84.00);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.00) -- ( 81.21, 80.07);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.24) -- ( 81.21, 89.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.16) -- ( 81.21, 85.24);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 95.70) -- ( 81.21, 99.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 99.63) -- ( 81.21, 95.70);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.55) -- ( 81.21, 86.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.47) -- ( 81.21, 82.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.29) -- ( 81.21, 84.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.21) -- ( 81.21, 80.29);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.62) -- ( 81.21, 83.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.55) -- ( 81.21, 79.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 78.78) -- ( 81.21, 82.70);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.70) -- ( 81.21, 78.78);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.17) -- ( 81.21, 83.10);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.10) -- ( 81.21, 79.17);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.00) -- ( 81.21, 82.92);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.92) -- ( 81.21, 79.00);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.09) -- ( 81.21, 91.01);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 91.01) -- ( 81.21, 87.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.15) -- ( 81.21, 88.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.08) -- ( 81.21, 84.15);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.74) -- ( 81.21, 84.66);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.66) -- ( 81.21, 80.74);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.67) -- ( 81.21, 90.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.59) -- ( 81.21, 86.67);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 78.53) -- ( 81.21, 82.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.45) -- ( 81.21, 78.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.85) -- ( 81.21, 84.77);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.77) -- ( 81.21, 80.85);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.25) -- ( 81.21, 84.18);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.18) -- ( 81.21, 80.25);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 92.97) -- ( 81.21, 96.90);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 96.90) -- ( 81.21, 92.97);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.45) -- ( 81.21, 93.38);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 93.38) -- ( 81.21, 89.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.27) -- ( 81.21, 83.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.19) -- ( 81.21, 79.27);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 78.86) -- ( 81.21, 82.79);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.79) -- ( 81.21, 78.86);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.53) -- ( 81.21, 87.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.45) -- ( 81.21, 83.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.02) -- ( 81.21, 90.94);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.94) -- ( 81.21, 87.02);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.13) -- ( 81.21, 89.06);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.06) -- ( 81.21, 85.13);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.62) -- ( 81.21, 88.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.55) -- ( 81.21, 84.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 97.62) -- ( 81.21,101.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29,101.54) -- ( 81.21, 97.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.91) -- ( 81.21, 90.83);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.83) -- ( 81.21, 86.91);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.65) -- ( 81.21, 87.58);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.58) -- ( 81.21, 83.65);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.14) -- ( 81.21, 92.07);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 92.07) -- ( 81.21, 88.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 91.09) -- ( 81.21, 95.02);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 95.02) -- ( 81.21, 91.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 78.39) -- ( 81.21, 82.32);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.32) -- ( 81.21, 78.39);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.95) -- ( 81.21, 85.87);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.87) -- ( 81.21, 81.95);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.34) -- ( 81.21, 86.27);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.27) -- ( 81.21, 82.34);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.97) -- ( 81.21, 85.89);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.89) -- ( 81.21, 81.97);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.45) -- ( 81.21, 92.37);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 92.37) -- ( 81.21, 88.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.38) -- ( 81.21, 90.30);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.30) -- ( 81.21, 86.38);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 98.06) -- ( 81.21,101.98);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29,101.98) -- ( 81.21, 98.06);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.07) -- ( 81.21, 84.00);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.00) -- ( 81.21, 80.07);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.75) -- ( 81.21, 84.67);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.67) -- ( 81.21, 80.75);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.06) -- ( 81.21, 82.98);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.98) -- ( 81.21, 79.06);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.07) -- ( 81.21, 83.00);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.00) -- ( 81.21, 79.07);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29,101.47) -- ( 81.21,105.40);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29,105.40) -- ( 81.21,101.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29,101.45) -- ( 81.21,105.37);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29,105.37) -- ( 81.21,101.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 97.62) -- ( 81.21,101.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29,101.54) -- ( 81.21, 97.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 91.30) -- ( 81.21, 95.22);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 95.22) -- ( 81.21, 91.30);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.28) -- ( 81.21, 93.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 93.21) -- ( 81.21, 89.28);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.70) -- ( 81.21, 93.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 93.63) -- ( 81.21, 89.70);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 95.64) -- ( 81.21, 99.56);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 99.56) -- ( 81.21, 95.64);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.38) -- ( 81.21, 91.31);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 91.31) -- ( 81.21, 87.38);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.67) -- ( 81.21, 85.60);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.60) -- ( 81.21, 81.67);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 78.60) -- ( 81.21, 82.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.53) -- ( 81.21, 78.60);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.22) -- ( 81.21, 83.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.14) -- ( 81.21, 79.22);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.50) -- ( 81.21, 85.43);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.43) -- ( 81.21, 81.50);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.62) -- ( 81.21, 87.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.54) -- ( 81.21, 83.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.82) -- ( 81.21, 83.74);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.74) -- ( 81.21, 79.82);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.50) -- ( 81.21, 86.42);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.42) -- ( 81.21, 82.50);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 78.97) -- ( 81.21, 82.90);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.90) -- ( 81.21, 78.97);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.86) -- ( 81.21, 83.79);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.79) -- ( 81.21, 79.86);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.38) -- ( 81.21, 83.30);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.30) -- ( 81.21, 79.38);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.21) -- ( 81.21, 88.13);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.13) -- ( 81.21, 84.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.65) -- ( 81.21, 86.57);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.57) -- ( 81.21, 82.65);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 78.56) -- ( 81.21, 82.48);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.48) -- ( 81.21, 78.56);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 78.88) -- ( 81.21, 82.80);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.80) -- ( 81.21, 78.88);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.24) -- ( 81.21, 84.17);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.17) -- ( 81.21, 80.24);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.20) -- ( 81.21, 84.12);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.12) -- ( 81.21, 80.20);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.69) -- ( 81.21, 85.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.62) -- ( 81.21, 81.69);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.98) -- ( 81.21, 91.91);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 91.91) -- ( 81.21, 87.98);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.45) -- ( 81.21, 83.37);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.37) -- ( 81.21, 79.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.11) -- ( 81.21, 91.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 91.03) -- ( 81.21, 87.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 95.98) -- ( 81.21, 99.90);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 99.90) -- ( 81.21, 95.98);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.69) -- ( 81.21, 84.61);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.61) -- ( 81.21, 80.69);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.22) -- ( 81.21, 84.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.14) -- ( 81.21, 80.22);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.07) -- ( 81.21, 87.00);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.00) -- ( 81.21, 83.07);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 79.35) -- ( 81.21, 83.28);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.28) -- ( 81.21, 79.35);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.70) -- ( 81.21, 84.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.62) -- ( 81.21, 80.70);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.19) -- ( 81.21, 84.12);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.12) -- ( 81.21, 80.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 80.84) -- ( 81.21, 84.76);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.76) -- ( 81.21, 80.84);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 93.44) -- ( 81.21, 97.37);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 97.37) -- ( 81.21, 93.44);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 78.39) -- ( 81.21, 82.32);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.32) -- ( 81.21, 78.39);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.63) -- ( 81.21, 86.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.55) -- ( 81.21, 82.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 78.52) -- ( 81.21, 82.44);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.44) -- ( 81.21, 78.52);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 79.25, 63.16) -- ( 79.25, 80.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 79.25, 51.75) -- ( 79.25, 36.04);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 68.11, 63.16) --
	( 68.11, 51.75) --
	( 90.39, 51.75) --
	( 90.39, 63.16) --
	( 68.11, 63.16) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 68.11, 57.40) -- ( 90.39, 57.40);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 31.71, 30.69) rectangle ( 97.08,108.01);
\end{scope}
\begin{scope}
\path[clip] (102.58, 30.69) rectangle (167.95,108.01);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (102.58, 30.69) rectangle (167.95,108.01);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (102.58, 36.29) --
	(167.95, 36.29);

\path[draw=drawColor,line width= 0.3pt,line join=round] (102.58, 62.99) --
	(167.95, 62.99);

\path[draw=drawColor,line width= 0.3pt,line join=round] (102.58, 89.69) --
	(167.95, 89.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (102.58, 49.64) --
	(167.95, 49.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (102.58, 76.34) --
	(167.95, 76.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (102.58,103.05) --
	(167.95,103.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.84, 30.69) --
	(114.84,108.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] (135.26, 30.69) --
	(135.26,108.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] (155.69, 30.69) --
	(155.69,108.01);
\definecolor{drawColor}{RGB}{51,51,51}

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 72.46) -- (116.80, 76.39);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.39) -- (116.80, 72.46);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 72.15) -- (116.80, 76.07);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.07) -- (116.80, 72.15);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.39) -- (116.80, 80.31);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 80.31) -- (116.80, 76.39);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 73.62) -- (116.80, 77.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 77.55) -- (116.80, 73.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 73.92) -- (116.80, 77.85);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 77.85) -- (116.80, 73.92);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 71.37) -- (116.80, 75.29);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.29) -- (116.80, 71.37);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.50) -- (116.80, 79.43);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.43) -- (116.80, 75.50);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 72.53) -- (116.80, 76.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.45) -- (116.80, 72.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 71.61) -- (116.80, 75.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.53) -- (116.80, 71.61);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 71.50) -- (116.80, 75.42);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.42) -- (116.80, 71.50);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 73.66) -- (116.80, 77.58);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 77.58) -- (116.80, 73.66);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 72.62) -- (116.80, 76.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.55) -- (116.80, 72.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 72.05) -- (116.80, 75.98);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.98) -- (116.80, 72.05);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.89) -- (116.80, 79.81);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.81) -- (116.80, 75.89);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 73.08) -- (116.80, 77.01);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 77.01) -- (116.80, 73.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 71.47) -- (116.80, 75.40);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.40) -- (116.80, 71.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 72.42) -- (116.80, 76.34);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.34) -- (116.80, 72.42);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 77.83) -- (116.80, 81.75);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 81.75) -- (116.80, 77.83);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 73.04) -- (116.80, 76.96);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.96) -- (116.80, 73.04);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 73.17) -- (116.80, 77.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 77.09) -- (116.80, 73.17);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 72.88) -- (116.80, 76.80);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.80) -- (116.80, 72.88);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 73.74) -- (116.80, 77.67);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 77.67) -- (116.80, 73.74);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 73.35) -- (116.80, 77.28);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 77.28) -- (116.80, 73.35);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 73.52) -- (116.80, 77.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 77.45) -- (116.80, 73.52);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 72.78) -- (116.80, 76.70);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.70) -- (116.80, 72.78);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 73.82) -- (116.80, 77.75);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 77.75) -- (116.80, 73.82);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 73.48) -- (116.80, 77.40);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 77.40) -- (116.80, 73.48);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 73.85) -- (116.80, 77.78);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 77.78) -- (116.80, 73.85);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 72.02) -- (116.80, 75.95);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.95) -- (116.80, 72.02);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 74.08) -- (116.80, 78.00);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.00) -- (116.80, 74.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 73.24) -- (116.80, 77.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 77.16) -- (116.80, 73.24);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 73.28) -- (116.80, 77.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 77.21) -- (116.80, 73.28);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 73.03) -- (116.80, 76.95);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.95) -- (116.80, 73.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 77.08) -- (116.80, 81.00);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 81.00) -- (116.80, 77.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 71.55) -- (116.80, 75.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.47) -- (116.80, 71.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 74.72) -- (116.80, 78.65);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.65) -- (116.80, 74.72);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 80.73) -- (116.80, 84.66);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 84.66) -- (116.80, 80.73);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 71.57) -- (116.80, 75.49);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.49) -- (116.80, 71.57);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 85.24) -- (116.80, 89.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 89.16) -- (116.80, 85.24);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 71.68) -- (116.80, 75.60);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.60) -- (116.80, 71.68);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 74.38) -- (116.80, 78.31);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.31) -- (116.80, 74.38);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 74.54) -- (116.80, 78.46);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.46) -- (116.80, 74.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 73.71) -- (116.80, 77.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 77.63) -- (116.80, 73.71);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.27) -- (116.80, 83.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 83.19) -- (116.80, 79.27);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 72.00) -- (116.80, 75.93);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.93) -- (116.80, 72.00);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 77.60) -- (116.80, 81.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 81.53) -- (116.80, 77.60);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 85.13) -- (116.80, 89.06);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 89.06) -- (116.80, 85.13);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 72.58) -- (116.80, 76.50);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.50) -- (116.80, 72.58);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 73.33) -- (116.80, 77.25);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 77.25) -- (116.80, 73.33);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 81.97) -- (116.80, 85.89);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 85.89) -- (116.80, 81.97);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 71.96) -- (116.80, 75.89);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.89) -- (116.80, 71.96);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 77.11) -- (116.80, 81.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 81.03) -- (116.80, 77.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 97.62) -- (116.80,101.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87,101.54) -- (116.80, 97.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 98.71) -- (116.80,102.64);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87,102.64) -- (116.80, 98.71);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 95.64) -- (116.80, 99.56);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 99.56) -- (116.80, 95.64);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.08) -- (116.80, 80.01);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 80.01) -- (116.80, 76.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 74.91) -- (116.80, 78.83);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.83) -- (116.80, 74.91);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 72.31) -- (116.80, 76.23);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.23) -- (116.80, 72.31);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 72.31) -- (116.80, 76.23);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.23) -- (116.80, 72.31);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 72.63) -- (116.80, 76.56);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.56) -- (116.80, 72.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 72.25) -- (116.80, 76.17);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.17) -- (116.80, 72.25);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 73.54) -- (116.80, 77.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 77.47) -- (116.80, 73.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 80.20) -- (116.80, 84.12);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 84.12) -- (116.80, 80.20);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.14) -- (116.80, 82.06);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 82.06) -- (116.80, 78.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 81.69) -- (116.80, 85.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 85.62) -- (116.80, 81.69);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.79) -- (116.80, 80.71);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 80.71) -- (116.80, 76.79);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 77.97) -- (116.80, 81.89);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 81.89) -- (116.80, 77.97);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 87.11) -- (116.80, 91.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 91.03) -- (116.80, 87.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 95.98) -- (116.80, 99.90);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 99.90) -- (116.80, 95.98);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 80.22) -- (116.80, 84.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 84.14) -- (116.80, 80.22);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 71.47) -- (116.80, 75.39);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.39) -- (116.80, 71.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 87.23) -- (116.80, 91.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 91.16) -- (116.80, 87.23);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 80.84) -- (116.80, 84.76);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 84.76) -- (116.80, 80.84);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.05) -- (116.80, 79.97);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.97) -- (116.80, 76.05);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 72.00) -- (116.80, 75.92);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.92) -- (116.80, 72.00);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.84, 58.20) -- (114.84, 73.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.84, 48.16) -- (114.84, 34.20);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (107.18, 58.20) --
	(107.18, 48.16) --
	(122.50, 48.16) --
	(122.50, 58.20) --
	(107.18, 58.20) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (107.18, 53.11) -- (122.50, 53.11);
\definecolor{drawColor}{RGB}{51,51,51}

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 79.03) -- (137.23, 82.96);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 82.96) -- (137.23, 79.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 82.81) -- (137.23, 86.73);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 86.73) -- (137.23, 82.81);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.76) -- (137.23, 84.68);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.68) -- (137.23, 80.76);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.46) -- (137.23, 84.39);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.39) -- (137.23, 80.46);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 77.23) -- (137.23, 81.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.16) -- (137.23, 77.23);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 76.48) -- (137.23, 80.40);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.40) -- (137.23, 76.48);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 76.86) -- (137.23, 80.78);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.78) -- (137.23, 76.86);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 77.23) -- (137.23, 81.15);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.15) -- (137.23, 77.23);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 78.16) -- (137.23, 82.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 82.09) -- (137.23, 78.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 77.85) -- (137.23, 81.78);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.78) -- (137.23, 77.85);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 82.66) -- (137.23, 86.58);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 86.58) -- (137.23, 82.66);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 77.03) -- (137.23, 80.95);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.95) -- (137.23, 77.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 79.34) -- (137.23, 83.26);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.26) -- (137.23, 79.34);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 79.23) -- (137.23, 83.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.16) -- (137.23, 79.23);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 78.51) -- (137.23, 82.43);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 82.43) -- (137.23, 78.51);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 76.55) -- (137.23, 80.48);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.48) -- (137.23, 76.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 77.98) -- (137.23, 81.91);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.91) -- (137.23, 77.98);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 78.50) -- (137.23, 82.42);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 82.42) -- (137.23, 78.50);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.79) -- (137.23, 84.71);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.71) -- (137.23, 80.79);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 77.79) -- (137.23, 81.71);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.71) -- (137.23, 77.79);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 78.41) -- (137.23, 82.33);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 82.33) -- (137.23, 78.41);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.55) -- (137.23, 85.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 85.47) -- (137.23, 81.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 77.37) -- (137.23, 81.29);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.29) -- (137.23, 77.37);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.19) -- (137.23, 85.12);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 85.12) -- (137.23, 81.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.32) -- (137.23, 85.24);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 85.24) -- (137.23, 81.32);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.04) -- (137.23, 86.96);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 86.96) -- (137.23, 83.04);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.20) -- (137.23, 85.13);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 85.13) -- (137.23, 81.20);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.26) -- (137.23, 85.18);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 85.18) -- (137.23, 81.26);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 77.66) -- (137.23, 81.58);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.58) -- (137.23, 77.66);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.90) -- (137.23, 85.83);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 85.83) -- (137.23, 81.90);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 79.97) -- (137.23, 83.90);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.90) -- (137.23, 79.97);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 77.88) -- (137.23, 81.81);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.81) -- (137.23, 77.88);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 76.54) -- (137.23, 80.46);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.46) -- (137.23, 76.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 78.21) -- (137.23, 82.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 82.14) -- (137.23, 78.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.88) -- (137.23, 84.81);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.81) -- (137.23, 80.88);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.65) -- (137.23, 84.57);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.57) -- (137.23, 80.65);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 82.82) -- (137.23, 86.74);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 86.74) -- (137.23, 82.82);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 78.07) -- (137.23, 81.99);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.99) -- (137.23, 78.07);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 79.16) -- (137.23, 83.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.08) -- (137.23, 79.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 78.16) -- (137.23, 82.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 82.08) -- (137.23, 78.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.23) -- (137.23, 84.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.16) -- (137.23, 80.23);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 79.04) -- (137.23, 82.96);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 82.96) -- (137.23, 79.04);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 78.88) -- (137.23, 82.80);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 82.80) -- (137.23, 78.88);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 76.59) -- (137.23, 80.51);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.51) -- (137.23, 76.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 79.28) -- (137.23, 83.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.21) -- (137.23, 79.28);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 78.28) -- (137.23, 82.20);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 82.20) -- (137.23, 78.28);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 77.75) -- (137.23, 81.68);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.68) -- (137.23, 77.75);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 76.82) -- (137.23, 80.74);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.74) -- (137.23, 76.82);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 95.70) -- (137.23, 99.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 99.63) -- (137.23, 95.70);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.74) -- (137.23, 84.66);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.66) -- (137.23, 80.74);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 78.53) -- (137.23, 82.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 82.45) -- (137.23, 78.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.25) -- (137.23, 84.18);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.18) -- (137.23, 80.25);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 92.97) -- (137.23, 96.90);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 96.90) -- (137.23, 92.97);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 89.45) -- (137.23, 93.38);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 93.38) -- (137.23, 89.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 87.02) -- (137.23, 90.94);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 90.94) -- (137.23, 87.02);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.62) -- (137.23, 88.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 88.55) -- (137.23, 84.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 86.91) -- (137.23, 90.83);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 90.83) -- (137.23, 86.91);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.65) -- (137.23, 87.58);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 87.58) -- (137.23, 83.65);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 88.14) -- (137.23, 92.07);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 92.07) -- (137.23, 88.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 78.39) -- (137.23, 82.32);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 82.32) -- (137.23, 78.39);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.95) -- (137.23, 85.87);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 85.87) -- (137.23, 81.95);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 88.45) -- (137.23, 92.37);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 92.37) -- (137.23, 88.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 86.38) -- (137.23, 90.30);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 90.30) -- (137.23, 86.38);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.75) -- (137.23, 84.67);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.67) -- (137.23, 80.75);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 82.65) -- (137.23, 86.57);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 86.57) -- (137.23, 82.65);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.69) -- (137.23, 84.61);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.61) -- (137.23, 80.69);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 77.07) -- (137.23, 80.99);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.99) -- (137.23, 77.07);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.07) -- (137.23, 87.00);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 87.00) -- (137.23, 83.07);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.61) -- (137.23, 84.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.53) -- (137.23, 80.61);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.19) -- (137.23, 84.12);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.12) -- (137.23, 80.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 78.39) -- (137.23, 82.32);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 82.32) -- (137.23, 78.39);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 82.63) -- (137.23, 86.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 86.55) -- (137.23, 82.63);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (135.26, 61.60) -- (135.26, 78.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (135.26, 50.39) -- (135.26, 35.20);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (127.60, 61.60) --
	(127.60, 50.39) --
	(142.92, 50.39) --
	(142.92, 61.60) --
	(127.60, 61.60) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (127.60, 56.00) -- (142.92, 56.00);
\definecolor{drawColor}{RGB}{51,51,51}

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 81.78) -- (157.65, 85.71);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.71) -- (157.65, 81.78);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 82.35) -- (157.65, 86.27);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.27) -- (157.65, 82.35);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.59) -- (157.65, 91.52);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 91.52) -- (157.65, 87.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 80.62) -- (157.65, 84.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.55) -- (157.65, 80.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 83.17) -- (157.65, 87.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.09) -- (157.65, 83.17);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 80.47) -- (157.65, 84.39);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.39) -- (157.65, 80.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.88) -- (157.65, 90.81);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.81) -- (157.65, 86.88);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.54) -- (157.65, 88.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.47) -- (157.65, 84.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.09) -- (157.65, 91.01);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 91.01) -- (157.65, 87.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 82.58) -- (157.65, 86.51);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.51) -- (157.65, 82.58);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.17) -- (157.65, 90.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.09) -- (157.65, 86.17);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.00) -- (157.65, 89.93);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.93) -- (157.65, 86.00);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 81.62) -- (157.65, 85.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.54) -- (157.65, 81.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.72) -- (157.65, 93.64);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 93.64) -- (157.65, 89.72);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.29) -- (157.65, 88.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.21) -- (157.65, 84.29);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 83.16) -- (157.65, 87.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.08) -- (157.65, 83.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 80.29) -- (157.65, 84.22);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.22) -- (157.65, 80.29);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.11) -- (157.65, 94.04);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 94.04) -- (157.65, 90.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 83.43) -- (157.65, 87.35);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.35) -- (157.65, 83.43);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 83.00) -- (157.65, 86.93);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.93) -- (157.65, 83.00);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 83.28) -- (157.65, 87.20);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.20) -- (157.65, 83.28);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.21) -- (157.65, 91.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 91.14) -- (157.65, 87.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 82.19) -- (157.65, 86.12);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.12) -- (157.65, 82.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 81.84) -- (157.65, 85.77);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.77) -- (157.65, 81.84);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 80.45) -- (157.65, 84.37);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.37) -- (157.65, 80.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 81.89) -- (157.65, 85.82);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.82) -- (157.65, 81.89);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 82.26) -- (157.65, 86.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.19) -- (157.65, 82.26);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.31) -- (157.65, 89.24);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.24) -- (157.65, 85.31);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 81.59) -- (157.65, 85.52);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.52) -- (157.65, 81.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 81.56) -- (157.65, 85.49);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.49) -- (157.65, 81.56);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 81.36) -- (157.65, 85.29);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.29) -- (157.65, 81.36);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 81.54) -- (157.65, 85.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.47) -- (157.65, 81.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.87) -- (157.65, 89.79);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.79) -- (157.65, 85.87);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 83.63) -- (157.65, 87.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.55) -- (157.65, 83.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 81.11) -- (157.65, 85.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.03) -- (157.65, 81.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.05) -- (157.65, 88.97);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.97) -- (157.65, 85.05);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.98) -- (157.65, 89.90);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.90) -- (157.65, 85.98);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 83.50) -- (157.65, 87.43);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.43) -- (157.65, 83.50);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 83.91) -- (157.65, 87.84);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.84) -- (157.65, 83.91);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 83.23) -- (157.65, 87.15);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.15) -- (157.65, 83.23);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.91) -- (157.65, 89.83);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.83) -- (157.65, 85.91);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 80.19) -- (157.65, 84.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.11) -- (157.65, 80.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 81.21) -- (157.65, 85.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.14) -- (157.65, 81.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 83.88) -- (157.65, 87.80);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.80) -- (157.65, 83.88);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 81.78) -- (157.65, 85.71);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.71) -- (157.65, 81.78);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 83.05) -- (157.65, 86.97);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.97) -- (157.65, 83.05);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 81.54) -- (157.65, 85.46);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.46) -- (157.65, 81.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.33) -- (157.65, 90.25);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.25) -- (157.65, 86.33);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 82.99) -- (157.65, 86.91);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.91) -- (157.65, 82.99);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.06) -- (157.65, 91.99);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 91.99) -- (157.65, 88.06);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.96) -- (157.65, 89.89);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.89) -- (157.65, 85.96);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.79) -- (157.65, 88.72);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.72) -- (157.65, 84.79);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 82.99) -- (157.65, 86.91);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.91) -- (157.65, 82.99);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 82.51) -- (157.65, 86.44);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.44) -- (157.65, 82.51);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 80.26) -- (157.65, 84.18);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.18) -- (157.65, 80.26);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.22) -- (157.65, 90.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.14) -- (157.65, 86.22);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.77) -- (157.65, 88.69);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.69) -- (157.65, 84.77);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.14) -- (157.65, 91.07);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 91.07) -- (157.65, 87.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 82.54) -- (157.65, 86.46);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.46) -- (157.65, 82.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.43) -- (157.65, 88.36);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.36) -- (157.65, 84.43);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 82.01) -- (157.65, 85.93);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.93) -- (157.65, 82.01);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 83.55) -- (157.65, 87.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.47) -- (157.65, 83.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 81.19) -- (157.65, 85.12);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.12) -- (157.65, 81.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 97.06) -- (157.65,100.99);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73,100.99) -- (157.65, 97.06);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.13) -- (157.65, 90.06);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.06) -- (157.65, 86.13);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 82.28) -- (157.65, 86.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.21) -- (157.65, 82.28);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.75) -- (157.65, 88.68);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.68) -- (157.65, 84.75);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.14) -- (157.65, 90.06);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.06) -- (157.65, 86.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 81.35) -- (157.65, 85.28);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.28) -- (157.65, 81.35);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 92.83) -- (157.65, 96.76);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 96.76) -- (157.65, 92.83);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 80.59) -- (157.65, 84.52);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.52) -- (157.65, 80.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.31) -- (157.65, 92.24);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 92.24) -- (157.65, 88.31);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 82.55) -- (157.65, 86.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.47) -- (157.65, 82.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 80.29) -- (157.65, 84.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.21) -- (157.65, 80.29);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.09) -- (157.65, 91.01);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 91.01) -- (157.65, 87.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.15) -- (157.65, 88.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.08) -- (157.65, 84.15);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.67) -- (157.65, 90.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.59) -- (157.65, 86.67);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 80.85) -- (157.65, 84.77);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.77) -- (157.65, 80.85);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 83.53) -- (157.65, 87.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.45) -- (157.65, 83.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 97.62) -- (157.65,101.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73,101.54) -- (157.65, 97.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 91.09) -- (157.65, 95.02);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 95.02) -- (157.65, 91.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 82.34) -- (157.65, 86.27);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.27) -- (157.65, 82.34);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 98.06) -- (157.65,101.98);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73,101.98) -- (157.65, 98.06);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73,101.49) -- (157.65,105.41);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73,105.41) -- (157.65,101.49);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73,102.54) -- (157.65,106.46);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73,106.46) -- (157.65,102.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73,101.45) -- (157.65,105.38);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73,105.38) -- (157.65,101.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73,101.49) -- (157.65,105.41);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73,105.41) -- (157.65,101.49);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73,101.47) -- (157.65,105.40);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73,105.40) -- (157.65,101.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73,101.45) -- (157.65,105.37);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73,105.37) -- (157.65,101.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 91.30) -- (157.65, 95.22);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 95.22) -- (157.65, 91.30);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.28) -- (157.65, 93.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 93.21) -- (157.65, 89.28);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.70) -- (157.65, 93.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 93.63) -- (157.65, 89.70);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.70) -- (157.65, 89.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.62) -- (157.65, 85.70);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.38) -- (157.65, 91.31);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 91.31) -- (157.65, 87.38);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 81.67) -- (157.65, 85.60);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.60) -- (157.65, 81.67);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 81.50) -- (157.65, 85.43);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.43) -- (157.65, 81.50);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 83.62) -- (157.65, 87.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.54) -- (157.65, 83.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 82.50) -- (157.65, 86.42);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.42) -- (157.65, 82.50);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.21) -- (157.65, 88.13);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.13) -- (157.65, 84.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 80.24) -- (157.65, 84.17);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.17) -- (157.65, 80.24);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.98) -- (157.65, 91.91);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 91.91) -- (157.65, 87.98);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 80.70) -- (157.65, 84.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.62) -- (157.65, 80.70);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 93.44) -- (157.65, 97.37);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 97.37) -- (157.65, 93.44);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (155.69, 64.76) -- (155.69, 82.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (155.69, 53.18) -- (155.69, 39.70);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (148.03, 64.76) --
	(148.03, 53.18) --
	(163.35, 53.18) --
	(163.35, 64.76) --
	(148.03, 64.76) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (148.03, 58.96) -- (163.35, 58.96);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (102.58, 30.69) rectangle (167.95,108.01);
\end{scope}
\begin{scope}
\path[clip] ( 31.71,108.01) rectangle ( 97.08,124.59);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 31.71,108.01) rectangle ( 97.08,124.59);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 64.40,113.27) {Maximum arity};
\end{scope}
\begin{scope}
\path[clip] (102.58,108.01) rectangle (167.95,124.59);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (102.58,108.01) rectangle (167.95,124.59);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (135.26,113.27) {Probabilistic};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (173.45,130.09);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 49.54, 27.94) --
	( 49.54, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 79.25, 27.94) --
	( 79.25, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (173.45,130.09);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 49.54, 19.68) {2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 79.25, 19.68) {3};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (173.45,130.09);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.84, 27.94) --
	(114.84, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (135.26, 27.94) --
	(135.26, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (155.69, 27.94) --
	(155.69, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (173.45,130.09);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (114.84, 19.68) {0.25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (135.26, 19.68) {0.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (155.69, 19.68) {0.75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (173.45,130.09);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 26.76, 46.61) {20};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 26.76, 73.31) {40};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 26.76,100.01) {60};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (173.45,130.09);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96, 49.64) --
	( 31.71, 49.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96, 76.34) --
	( 31.71, 76.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,103.05) --
	( 31.71,103.05);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (173.45,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08, 69.35) {Inference time (s)};
\end{scope}
\end{tikzpicture}