### 2020-05-15 text/paper2/bars.tex
Description of the experiments

% Created by tikzDevice version 0.12.3 on 2020-05-14 17:55:45
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (173.45,130.09);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (173.45,130.09);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( -0.00,  0.00) rectangle (173.45,130.09);
\end{scope}
\begin{scope}
\path[clip] ( 31.71, 30.69) rectangle ( 97.08,108.01);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.71, 30.69) rectangle ( 97.08,108.01);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.71, 36.97) --
	( 97.08, 36.97);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.71, 64.98) --
	( 97.08, 64.98);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.71, 92.98) --
	( 97.08, 92.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.71, 50.98) --
	( 97.08, 50.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.71, 78.98) --
	( 97.08, 78.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.71,106.99) --
	( 97.08,106.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 49.54, 30.69) --
	( 49.54,108.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 79.25, 30.69) --
	( 79.25,108.01);
\definecolor{drawColor}{RGB}{51,51,51}

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.50) -- ( 51.50, 78.42);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 78.42) -- ( 51.50, 74.50);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 73.55) -- ( 51.50, 77.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 77.47) -- ( 51.50, 73.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 83.41) -- ( 51.50, 87.33);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 87.33) -- ( 51.50, 83.41);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 78.42) -- ( 51.50, 82.34);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 82.34) -- ( 51.50, 78.42);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 76.11) -- ( 51.50, 80.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 80.03) -- ( 51.50, 76.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.48) -- ( 51.50, 78.40);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 78.40) -- ( 51.50, 74.48);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 76.90) -- ( 51.50, 80.82);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 80.82) -- ( 51.50, 76.90);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 72.95) -- ( 51.50, 76.87);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 76.87) -- ( 51.50, 72.95);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 77.09) -- ( 51.50, 81.02);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 81.02) -- ( 51.50, 77.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.11) -- ( 51.50, 78.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 78.03) -- ( 51.50, 74.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 76.69) -- ( 51.50, 80.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 80.62) -- ( 51.50, 76.69);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 75.42) -- ( 51.50, 79.34);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 79.34) -- ( 51.50, 75.42);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 73.74) -- ( 51.50, 77.67);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 77.67) -- ( 51.50, 73.74);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.11) -- ( 51.50, 78.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 78.03) -- ( 51.50, 74.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.03) -- ( 51.50, 77.96);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 77.96) -- ( 51.50, 74.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 76.80) -- ( 51.50, 80.72);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 80.72) -- ( 51.50, 76.80);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.58) -- ( 51.50, 78.50);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 78.50) -- ( 51.50, 74.58);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 73.04) -- ( 51.50, 76.96);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 76.96) -- ( 51.50, 73.04);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.11) -- ( 51.50, 78.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 78.03) -- ( 51.50, 74.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.24) -- ( 51.50, 78.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 78.16) -- ( 51.50, 74.24);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 73.75) -- ( 51.50, 77.68);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 77.68) -- ( 51.50, 73.75);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 73.33) -- ( 51.50, 77.25);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 77.25) -- ( 51.50, 73.33);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 75.34) -- ( 51.50, 79.26);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 79.26) -- ( 51.50, 75.34);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 75.66) -- ( 51.50, 79.58);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 79.58) -- ( 51.50, 75.66);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 72.71) -- ( 51.50, 76.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 76.63) -- ( 51.50, 72.71);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 77.76) -- ( 51.50, 81.69);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 81.69) -- ( 51.50, 77.76);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 76.89) -- ( 51.50, 80.82);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 80.82) -- ( 51.50, 76.89);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.43) -- ( 51.50, 78.36);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 78.36) -- ( 51.50, 74.43);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 73.36) -- ( 51.50, 77.29);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 77.29) -- ( 51.50, 73.36);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 73.28) -- ( 51.50, 77.20);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 77.20) -- ( 51.50, 73.28);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 73.81) -- ( 51.50, 77.73);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 77.73) -- ( 51.50, 73.81);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 76.10) -- ( 51.50, 80.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 80.03) -- ( 51.50, 76.10);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 75.24) -- ( 51.50, 79.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 79.16) -- ( 51.50, 75.24);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 78.38) -- ( 51.50, 82.31);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 82.31) -- ( 51.50, 78.38);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 72.51) -- ( 51.50, 76.43);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 76.43) -- ( 51.50, 72.51);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.85) -- ( 51.50, 78.78);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 78.78) -- ( 51.50, 74.85);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 72.89) -- ( 51.50, 76.82);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 76.82) -- ( 51.50, 72.89);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 73.54) -- ( 51.50, 77.46);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 77.46) -- ( 51.50, 73.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 79.88) -- ( 51.50, 83.81);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 83.81) -- ( 51.50, 79.88);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58,102.54) -- ( 51.50,106.46);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58,106.46) -- ( 51.50,102.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 88.89) -- ( 51.50, 92.81);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 92.81) -- ( 51.50, 88.89);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 73.05) -- ( 51.50, 76.97);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 76.97) -- ( 51.50, 73.05);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 73.11) -- ( 51.50, 77.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 77.03) -- ( 51.50, 73.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 73.17) -- ( 51.50, 77.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 77.09) -- ( 51.50, 73.17);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 72.85) -- ( 51.50, 76.77);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 76.77) -- ( 51.50, 72.85);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 80.27) -- ( 51.50, 84.20);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 84.20) -- ( 51.50, 80.27);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 83.55) -- ( 51.50, 87.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 87.47) -- ( 51.50, 83.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 90.50) -- ( 51.50, 94.42);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 94.42) -- ( 51.50, 90.50);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 76.52) -- ( 51.50, 80.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 80.45) -- ( 51.50, 76.52);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 75.93) -- ( 51.50, 79.85);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 79.85) -- ( 51.50, 75.93);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 73.09) -- ( 51.50, 77.02);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 77.02) -- ( 51.50, 73.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 73.21) -- ( 51.50, 77.13);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 77.13) -- ( 51.50, 73.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 77.77) -- ( 51.50, 81.69);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 81.69) -- ( 51.50, 77.77);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 73.08) -- ( 51.50, 77.01);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 77.01) -- ( 51.50, 73.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 90.09) -- ( 51.50, 94.01);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 94.01) -- ( 51.50, 90.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 74.16) -- ( 51.50, 78.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 78.08) -- ( 51.50, 74.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 79.84) -- ( 51.50, 83.77);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 83.77) -- ( 51.50, 79.84);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 78.15) -- ( 51.50, 82.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 82.08) -- ( 51.50, 78.15);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 93.27) -- ( 51.50, 97.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 47.58, 97.19) -- ( 51.50, 93.27);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 49.54, 58.91) -- ( 49.54, 74.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 49.54, 48.64) -- ( 49.54, 34.20);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 38.40, 58.91) --
	( 38.40, 48.64) --
	( 60.68, 48.64) --
	( 60.68, 58.91) --
	( 38.40, 58.91) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 38.40, 53.42) -- ( 60.68, 53.42);
\definecolor{drawColor}{RGB}{51,51,51}

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.78) -- ( 81.21, 88.71);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.71) -- ( 81.21, 84.78);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.38) -- ( 81.21, 89.30);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.30) -- ( 81.21, 85.38);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.95) -- ( 81.21, 85.87);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.87) -- ( 81.21, 81.95);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.90) -- ( 81.21, 85.82);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.82) -- ( 81.21, 81.90);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.85) -- ( 81.21, 89.78);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.78) -- ( 81.21, 85.85);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.71) -- ( 81.21, 87.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.63) -- ( 81.21, 83.71);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.40) -- ( 81.21, 87.32);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.32) -- ( 81.21, 83.40);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.88) -- ( 81.21, 94.80);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 94.80) -- ( 81.21, 90.88);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.57) -- ( 81.21, 87.49);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.49) -- ( 81.21, 83.57);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.24) -- ( 81.21, 90.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.16) -- ( 81.21, 86.24);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.08) -- ( 81.21, 86.01);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.01) -- ( 81.21, 82.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.15) -- ( 81.21, 86.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.08) -- ( 81.21, 82.15);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.14) -- ( 81.21, 86.07);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.07) -- ( 81.21, 82.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.13) -- ( 81.21, 94.06);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 94.06) -- ( 81.21, 90.13);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.67) -- ( 81.21, 91.60);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 91.60) -- ( 81.21, 87.67);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.35) -- ( 81.21, 94.27);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 94.27) -- ( 81.21, 90.35);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.62) -- ( 81.21, 89.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.55) -- ( 81.21, 85.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.38) -- ( 81.21, 93.30);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 93.30) -- ( 81.21, 89.38);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.21) -- ( 81.21, 93.13);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 93.13) -- ( 81.21, 89.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.61) -- ( 81.21, 88.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.53) -- ( 81.21, 84.61);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 93.10) -- ( 81.21, 97.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 97.03) -- ( 81.21, 93.10);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.41) -- ( 81.21, 91.33);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 91.33) -- ( 81.21, 87.41);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.23) -- ( 81.21, 90.15);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.15) -- ( 81.21, 86.23);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.22) -- ( 81.21, 87.15);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.15) -- ( 81.21, 83.22);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 93.52) -- ( 81.21, 97.44);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 97.44) -- ( 81.21, 93.52);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.51) -- ( 81.21, 90.43);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.43) -- ( 81.21, 86.51);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.70) -- ( 81.21, 89.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.63) -- ( 81.21, 85.70);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.06) -- ( 81.21, 89.98);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.98) -- ( 81.21, 86.06);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.27) -- ( 81.21, 86.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.19) -- ( 81.21, 82.27);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.35) -- ( 81.21, 90.28);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.28) -- ( 81.21, 86.35);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.22) -- ( 81.21, 86.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.14) -- ( 81.21, 82.22);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.48) -- ( 81.21, 94.40);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 94.40) -- ( 81.21, 90.48);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.74) -- ( 81.21, 85.66);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.66) -- ( 81.21, 81.74);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.21) -- ( 81.21, 89.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.14) -- ( 81.21, 85.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.82) -- ( 81.21, 85.75);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.75) -- ( 81.21, 81.82);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.84) -- ( 81.21, 88.77);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.77) -- ( 81.21, 84.84);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.38) -- ( 81.21, 87.31);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.31) -- ( 81.21, 83.38);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.11) -- ( 81.21, 86.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.03) -- ( 81.21, 82.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.90) -- ( 81.21, 88.82);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.82) -- ( 81.21, 84.90);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.68) -- ( 81.21, 85.61);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.61) -- ( 81.21, 81.68);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.29) -- ( 81.21, 89.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.21) -- ( 81.21, 85.29);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.48) -- ( 81.21, 92.41);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 92.41) -- ( 81.21, 88.48);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.59) -- ( 81.21, 88.51);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.51) -- ( 81.21, 84.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.55) -- ( 81.21, 88.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.47) -- ( 81.21, 84.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.00) -- ( 81.21, 86.92);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.92) -- ( 81.21, 83.00);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.34) -- ( 81.21, 88.27);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.27) -- ( 81.21, 84.34);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.53) -- ( 81.21, 88.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.45) -- ( 81.21, 84.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.06) -- ( 81.21, 85.98);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.98) -- ( 81.21, 82.06);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.07) -- ( 81.21, 92.99);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 92.99) -- ( 81.21, 89.07);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.56) -- ( 81.21, 86.48);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.48) -- ( 81.21, 82.56);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.74) -- ( 81.21, 87.66);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.66) -- ( 81.21, 83.74);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.71) -- ( 81.21, 90.64);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.64) -- ( 81.21, 86.71);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.08) -- ( 81.21, 88.00);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.00) -- ( 81.21, 84.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.21) -- ( 81.21, 92.13);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 92.13) -- ( 81.21, 88.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.18) -- ( 81.21, 93.10);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 93.10) -- ( 81.21, 89.18);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.95) -- ( 81.21, 86.87);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.87) -- ( 81.21, 82.95);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.59) -- ( 81.21, 90.51);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.51) -- ( 81.21, 86.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.02) -- ( 81.21, 90.94);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.94) -- ( 81.21, 87.02);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.30) -- ( 81.21, 90.22);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.22) -- ( 81.21, 86.30);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.11) -- ( 81.21, 93.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 93.03) -- ( 81.21, 89.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.54) -- ( 81.21, 88.46);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.46) -- ( 81.21, 84.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.69) -- ( 81.21, 86.61);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.61) -- ( 81.21, 82.69);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.89) -- ( 81.21, 85.82);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.82) -- ( 81.21, 81.89);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.73) -- ( 81.21, 85.66);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.66) -- ( 81.21, 81.73);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.11) -- ( 81.21, 87.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.03) -- ( 81.21, 83.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.85) -- ( 81.21, 86.77);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.77) -- ( 81.21, 82.85);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.16) -- ( 81.21, 88.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.09) -- ( 81.21, 84.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.18) -- ( 81.21, 88.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.11) -- ( 81.21, 84.18);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.30) -- ( 81.21, 88.22);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.22) -- ( 81.21, 84.30);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.08) -- ( 81.21, 87.00);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.00) -- ( 81.21, 83.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.10) -- ( 81.21, 90.02);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.02) -- ( 81.21, 86.10);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.17) -- ( 81.21, 88.10);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.10) -- ( 81.21, 84.17);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.98) -- ( 81.21, 90.90);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.90) -- ( 81.21, 86.98);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.23) -- ( 81.21, 88.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.16) -- ( 81.21, 84.23);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.78) -- ( 81.21, 88.70);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.70) -- ( 81.21, 84.78);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.11) -- ( 81.21, 90.04);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.04) -- ( 81.21, 86.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.85) -- ( 81.21, 85.77);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.77) -- ( 81.21, 81.85);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.91) -- ( 81.21, 88.83);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.83) -- ( 81.21, 84.91);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.89) -- ( 81.21, 86.81);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.81) -- ( 81.21, 82.89);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.53) -- ( 81.21, 88.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.45) -- ( 81.21, 84.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.55) -- ( 81.21, 93.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 93.47) -- ( 81.21, 89.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.05) -- ( 81.21, 89.97);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.97) -- ( 81.21, 86.05);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.94) -- ( 81.21, 85.87);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.87) -- ( 81.21, 81.94);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 91.37) -- ( 81.21, 95.29);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 95.29) -- ( 81.21, 91.37);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.84) -- ( 81.21, 87.76);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.76) -- ( 81.21, 83.84);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.17) -- ( 81.21, 93.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 93.09) -- ( 81.21, 89.17);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.59) -- ( 81.21, 87.52);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.52) -- ( 81.21, 83.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.94) -- ( 81.21, 91.87);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 91.87) -- ( 81.21, 87.94);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.76) -- ( 81.21, 86.68);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.68) -- ( 81.21, 82.76);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.05) -- ( 81.21, 89.97);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.97) -- ( 81.21, 86.05);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.55) -- ( 81.21, 89.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.47) -- ( 81.21, 85.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.87) -- ( 81.21, 89.79);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.79) -- ( 81.21, 85.87);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.18) -- ( 81.21, 87.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.11) -- ( 81.21, 83.18);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.44) -- ( 81.21, 93.36);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 93.36) -- ( 81.21, 89.44);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.14) -- ( 81.21, 86.07);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.07) -- ( 81.21, 82.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.03) -- ( 81.21, 85.95);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.95) -- ( 81.21, 82.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.91) -- ( 81.21, 91.84);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 91.84) -- ( 81.21, 87.91);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.16) -- ( 81.21, 87.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.08) -- ( 81.21, 83.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.90) -- ( 81.21, 85.83);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.83) -- ( 81.21, 81.90);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.40) -- ( 81.21, 94.33);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 94.33) -- ( 81.21, 90.40);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.74) -- ( 81.21, 85.66);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.66) -- ( 81.21, 81.74);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.58) -- ( 81.21, 89.50);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.50) -- ( 81.21, 85.58);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.56) -- ( 81.21, 91.49);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 91.49) -- ( 81.21, 87.56);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.02) -- ( 81.21, 88.94);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.94) -- ( 81.21, 85.02);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.63) -- ( 81.21, 90.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.55) -- ( 81.21, 86.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.16) -- ( 81.21, 88.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.09) -- ( 81.21, 84.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29,100.81) -- ( 81.21,104.73);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29,104.73) -- ( 81.21,100.81);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.34) -- ( 81.21, 93.27);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 93.27) -- ( 81.21, 89.34);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.30) -- ( 81.21, 89.23);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.23) -- ( 81.21, 85.30);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.90) -- ( 81.21, 91.82);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 91.82) -- ( 81.21, 87.90);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.35) -- ( 81.21, 93.27);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 93.27) -- ( 81.21, 89.35);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.16) -- ( 81.21, 86.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.09) -- ( 81.21, 82.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.33) -- ( 81.21, 88.25);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.25) -- ( 81.21, 84.33);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.86) -- ( 81.21, 86.78);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.78) -- ( 81.21, 82.86);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.68) -- ( 81.21, 87.61);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.61) -- ( 81.21, 83.68);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 96.37) -- ( 81.21,100.29);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29,100.29) -- ( 81.21, 96.37);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.53) -- ( 81.21, 87.46);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.46) -- ( 81.21, 83.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 91.63) -- ( 81.21, 95.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 95.55) -- ( 81.21, 91.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.99) -- ( 81.21, 86.92);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.92) -- ( 81.21, 82.99);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.40) -- ( 81.21, 92.33);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 92.33) -- ( 81.21, 88.40);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 99.38) -- ( 81.21,103.31);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29,103.31) -- ( 81.21, 99.38);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.59) -- ( 81.21, 89.51);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.51) -- ( 81.21, 85.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.21) -- ( 81.21, 87.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.14) -- ( 81.21, 83.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.52) -- ( 81.21, 86.44);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.44) -- ( 81.21, 82.52);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.63) -- ( 81.21, 85.56);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.56) -- ( 81.21, 81.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.05) -- ( 81.21, 85.97);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.97) -- ( 81.21, 82.05);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.99) -- ( 81.21, 86.91);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.91) -- ( 81.21, 82.99);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.70) -- ( 81.21, 87.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.62) -- ( 81.21, 83.70);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.92) -- ( 81.21, 85.85);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.85) -- ( 81.21, 81.92);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.94) -- ( 81.21, 85.87);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.87) -- ( 81.21, 81.94);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29,101.39) -- ( 81.21,105.31);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29,105.31) -- ( 81.21,101.39);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 94.76) -- ( 81.21, 98.69);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 98.69) -- ( 81.21, 94.76);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 92.65) -- ( 81.21, 96.57);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 96.57) -- ( 81.21, 92.65);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 93.09) -- ( 81.21, 97.01);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 97.01) -- ( 81.21, 93.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 99.31) -- ( 81.21,103.23);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29,103.23) -- ( 81.21, 99.31);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.65) -- ( 81.21, 94.58);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 94.58) -- ( 81.21, 90.65);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.67) -- ( 81.21, 88.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.59) -- ( 81.21, 84.67);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.09) -- ( 81.21, 86.01);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.01) -- ( 81.21, 82.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.49) -- ( 81.21, 88.41);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.41) -- ( 81.21, 84.49);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.70) -- ( 81.21, 90.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.63) -- ( 81.21, 86.70);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.72) -- ( 81.21, 86.65);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.65) -- ( 81.21, 82.72);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.53) -- ( 81.21, 89.46);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.46) -- ( 81.21, 85.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.84) -- ( 81.21, 85.76);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.76) -- ( 81.21, 81.84);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.77) -- ( 81.21, 86.69);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.69) -- ( 81.21, 82.77);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.26) -- ( 81.21, 86.18);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.18) -- ( 81.21, 82.26);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.33) -- ( 81.21, 91.25);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 91.25) -- ( 81.21, 87.33);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.69) -- ( 81.21, 89.61);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.61) -- ( 81.21, 85.69);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.73) -- ( 81.21, 85.66);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.66) -- ( 81.21, 81.73);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.17) -- ( 81.21, 87.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.09) -- ( 81.21, 83.17);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.12) -- ( 81.21, 87.04);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.04) -- ( 81.21, 83.12);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.69) -- ( 81.21, 88.61);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.61) -- ( 81.21, 84.69);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 91.28) -- ( 81.21, 95.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 95.21) -- ( 81.21, 91.28);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.34) -- ( 81.21, 86.26);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.26) -- ( 81.21, 82.34);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.37) -- ( 81.21, 94.29);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 94.29) -- ( 81.21, 90.37);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 99.67) -- ( 81.21,103.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29,103.59) -- ( 81.21, 99.67);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.63) -- ( 81.21, 87.56);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.56) -- ( 81.21, 83.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.14) -- ( 81.21, 87.07);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.07) -- ( 81.21, 83.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.14) -- ( 81.21, 90.06);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.06) -- ( 81.21, 86.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.23) -- ( 81.21, 86.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.16) -- ( 81.21, 82.23);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.64) -- ( 81.21, 87.57);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.57) -- ( 81.21, 83.64);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.12) -- ( 81.21, 87.04);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.04) -- ( 81.21, 83.12);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.79) -- ( 81.21, 87.71);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.71) -- ( 81.21, 83.79);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 97.01) -- ( 81.21,100.93);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29,100.93) -- ( 81.21, 97.01);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.67) -- ( 81.21, 89.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.59) -- ( 81.21, 85.67);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.04) -- ( 81.21, 93.97);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 93.97) -- ( 81.21, 90.04);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.39) -- ( 81.21, 93.32);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 93.32) -- ( 81.21, 89.39);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.02) -- ( 81.21, 85.94);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.94) -- ( 81.21, 82.02);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.59) -- ( 81.21, 92.51);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 92.51) -- ( 81.21, 88.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.22) -- ( 81.21, 93.15);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 93.15) -- ( 81.21, 89.22);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 92.53) -- ( 81.21, 96.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 96.45) -- ( 81.21, 92.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.87) -- ( 81.21, 91.79);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 91.79) -- ( 81.21, 87.87);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.02) -- ( 81.21, 85.94);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.94) -- ( 81.21, 82.02);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 91.38) -- ( 81.21, 95.30);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 95.30) -- ( 81.21, 91.38);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.33) -- ( 81.21, 92.25);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 92.25) -- ( 81.21, 88.33);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.83) -- ( 81.21, 92.75);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 92.75) -- ( 81.21, 88.83);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.10) -- ( 81.21, 91.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 91.03) -- ( 81.21, 87.10);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 90.05) -- ( 81.21, 93.98);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 93.98) -- ( 81.21, 90.05);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.15) -- ( 81.21, 89.07);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 89.07) -- ( 81.21, 85.15);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 84.87) -- ( 81.21, 88.79);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 88.79) -- ( 81.21, 84.87);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.64) -- ( 81.21, 85.56);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.56) -- ( 81.21, 81.64);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.16) -- ( 81.21, 86.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.08) -- ( 81.21, 82.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.79) -- ( 81.21, 87.72);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.72) -- ( 81.21, 83.79);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 91.15) -- ( 81.21, 95.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 95.08) -- ( 81.21, 91.15);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 83.79) -- ( 81.21, 87.72);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 87.72) -- ( 81.21, 83.79);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 81.70) -- ( 81.21, 85.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 85.63) -- ( 81.21, 81.70);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 82.65) -- ( 81.21, 86.57);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] ( 77.29, 86.57) -- ( 81.21, 82.65);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 79.25, 65.34) -- ( 79.25, 83.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 79.25, 53.18) -- ( 79.25, 36.71);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 68.11, 65.34) --
	( 68.11, 53.18) --
	( 90.39, 53.18) --
	( 90.39, 65.34) --
	( 68.11, 65.34) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 68.11, 59.03) -- ( 90.39, 59.03);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 31.71, 30.69) rectangle ( 97.08,108.01);
\end{scope}
\begin{scope}
\path[clip] (102.58, 30.69) rectangle (167.95,108.01);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (102.58, 30.69) rectangle (167.95,108.01);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (102.58, 36.97) --
	(167.95, 36.97);

\path[draw=drawColor,line width= 0.3pt,line join=round] (102.58, 64.98) --
	(167.95, 64.98);

\path[draw=drawColor,line width= 0.3pt,line join=round] (102.58, 92.98) --
	(167.95, 92.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (102.58, 50.98) --
	(167.95, 50.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (102.58, 78.98) --
	(167.95, 78.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (102.58,106.99) --
	(167.95,106.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.84, 30.69) --
	(114.84,108.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] (135.26, 30.69) --
	(135.26,108.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] (155.69, 30.69) --
	(155.69,108.01);
\definecolor{drawColor}{RGB}{51,51,51}

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.01) -- (116.80, 78.93);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.93) -- (116.80, 75.01);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 74.68) -- (116.80, 78.60);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.60) -- (116.80, 74.68);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.12) -- (116.80, 83.05);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 83.05) -- (116.80, 79.12);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.22) -- (116.80, 80.15);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 80.15) -- (116.80, 76.22);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.54) -- (116.80, 80.46);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 80.46) -- (116.80, 76.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.20) -- (116.80, 82.12);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 82.12) -- (116.80, 78.20);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.08) -- (116.80, 79.00);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.00) -- (116.80, 75.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 74.11) -- (116.80, 78.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.03) -- (116.80, 74.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.26) -- (116.80, 80.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 80.19) -- (116.80, 76.26);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.17) -- (116.80, 79.10);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.10) -- (116.80, 75.17);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 74.58) -- (116.80, 78.50);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.50) -- (116.80, 74.58);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.60) -- (116.80, 82.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 82.53) -- (116.80, 78.60);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.66) -- (116.80, 79.58);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.58) -- (116.80, 75.66);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 74.96) -- (116.80, 78.89);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.89) -- (116.80, 74.96);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 80.63) -- (116.80, 84.56);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 84.56) -- (116.80, 80.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.61) -- (116.80, 79.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.54) -- (116.80, 75.61);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.75) -- (116.80, 79.67);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.67) -- (116.80, 75.75);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.44) -- (116.80, 79.36);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.36) -- (116.80, 75.44);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.35) -- (116.80, 80.27);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 80.27) -- (116.80, 76.35);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.94) -- (116.80, 79.87);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.87) -- (116.80, 75.94);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.12) -- (116.80, 80.05);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 80.05) -- (116.80, 76.12);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.34) -- (116.80, 79.26);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.26) -- (116.80, 75.34);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.44) -- (116.80, 80.36);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 80.36) -- (116.80, 76.44);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.07) -- (116.80, 80.00);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 80.00) -- (116.80, 76.07);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.47) -- (116.80, 80.39);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 80.39) -- (116.80, 76.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 74.55) -- (116.80, 78.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.47) -- (116.80, 74.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.70) -- (116.80, 80.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 80.63) -- (116.80, 76.70);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.82) -- (116.80, 79.74);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.74) -- (116.80, 75.82);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.87) -- (116.80, 79.79);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.79) -- (116.80, 75.87);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.60) -- (116.80, 79.52);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.52) -- (116.80, 75.60);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.85) -- (116.80, 83.77);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 83.77) -- (116.80, 79.85);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 77.38) -- (116.80, 81.30);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 81.30) -- (116.80, 77.38);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 83.68) -- (116.80, 87.61);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 87.61) -- (116.80, 83.68);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 88.40) -- (116.80, 92.33);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 92.33) -- (116.80, 88.40);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 74.19) -- (116.80, 78.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.11) -- (116.80, 74.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 74.48) -- (116.80, 78.41);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.41) -- (116.80, 74.48);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.88) -- (116.80, 83.81);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 83.81) -- (116.80, 79.88);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87,101.39) -- (116.80,105.31);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87,105.31) -- (116.80,101.39);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87,102.54) -- (116.80,106.46);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87,106.46) -- (116.80,102.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 99.31) -- (116.80,103.23);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87,103.23) -- (116.80, 99.31);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.81) -- (116.80, 82.73);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 82.73) -- (116.80, 78.81);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 77.57) -- (116.80, 81.50);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 81.50) -- (116.80, 77.57);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 74.85) -- (116.80, 78.77);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.77) -- (116.80, 74.85);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 74.84) -- (116.80, 78.77);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.77) -- (116.80, 74.84);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.19) -- (116.80, 79.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.11) -- (116.80, 75.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 74.78) -- (116.80, 78.70);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.70) -- (116.80, 74.78);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.14) -- (116.80, 80.06);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 80.06) -- (116.80, 76.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 83.12) -- (116.80, 87.04);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 87.04) -- (116.80, 83.12);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 80.96) -- (116.80, 84.89);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 84.89) -- (116.80, 80.96);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 84.69) -- (116.80, 88.61);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 88.61) -- (116.80, 84.69);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.54) -- (116.80, 83.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 83.47) -- (116.80, 79.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 80.78) -- (116.80, 84.71);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 84.71) -- (116.80, 80.78);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 90.37) -- (116.80, 94.29);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 94.29) -- (116.80, 90.37);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 99.67) -- (116.80,103.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87,103.59) -- (116.80, 99.67);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 83.14) -- (116.80, 87.07);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 87.07) -- (116.80, 83.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 90.50) -- (116.80, 94.42);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 94.42) -- (116.80, 90.50);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 83.79) -- (116.80, 87.71);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 87.71) -- (116.80, 83.79);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.77) -- (116.80, 82.70);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 82.70) -- (116.80, 78.77);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 74.52) -- (116.80, 78.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.45) -- (116.80, 74.52);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 74.78) -- (116.80, 78.71);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.71) -- (116.80, 74.78);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.33) -- (116.80, 79.25);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.25) -- (116.80, 75.33);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.24) -- (116.80, 82.17);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 82.17) -- (116.80, 78.24);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.25) -- (116.80, 79.18);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.18) -- (116.80, 75.25);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.55) -- (116.80, 79.48);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.48) -- (116.80, 75.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.18) -- (116.80, 79.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.11) -- (116.80, 75.18);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.22) -- (116.80, 82.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 82.14) -- (116.80, 78.22);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 77.74) -- (116.80, 81.67);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 81.67) -- (116.80, 77.74);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 75.24) -- (116.80, 79.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.16) -- (116.80, 75.24);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 85.15) -- (116.80, 89.07);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 89.07) -- (116.80, 85.15);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 76.03) -- (116.80, 79.96);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 79.96) -- (116.80, 76.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.69) -- (116.80, 82.61);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 82.61) -- (116.80, 78.69);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 91.15) -- (116.80, 95.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 95.08) -- (116.80, 91.15);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 78.15) -- (116.80, 82.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (112.87, 82.08) -- (116.80, 78.15);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.84, 60.05) -- (114.84, 76.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.84, 49.37) -- (114.84, 34.20);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (107.18, 60.05) --
	(107.18, 49.37) --
	(122.50, 49.37) --
	(122.50, 60.05) --
	(107.18, 60.05) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (107.18, 54.57) -- (122.50, 54.57);
\definecolor{drawColor}{RGB}{51,51,51}

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.90) -- (137.23, 85.82);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 85.82) -- (137.23, 81.90);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 85.85) -- (137.23, 89.78);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 89.78) -- (137.23, 85.85);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.71) -- (137.23, 87.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 87.63) -- (137.23, 83.71);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.40) -- (137.23, 87.32);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 87.32) -- (137.23, 83.40);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.01) -- (137.23, 83.94);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.94) -- (137.23, 80.01);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 79.22) -- (137.23, 83.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.14) -- (137.23, 79.22);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 79.62) -- (137.23, 83.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.54) -- (137.23, 79.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.00) -- (137.23, 83.93);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.93) -- (137.23, 80.00);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.99) -- (137.23, 84.91);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.91) -- (137.23, 80.99);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.66) -- (137.23, 84.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.59) -- (137.23, 80.66);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 85.70) -- (137.23, 89.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 89.63) -- (137.23, 85.70);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 79.80) -- (137.23, 83.72);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.72) -- (137.23, 79.80);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 82.22) -- (137.23, 86.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 86.14) -- (137.23, 82.22);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 82.11) -- (137.23, 86.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 86.03) -- (137.23, 82.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.35) -- (137.23, 85.27);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 85.27) -- (137.23, 81.35);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 79.30) -- (137.23, 83.22);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.22) -- (137.23, 79.30);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.80) -- (137.23, 84.72);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.72) -- (137.23, 80.80);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.34) -- (137.23, 85.26);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 85.26) -- (137.23, 81.34);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 79.00) -- (137.23, 82.92);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 82.92) -- (137.23, 79.00);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.74) -- (137.23, 87.66);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 87.66) -- (137.23, 83.74);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.59) -- (137.23, 84.52);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.52) -- (137.23, 80.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.24) -- (137.23, 85.17);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 85.17) -- (137.23, 81.24);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.54) -- (137.23, 88.46);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 88.46) -- (137.23, 84.54);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.15) -- (137.23, 84.07);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.07) -- (137.23, 80.15);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.16) -- (137.23, 88.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 88.09) -- (137.23, 84.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.30) -- (137.23, 88.22);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 88.22) -- (137.23, 84.30);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 86.10) -- (137.23, 90.02);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 90.02) -- (137.23, 86.10);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.17) -- (137.23, 88.10);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 88.10) -- (137.23, 84.17);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.23) -- (137.23, 88.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 88.16) -- (137.23, 84.23);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.45) -- (137.23, 84.38);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.38) -- (137.23, 80.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.91) -- (137.23, 88.83);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 88.83) -- (137.23, 84.91);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 82.89) -- (137.23, 86.81);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 86.81) -- (137.23, 82.89);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.69) -- (137.23, 84.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.62) -- (137.23, 80.69);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 79.28) -- (137.23, 83.20);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.20) -- (137.23, 79.28);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.04) -- (137.23, 84.96);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.96) -- (137.23, 81.04);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.84) -- (137.23, 87.76);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 87.76) -- (137.23, 83.84);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.59) -- (137.23, 87.52);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 87.52) -- (137.23, 83.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 85.87) -- (137.23, 89.79);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 89.79) -- (137.23, 85.87);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.89) -- (137.23, 84.81);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.81) -- (137.23, 80.89);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 82.03) -- (137.23, 85.95);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 85.95) -- (137.23, 82.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.98) -- (137.23, 84.90);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.90) -- (137.23, 80.98);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.16) -- (137.23, 87.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 87.08) -- (137.23, 83.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.90) -- (137.23, 85.83);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 85.83) -- (137.23, 81.90);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.74) -- (137.23, 85.66);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 85.66) -- (137.23, 81.74);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 79.12) -- (137.23, 83.05);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.05) -- (137.23, 79.12);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 79.34) -- (137.23, 83.26);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.26) -- (137.23, 79.34);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 82.16) -- (137.23, 86.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 86.09) -- (137.23, 82.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.11) -- (137.23, 85.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 85.03) -- (137.23, 81.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.56) -- (137.23, 84.48);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.48) -- (137.23, 80.56);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 79.58) -- (137.23, 83.50);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.50) -- (137.23, 79.58);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 99.38) -- (137.23,103.31);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30,103.31) -- (137.23, 99.38);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 79.19) -- (137.23, 83.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.11) -- (137.23, 79.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.70) -- (137.23, 87.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 87.62) -- (137.23, 83.70);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 85.69) -- (137.23, 89.61);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 89.61) -- (137.23, 85.69);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.63) -- (137.23, 87.56);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 87.56) -- (137.23, 83.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 79.84) -- (137.23, 83.76);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.76) -- (137.23, 79.84);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 86.14) -- (137.23, 90.06);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 90.06) -- (137.23, 86.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.55) -- (137.23, 87.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 87.47) -- (137.23, 83.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.12) -- (137.23, 87.04);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 87.04) -- (137.23, 83.12);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.23) -- (137.23, 85.15);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 85.15) -- (137.23, 81.23);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 85.67) -- (137.23, 89.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 89.59) -- (137.23, 85.67);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.56) -- (137.23, 85.49);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 85.49) -- (137.23, 81.56);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 80.35) -- (137.23, 84.28);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 84.28) -- (137.23, 80.35);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 90.09) -- (137.23, 94.01);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 94.01) -- (137.23, 90.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 79.84) -- (137.23, 83.77);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 83.77) -- (137.23, 79.84);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 81.70) -- (137.23, 85.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 85.63) -- (137.23, 81.70);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 82.65) -- (137.23, 86.57);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (133.30, 86.57) -- (137.23, 82.65);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (135.26, 63.50) -- (135.26, 80.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (135.26, 51.90) -- (135.26, 35.59);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (127.60, 63.50) --
	(127.60, 51.90) --
	(142.92, 51.90) --
	(142.92, 63.50) --
	(127.60, 63.50) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (127.60, 57.67) -- (142.92, 57.67);
\definecolor{drawColor}{RGB}{51,51,51}

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.78) -- (157.65, 88.71);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.71) -- (157.65, 84.78);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.38) -- (157.65, 89.30);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.30) -- (157.65, 85.38);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.88) -- (157.65, 94.80);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 94.80) -- (157.65, 90.88);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.24) -- (157.65, 90.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.16) -- (157.65, 86.24);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.13) -- (157.65, 94.06);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 94.06) -- (157.65, 90.13);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.67) -- (157.65, 91.60);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 91.60) -- (157.65, 87.67);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.35) -- (157.65, 94.27);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 94.27) -- (157.65, 90.35);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.62) -- (157.65, 89.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.55) -- (157.65, 85.62);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.38) -- (157.65, 93.30);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 93.30) -- (157.65, 89.38);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.21) -- (157.65, 93.13);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 93.13) -- (157.65, 89.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.61) -- (157.65, 88.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.53) -- (157.65, 84.61);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 93.10) -- (157.65, 97.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 97.03) -- (157.65, 93.10);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.41) -- (157.65, 91.33);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 91.33) -- (157.65, 87.41);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.23) -- (157.65, 90.15);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.15) -- (157.65, 86.23);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 93.52) -- (157.65, 97.44);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 97.44) -- (157.65, 93.52);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.51) -- (157.65, 90.43);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.43) -- (157.65, 86.51);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.06) -- (157.65, 89.98);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.98) -- (157.65, 86.06);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.35) -- (157.65, 90.28);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.28) -- (157.65, 86.35);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.48) -- (157.65, 94.40);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 94.40) -- (157.65, 90.48);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.21) -- (157.65, 89.14);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.14) -- (157.65, 85.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.84) -- (157.65, 88.77);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.77) -- (157.65, 84.84);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.90) -- (157.65, 88.82);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.82) -- (157.65, 84.90);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.29) -- (157.65, 89.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.21) -- (157.65, 85.29);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.48) -- (157.65, 92.41);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 92.41) -- (157.65, 88.48);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.59) -- (157.65, 88.51);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.51) -- (157.65, 84.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.55) -- (157.65, 88.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.47) -- (157.65, 84.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.34) -- (157.65, 88.27);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.27) -- (157.65, 84.34);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.53) -- (157.65, 88.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.45) -- (157.65, 84.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.07) -- (157.65, 92.99);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 92.99) -- (157.65, 89.07);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.71) -- (157.65, 90.64);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.64) -- (157.65, 86.71);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.08) -- (157.65, 88.00);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.00) -- (157.65, 84.08);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.21) -- (157.65, 92.13);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 92.13) -- (157.65, 88.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.18) -- (157.65, 93.10);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 93.10) -- (157.65, 89.18);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.59) -- (157.65, 90.51);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.51) -- (157.65, 86.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.02) -- (157.65, 90.94);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.94) -- (157.65, 87.02);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.30) -- (157.65, 90.22);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.22) -- (157.65, 86.30);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.11) -- (157.65, 93.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 93.03) -- (157.65, 89.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.18) -- (157.65, 88.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.11) -- (157.65, 84.18);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.98) -- (157.65, 90.90);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.90) -- (157.65, 86.98);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.78) -- (157.65, 88.70);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.70) -- (157.65, 84.78);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.11) -- (157.65, 90.04);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.04) -- (157.65, 86.11);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.53) -- (157.65, 88.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.45) -- (157.65, 84.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.55) -- (157.65, 93.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 93.47) -- (157.65, 89.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.05) -- (157.65, 89.97);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.97) -- (157.65, 86.05);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 91.37) -- (157.65, 95.29);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 95.29) -- (157.65, 91.37);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.17) -- (157.65, 93.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 93.09) -- (157.65, 89.17);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.94) -- (157.65, 91.87);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 91.87) -- (157.65, 87.94);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.05) -- (157.65, 89.97);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.97) -- (157.65, 86.05);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.55) -- (157.65, 89.47);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.47) -- (157.65, 85.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.44) -- (157.65, 93.36);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 93.36) -- (157.65, 89.44);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.91) -- (157.65, 91.84);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 91.84) -- (157.65, 87.91);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.40) -- (157.65, 94.33);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 94.33) -- (157.65, 90.40);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.58) -- (157.65, 89.50);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.50) -- (157.65, 85.58);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.56) -- (157.65, 91.49);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 91.49) -- (157.65, 87.56);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.02) -- (157.65, 88.94);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.94) -- (157.65, 85.02);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.63) -- (157.65, 90.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.55) -- (157.65, 86.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.16) -- (157.65, 88.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.09) -- (157.65, 84.16);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73,100.81) -- (157.65,104.73);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73,104.73) -- (157.65,100.81);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.34) -- (157.65, 93.27);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 93.27) -- (157.65, 89.34);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.30) -- (157.65, 89.23);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.23) -- (157.65, 85.30);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.90) -- (157.65, 91.82);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 91.82) -- (157.65, 87.90);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.35) -- (157.65, 93.27);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 93.27) -- (157.65, 89.35);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.33) -- (157.65, 88.25);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.25) -- (157.65, 84.33);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 96.37) -- (157.65,100.29);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73,100.29) -- (157.65, 96.37);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 91.63) -- (157.65, 95.55);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 95.55) -- (157.65, 91.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.59) -- (157.65, 89.51);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.51) -- (157.65, 85.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 94.76) -- (157.65, 98.69);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 98.69) -- (157.65, 94.76);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 92.65) -- (157.65, 96.57);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 96.57) -- (157.65, 92.65);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 93.09) -- (157.65, 97.01);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 97.01) -- (157.65, 93.09);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.89) -- (157.65, 92.81);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 92.81) -- (157.65, 88.89);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.65) -- (157.65, 94.58);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 94.58) -- (157.65, 90.65);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.67) -- (157.65, 88.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.59) -- (157.65, 84.67);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.49) -- (157.65, 88.41);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.41) -- (157.65, 84.49);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 86.70) -- (157.65, 90.63);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.63) -- (157.65, 86.70);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 85.53) -- (157.65, 89.46);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.46) -- (157.65, 85.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.33) -- (157.65, 91.25);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 91.25) -- (157.65, 87.33);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 91.28) -- (157.65, 95.21);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 95.21) -- (157.65, 91.28);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 97.01) -- (157.65,100.93);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73,100.93) -- (157.65, 97.01);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.04) -- (157.65, 93.97);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 93.97) -- (157.65, 90.04);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.39) -- (157.65, 93.32);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 93.32) -- (157.65, 89.39);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.59) -- (157.65, 92.51);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 92.51) -- (157.65, 88.59);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 89.22) -- (157.65, 93.15);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 93.15) -- (157.65, 89.22);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 92.53) -- (157.65, 96.45);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 96.45) -- (157.65, 92.53);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.87) -- (157.65, 91.79);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 91.79) -- (157.65, 87.87);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 91.38) -- (157.65, 95.30);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 95.30) -- (157.65, 91.38);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.33) -- (157.65, 92.25);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 92.25) -- (157.65, 88.33);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.83) -- (157.65, 92.75);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 92.75) -- (157.65, 88.83);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 87.10) -- (157.65, 91.03);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 91.03) -- (157.65, 87.10);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 90.05) -- (157.65, 93.98);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 93.98) -- (157.65, 90.05);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 84.87) -- (157.65, 88.79);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 88.79) -- (157.65, 84.87);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 93.27) -- (157.65, 97.19);

\path[draw=drawColor,draw opacity=0.25,line width= 0.4pt,line join=round,line cap=round] (153.73, 97.19) -- (157.65, 93.27);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (155.69, 67.06) -- (155.69, 85.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (155.69, 54.45) -- (155.69, 41.13);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (148.03, 67.06) --
	(148.03, 54.45) --
	(163.35, 54.45) --
	(163.35, 67.06) --
	(148.03, 67.06) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (148.03, 60.57) -- (163.35, 60.57);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (102.58, 30.69) rectangle (167.95,108.01);
\end{scope}
\begin{scope}
\path[clip] ( 31.71,108.01) rectangle ( 97.08,124.59);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 31.71,108.01) rectangle ( 97.08,124.59);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 64.40,113.27) {Maximum arity};
\end{scope}
\begin{scope}
\path[clip] (102.58,108.01) rectangle (167.95,124.59);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (102.58,108.01) rectangle (167.95,124.59);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (135.26,113.27) {Probabilistic};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (173.45,130.09);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 49.54, 27.94) --
	( 49.54, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 79.25, 27.94) --
	( 79.25, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (173.45,130.09);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 49.54, 19.68) {2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 79.25, 19.68) {3};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (173.45,130.09);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.84, 27.94) --
	(114.84, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (135.26, 27.94) --
	(135.26, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (155.69, 27.94) --
	(155.69, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (173.45,130.09);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (114.84, 19.68) {0.25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (135.26, 19.68) {0.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (155.69, 19.68) {0.75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (173.45,130.09);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 26.76, 47.95) {20};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 26.76, 75.95) {40};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 26.76,103.95) {60};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (173.45,130.09);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96, 50.98) --
	( 31.71, 50.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96, 78.98) --
	( 31.71, 78.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,106.99) --
	( 31.71,106.99);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (173.45,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08, 69.35) {Inference time (s)};
\end{scope}
\end{tikzpicture}