### 2020-05-30 text/paper2/algorithms.tex
New plots

% Created by tikzDevice version 0.12.3 on 2020-05-30 13:08:25
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (346.90,346.90);
\begin{scope}
\path[clip] ( 12.91,260.17) rectangle (124.24,346.90);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 12.91,260.17) rectangle (124.24,346.90);
\end{scope}
\begin{scope}
\path[clip] ( 34.60,290.86) rectangle (118.74,341.40);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 34.60,290.86) rectangle (118.74,341.40);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.60,295.44) --
	(118.74,295.44);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.60,307.95) --
	(118.74,307.95);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.60,320.47) --
	(118.74,320.47);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.60,332.98) --
	(118.74,332.98);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 51.17,290.86) --
	( 51.17,341.40);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 76.67,290.86) --
	( 76.67,341.40);

\path[draw=drawColor,line width= 0.3pt,line join=round] (102.17,290.86) --
	(102.17,341.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.60,301.70) --
	(118.74,301.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.60,314.21) --
	(118.74,314.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.60,326.72) --
	(118.74,326.72);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.60,339.24) --
	(118.74,339.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 38.43,290.86) --
	( 38.43,341.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 63.92,290.86) --
	( 63.92,341.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 89.42,290.86) --
	( 89.42,341.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.91,290.86) --
	(114.91,341.40);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 38.43,313.83) --
	( 63.92,305.69) --
	(114.91,313.71);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 38.43,295.89) --
	( 63.92,293.16) --
	(114.91,303.91);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 38.43,305.50) --
	( 63.92,302.49) --
	(114.91,313.03);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 38.43,328.83) --
	( 63.92,329.45) --
	(114.91,339.10);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt ,line join=round] ( 38.43,320.24) --
	( 63.92,323.43) --
	(114.91,332.99);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt on 4pt off 3pt ,line join=round] ( 38.43,330.32) --
	( 63.92,328.96) --
	(114.91,338.63);
\definecolor{drawColor}{RGB}{27,158,119}
\definecolor{fillColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 38.43,313.83) circle (  1.96);
\definecolor{drawColor}{RGB}{217,95,2}
\definecolor{fillColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 38.43,295.89) circle (  1.96);
\definecolor{drawColor}{RGB}{117,112,179}
\definecolor{fillColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 38.43,305.50) circle (  1.96);
\definecolor{drawColor}{RGB}{231,41,138}
\definecolor{fillColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 38.43,328.83) circle (  1.96);
\definecolor{drawColor}{RGB}{102,166,30}
\definecolor{fillColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 38.43,320.24) circle (  1.96);
\definecolor{drawColor}{RGB}{230,171,2}
\definecolor{fillColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 38.43,330.32) circle (  1.96);
\definecolor{drawColor}{RGB}{27,158,119}
\definecolor{fillColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 63.92,305.69) circle (  1.96);
\definecolor{drawColor}{RGB}{217,95,2}
\definecolor{fillColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 63.92,293.16) circle (  1.96);
\definecolor{drawColor}{RGB}{117,112,179}
\definecolor{fillColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 63.92,302.49) circle (  1.96);
\definecolor{drawColor}{RGB}{231,41,138}
\definecolor{fillColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 63.92,329.45) circle (  1.96);
\definecolor{drawColor}{RGB}{102,166,30}
\definecolor{fillColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 63.92,323.43) circle (  1.96);
\definecolor{drawColor}{RGB}{230,171,2}
\definecolor{fillColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 63.92,328.96) circle (  1.96);
\definecolor{drawColor}{RGB}{27,158,119}
\definecolor{fillColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (114.91,313.71) circle (  1.96);
\definecolor{drawColor}{RGB}{217,95,2}
\definecolor{fillColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (114.91,303.91) circle (  1.96);
\definecolor{drawColor}{RGB}{117,112,179}
\definecolor{fillColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (114.91,313.03) circle (  1.96);
\definecolor{drawColor}{RGB}{231,41,138}
\definecolor{fillColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (114.91,339.10) circle (  1.96);
\definecolor{drawColor}{RGB}{102,166,30}
\definecolor{fillColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (114.91,332.99) circle (  1.96);
\definecolor{drawColor}{RGB}{230,171,2}
\definecolor{fillColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (114.91,338.63) circle (  1.96);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 34.60,290.86) rectangle (118.74,341.40);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.65,298.67) {7.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.65,311.18) {7.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.65,323.69) {8.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.65,336.21) {8.5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.85,301.70) --
	( 34.60,301.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.85,314.21) --
	( 34.60,314.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.85,326.72) --
	( 34.60,326.72);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.85,339.24) --
	( 34.60,339.24);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 38.43,288.11) --
	( 38.43,290.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 63.92,288.11) --
	( 63.92,290.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 89.42,288.11) --
	( 89.42,290.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.91,288.11) --
	(114.91,290.86);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 38.43,279.85) {2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 63.92,279.85) {4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 89.42,279.85) {6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (114.91,279.85) {8};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 76.67,267.81) {Predicates};
\end{scope}
\begin{scope}
\path[clip] (124.24,260.17) rectangle (235.57,346.90);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (124.24,260.17) rectangle (235.57,346.90);
\end{scope}
\begin{scope}
\path[clip] (145.93,290.86) rectangle (230.07,341.40);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (145.93,290.86) rectangle (230.07,341.40);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (145.93,298.88) --
	(230.07,298.88);

\path[draw=drawColor,line width= 0.3pt,line join=round] (145.93,309.45) --
	(230.07,309.45);

\path[draw=drawColor,line width= 0.3pt,line join=round] (145.93,320.03) --
	(230.07,320.03);

\path[draw=drawColor,line width= 0.3pt,line join=round] (145.93,330.60) --
	(230.07,330.60);

\path[draw=drawColor,line width= 0.3pt,line join=round] (145.93,341.17) --
	(230.07,341.17);

\path[draw=drawColor,line width= 0.3pt,line join=round] (162.50,290.86) --
	(162.50,341.40);

\path[draw=drawColor,line width= 0.3pt,line join=round] (188.00,290.86) --
	(188.00,341.40);

\path[draw=drawColor,line width= 0.3pt,line join=round] (213.49,290.86) --
	(213.49,341.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (145.93,293.59) --
	(230.07,293.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (145.93,304.16) --
	(230.07,304.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (145.93,314.74) --
	(230.07,314.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (145.93,325.31) --
	(230.07,325.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (145.93,335.89) --
	(230.07,335.89);

\path[draw=drawColor,line width= 0.6pt,line join=round] (149.75,290.86) --
	(149.75,341.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (175.25,290.86) --
	(175.25,341.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (200.75,290.86) --
	(200.75,341.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (226.24,290.86) --
	(226.24,341.40);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (149.75,319.74) --
	(175.25,314.24) --
	(226.24,300.18);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (149.75,309.41) --
	(175.25,305.28) --
	(226.24,293.16);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (149.75,316.71) --
	(175.25,313.05) --
	(226.24,301.69);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (149.75,336.45) --
	(175.25,339.10) --
	(226.24,320.12);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt ,line join=round] (149.75,333.77) --
	(175.25,329.89) --
	(226.24,317.60);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt on 4pt off 3pt ,line join=round] (149.75,337.33) --
	(175.25,336.51) --
	(226.24,322.03);
\definecolor{drawColor}{RGB}{27,158,119}
\definecolor{fillColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (149.75,319.74) circle (  1.96);
\definecolor{drawColor}{RGB}{217,95,2}
\definecolor{fillColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (149.75,309.41) circle (  1.96);
\definecolor{drawColor}{RGB}{117,112,179}
\definecolor{fillColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (149.75,316.71) circle (  1.96);
\definecolor{drawColor}{RGB}{231,41,138}
\definecolor{fillColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (149.75,336.45) circle (  1.96);
\definecolor{drawColor}{RGB}{102,166,30}
\definecolor{fillColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (149.75,333.77) circle (  1.96);
\definecolor{drawColor}{RGB}{230,171,2}
\definecolor{fillColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (149.75,337.33) circle (  1.96);
\definecolor{drawColor}{RGB}{27,158,119}
\definecolor{fillColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (175.25,314.24) circle (  1.96);
\definecolor{drawColor}{RGB}{217,95,2}
\definecolor{fillColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (175.25,305.28) circle (  1.96);
\definecolor{drawColor}{RGB}{117,112,179}
\definecolor{fillColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (175.25,313.05) circle (  1.96);
\definecolor{drawColor}{RGB}{231,41,138}
\definecolor{fillColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (175.25,339.10) circle (  1.96);
\definecolor{drawColor}{RGB}{102,166,30}
\definecolor{fillColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (175.25,329.89) circle (  1.96);
\definecolor{drawColor}{RGB}{230,171,2}
\definecolor{fillColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (175.25,336.51) circle (  1.96);
\definecolor{drawColor}{RGB}{27,158,119}
\definecolor{fillColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (226.24,300.18) circle (  1.96);
\definecolor{drawColor}{RGB}{217,95,2}
\definecolor{fillColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (226.24,293.16) circle (  1.96);
\definecolor{drawColor}{RGB}{117,112,179}
\definecolor{fillColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (226.24,301.69) circle (  1.96);
\definecolor{drawColor}{RGB}{231,41,138}
\definecolor{fillColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (226.24,320.12) circle (  1.96);
\definecolor{drawColor}{RGB}{102,166,30}
\definecolor{fillColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (226.24,317.60) circle (  1.96);
\definecolor{drawColor}{RGB}{230,171,2}
\definecolor{fillColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (226.24,322.03) circle (  1.96);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (145.93,290.86) rectangle (230.07,341.40);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (140.98,290.56) {6.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (140.98,301.13) {7.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (140.98,311.71) {7.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (140.98,322.28) {8.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (140.98,332.86) {8.5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (143.18,293.59) --
	(145.93,293.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (143.18,304.16) --
	(145.93,304.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (143.18,314.74) --
	(145.93,314.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (143.18,325.31) --
	(145.93,325.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (143.18,335.89) --
	(145.93,335.89);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (149.75,288.11) --
	(149.75,290.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] (175.25,288.11) --
	(175.25,290.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] (200.75,288.11) --
	(200.75,290.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] (226.24,288.11) --
	(226.24,290.86);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (149.75,279.85) {2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (175.25,279.85) {4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (200.75,279.85) {6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (226.24,279.85) {8};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (188.00,267.81) {Variables};
\end{scope}
\begin{scope}
\path[clip] (235.57,260.17) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (235.57,260.17) rectangle (346.90,346.90);
\end{scope}
\begin{scope}
\path[clip] (250.42,290.86) rectangle (341.40,341.40);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (250.42,290.86) rectangle (341.40,341.40);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (250.42,301.05) --
	(341.40,301.05);

\path[draw=drawColor,line width= 0.3pt,line join=round] (250.42,314.08) --
	(341.40,314.08);

\path[draw=drawColor,line width= 0.3pt,line join=round] (250.42,327.12) --
	(341.40,327.12);

\path[draw=drawColor,line width= 0.3pt,line join=round] (250.42,340.15) --
	(341.40,340.15);

\path[draw=drawColor,line width= 0.3pt,line join=round] (268.34,290.86) --
	(268.34,341.40);

\path[draw=drawColor,line width= 0.3pt,line join=round] (295.91,290.86) --
	(295.91,341.40);

\path[draw=drawColor,line width= 0.3pt,line join=round] (323.48,290.86) --
	(323.48,341.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.42,294.53) --
	(341.40,294.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.42,307.57) --
	(341.40,307.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.42,320.60) --
	(341.40,320.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.42,333.64) --
	(341.40,333.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.55,290.86) --
	(254.55,341.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (282.12,290.86) --
	(282.12,341.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (309.69,290.86) --
	(309.69,341.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (337.26,290.86) --
	(337.26,341.40);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.55,313.91) --
	(282.12,324.40) --
	(337.26,297.39);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (254.55,307.70) --
	(282.12,318.43) --
	(337.26,293.16);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (254.55,313.15) --
	(282.12,322.72) --
	(337.26,297.68);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (254.55,325.94) --
	(282.12,339.01) --
	(337.26,308.68);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt ,line join=round] (254.55,322.23) --
	(282.12,334.34) --
	(337.26,307.90);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt on 4pt off 3pt ,line join=round] (254.55,326.00) --
	(282.12,339.10) --
	(337.26,307.80);
\definecolor{drawColor}{RGB}{27,158,119}
\definecolor{fillColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (254.55,313.91) circle (  1.96);
\definecolor{drawColor}{RGB}{217,95,2}
\definecolor{fillColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (254.55,307.70) circle (  1.96);
\definecolor{drawColor}{RGB}{117,112,179}
\definecolor{fillColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (254.55,313.15) circle (  1.96);
\definecolor{drawColor}{RGB}{231,41,138}
\definecolor{fillColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (254.55,325.94) circle (  1.96);
\definecolor{drawColor}{RGB}{102,166,30}
\definecolor{fillColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (254.55,322.23) circle (  1.96);
\definecolor{drawColor}{RGB}{230,171,2}
\definecolor{fillColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (254.55,326.00) circle (  1.96);
\definecolor{drawColor}{RGB}{27,158,119}
\definecolor{fillColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (282.12,324.40) circle (  1.96);
\definecolor{drawColor}{RGB}{217,95,2}
\definecolor{fillColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (282.12,318.43) circle (  1.96);
\definecolor{drawColor}{RGB}{117,112,179}
\definecolor{fillColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (282.12,322.72) circle (  1.96);
\definecolor{drawColor}{RGB}{231,41,138}
\definecolor{fillColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (282.12,339.01) circle (  1.96);
\definecolor{drawColor}{RGB}{102,166,30}
\definecolor{fillColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (282.12,334.34) circle (  1.96);
\definecolor{drawColor}{RGB}{230,171,2}
\definecolor{fillColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (282.12,339.10) circle (  1.96);
\definecolor{drawColor}{RGB}{27,158,119}
\definecolor{fillColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (337.26,297.39) circle (  1.96);
\definecolor{drawColor}{RGB}{217,95,2}
\definecolor{fillColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (337.26,293.16) circle (  1.96);
\definecolor{drawColor}{RGB}{117,112,179}
\definecolor{fillColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (337.26,297.68) circle (  1.96);
\definecolor{drawColor}{RGB}{231,41,138}
\definecolor{fillColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (337.26,308.68) circle (  1.96);
\definecolor{drawColor}{RGB}{102,166,30}
\definecolor{fillColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (337.26,307.90) circle (  1.96);
\definecolor{drawColor}{RGB}{230,171,2}
\definecolor{fillColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (337.26,307.80) circle (  1.96);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (250.42,290.86) rectangle (341.40,341.40);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (245.47,291.50) {6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (245.47,304.53) {7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (245.47,317.57) {8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (245.47,330.61) {9};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (247.67,294.53) --
	(250.42,294.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (247.67,307.57) --
	(250.42,307.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] (247.67,320.60) --
	(250.42,320.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (247.67,333.64) --
	(250.42,333.64);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.55,288.11) --
	(254.55,290.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] (282.12,288.11) --
	(282.12,290.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] (309.69,288.11) --
	(309.69,290.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] (337.26,288.11) --
	(337.26,290.86);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (254.55,279.85) {2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (282.12,279.85) {4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (309.69,279.85) {6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (337.26,279.85) {8};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (295.91,267.81) {Nodes};
\end{scope}
\begin{scope}
\path[clip] ( 12.91,173.45) rectangle (124.24,260.17);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 12.91,173.45) rectangle (124.24,260.17);
\end{scope}
\begin{scope}
\path[clip] ( 27.76,204.13) rectangle (118.74,254.67);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 27.76,204.13) rectangle (118.74,254.67);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.76,211.92) --
	(118.74,211.92);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.76,224.78) --
	(118.74,224.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.76,237.63) --
	(118.74,237.63);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.76,250.49) --
	(118.74,250.49);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 52.57,204.13) --
	( 52.57,254.67);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 93.93,204.13) --
	( 93.93,254.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.76,205.49) --
	(118.74,205.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.76,218.35) --
	(118.74,218.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.76,231.21) --
	(118.74,231.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.76,244.06) --
	(118.74,244.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.89,204.13) --
	( 31.89,254.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 73.25,204.13) --
	( 73.25,254.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.60,204.13) --
	(114.60,254.67);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.89,206.50) --
	( 73.25,232.39) --
	(114.60,246.53);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 31.89,206.43) --
	( 73.25,230.60) --
	(114.60,244.33);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 31.89,206.59) --
	( 73.25,232.14) --
	(114.60,246.19);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 31.89,206.63) --
	( 73.25,235.27) --
	(114.60,252.37);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt ,line join=round] ( 31.89,206.61) --
	( 73.25,234.81) --
	(114.60,250.70);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt on 4pt off 3pt ,line join=round] ( 31.89,206.61) --
	( 73.25,235.71) --
	(114.60,251.98);
\definecolor{drawColor}{RGB}{27,158,119}
\definecolor{fillColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 31.89,206.50) circle (  1.96);
\definecolor{drawColor}{RGB}{217,95,2}
\definecolor{fillColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 31.89,206.43) circle (  1.96);
\definecolor{drawColor}{RGB}{117,112,179}
\definecolor{fillColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 31.89,206.59) circle (  1.96);
\definecolor{drawColor}{RGB}{231,41,138}
\definecolor{fillColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 31.89,206.63) circle (  1.96);
\definecolor{drawColor}{RGB}{102,166,30}
\definecolor{fillColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 31.89,206.61) circle (  1.96);
\definecolor{drawColor}{RGB}{230,171,2}
\definecolor{fillColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 31.89,206.61) circle (  1.96);
\definecolor{drawColor}{RGB}{27,158,119}
\definecolor{fillColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 73.25,232.39) circle (  1.96);
\definecolor{drawColor}{RGB}{217,95,2}
\definecolor{fillColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 73.25,230.60) circle (  1.96);
\definecolor{drawColor}{RGB}{117,112,179}
\definecolor{fillColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 73.25,232.14) circle (  1.96);
\definecolor{drawColor}{RGB}{231,41,138}
\definecolor{fillColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 73.25,235.27) circle (  1.96);
\definecolor{drawColor}{RGB}{102,166,30}
\definecolor{fillColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 73.25,234.81) circle (  1.96);
\definecolor{drawColor}{RGB}{230,171,2}
\definecolor{fillColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 73.25,235.71) circle (  1.96);
\definecolor{drawColor}{RGB}{27,158,119}
\definecolor{fillColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (114.60,246.53) circle (  1.96);
\definecolor{drawColor}{RGB}{217,95,2}
\definecolor{fillColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (114.60,244.33) circle (  1.96);
\definecolor{drawColor}{RGB}{117,112,179}
\definecolor{fillColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (114.60,246.19) circle (  1.96);
\definecolor{drawColor}{RGB}{231,41,138}
\definecolor{fillColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (114.60,252.37) circle (  1.96);
\definecolor{drawColor}{RGB}{102,166,30}
\definecolor{fillColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (114.60,250.70) circle (  1.96);
\definecolor{drawColor}{RGB}{230,171,2}
\definecolor{fillColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (114.60,251.98) circle (  1.96);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 27.76,204.13) rectangle (118.74,254.67);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.81,202.46) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.81,215.32) {3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.81,228.17) {6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.81,241.03) {9};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 25.01,205.49) --
	( 27.76,205.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 25.01,218.35) --
	( 27.76,218.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 25.01,231.21) --
	( 27.76,231.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 25.01,244.06) --
	( 27.76,244.06);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.89,201.38) --
	( 31.89,204.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 73.25,201.38) --
	( 73.25,204.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.60,201.38) --
	(114.60,204.13);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 31.89,193.12) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 73.25,193.12) {2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (114.60,193.12) {3};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 73.25,181.09) {Maximum arity};
\end{scope}
\begin{scope}
\path[clip] (124.24,173.45) rectangle (235.57,260.17);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (124.24,173.45) rectangle (235.57,260.17);
\end{scope}
\begin{scope}
\path[clip] (139.09,204.13) rectangle (230.07,254.67);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (139.09,204.13) rectangle (230.07,254.67);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (139.09,211.84) --
	(230.07,211.84);

\path[draw=drawColor,line width= 0.3pt,line join=round] (139.09,228.56) --
	(230.07,228.56);

\path[draw=drawColor,line width= 0.3pt,line join=round] (139.09,245.29) --
	(230.07,245.29);

\path[draw=drawColor,line width= 0.3pt,line join=round] (157.01,204.13) --
	(157.01,254.67);

\path[draw=drawColor,line width= 0.3pt,line join=round] (184.58,204.13) --
	(184.58,254.67);

\path[draw=drawColor,line width= 0.3pt,line join=round] (212.15,204.13) --
	(212.15,254.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.09,220.20) --
	(230.07,220.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.09,236.93) --
	(230.07,236.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.09,253.65) --
	(230.07,253.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (143.22,204.13) --
	(143.22,254.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (170.79,204.13) --
	(170.79,254.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (198.36,204.13) --
	(198.36,254.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (225.93,204.13) --
	(225.93,254.67);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (143.22,214.12) --
	(170.79,232.73) --
	(225.93,234.71);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (143.22,206.43) --
	(170.79,225.85) --
	(225.93,227.75);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (143.22,211.87) --
	(170.79,232.76) --
	(225.93,233.78);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (143.22,228.34) --
	(170.79,250.01) --
	(225.93,252.21);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt ,line join=round] (143.22,225.19) --
	(170.79,244.09) --
	(225.93,248.77);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt on 4pt off 3pt ,line join=round] (143.22,228.37) --
	(170.79,249.10) --
	(225.93,252.37);
\definecolor{drawColor}{RGB}{27,158,119}
\definecolor{fillColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (143.22,214.12) circle (  1.96);
\definecolor{drawColor}{RGB}{217,95,2}
\definecolor{fillColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (143.22,206.43) circle (  1.96);
\definecolor{drawColor}{RGB}{117,112,179}
\definecolor{fillColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (143.22,211.87) circle (  1.96);
\definecolor{drawColor}{RGB}{231,41,138}
\definecolor{fillColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (143.22,228.34) circle (  1.96);
\definecolor{drawColor}{RGB}{102,166,30}
\definecolor{fillColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (143.22,225.19) circle (  1.96);
\definecolor{drawColor}{RGB}{230,171,2}
\definecolor{fillColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (143.22,228.37) circle (  1.96);
\definecolor{drawColor}{RGB}{27,158,119}
\definecolor{fillColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (170.79,232.73) circle (  1.96);
\definecolor{drawColor}{RGB}{217,95,2}
\definecolor{fillColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (170.79,225.85) circle (  1.96);
\definecolor{drawColor}{RGB}{117,112,179}
\definecolor{fillColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (170.79,232.76) circle (  1.96);
\definecolor{drawColor}{RGB}{231,41,138}
\definecolor{fillColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (170.79,250.01) circle (  1.96);
\definecolor{drawColor}{RGB}{102,166,30}
\definecolor{fillColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (170.79,244.09) circle (  1.96);
\definecolor{drawColor}{RGB}{230,171,2}
\definecolor{fillColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (170.79,249.10) circle (  1.96);
\definecolor{drawColor}{RGB}{27,158,119}
\definecolor{fillColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (225.93,234.71) circle (  1.96);
\definecolor{drawColor}{RGB}{217,95,2}
\definecolor{fillColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (225.93,227.75) circle (  1.96);
\definecolor{drawColor}{RGB}{117,112,179}
\definecolor{fillColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (225.93,233.78) circle (  1.96);
\definecolor{drawColor}{RGB}{231,41,138}
\definecolor{fillColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (225.93,252.21) circle (  1.96);
\definecolor{drawColor}{RGB}{102,166,30}
\definecolor{fillColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (225.93,248.77) circle (  1.96);
\definecolor{drawColor}{RGB}{230,171,2}
\definecolor{fillColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (225.93,252.37) circle (  1.96);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (139.09,204.13) rectangle (230.07,254.67);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (134.14,217.17) {7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (134.14,233.90) {8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (134.14,250.62) {9};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (136.34,220.20) --
	(139.09,220.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (136.34,236.93) --
	(139.09,236.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (136.34,253.65) --
	(139.09,253.65);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (143.22,201.38) --
	(143.22,204.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (170.79,201.38) --
	(170.79,204.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (198.36,201.38) --
	(198.36,204.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (225.93,201.38) --
	(225.93,204.13);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (143.22,193.12) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (170.79,193.12) {200};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (198.36,193.12) {300};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (225.93,193.12) {400};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (184.58,181.09) {Domain size};
\end{scope}
\begin{scope}
\path[clip] (235.57,173.45) rectangle (346.90,260.17);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (235.57,173.45) rectangle (346.90,260.17);
\end{scope}
\begin{scope}
\path[clip] (250.42,204.13) rectangle (341.40,254.67);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (250.42,204.13) rectangle (341.40,254.67);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (250.42,207.43) --
	(341.40,207.43);

\path[draw=drawColor,line width= 0.3pt,line join=round] (250.42,223.92) --
	(341.40,223.92);

\path[draw=drawColor,line width= 0.3pt,line join=round] (250.42,240.40) --
	(341.40,240.40);

\path[draw=drawColor,line width= 0.3pt,line join=round] (254.55,204.13) --
	(254.55,254.67);

\path[draw=drawColor,line width= 0.3pt,line join=round] (271.09,204.13) --
	(271.09,254.67);

\path[draw=drawColor,line width= 0.3pt,line join=round] (287.64,204.13) --
	(287.64,254.67);

\path[draw=drawColor,line width= 0.3pt,line join=round] (304.18,204.13) --
	(304.18,254.67);

\path[draw=drawColor,line width= 0.3pt,line join=round] (320.72,204.13) --
	(320.72,254.67);

\path[draw=drawColor,line width= 0.3pt,line join=round] (337.26,204.13) --
	(337.26,254.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.42,215.67) --
	(341.40,215.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.42,232.16) --
	(341.40,232.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.42,248.64) --
	(341.40,248.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (262.82,204.13) --
	(262.82,254.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (279.36,204.13) --
	(279.36,254.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (295.91,204.13) --
	(295.91,254.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (312.45,204.13) --
	(312.45,254.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (328.99,204.13) --
	(328.99,254.67);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.55,212.10) --
	(295.91,221.05) --
	(337.26,235.32);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (254.55,206.43) --
	(295.91,214.71) --
	(337.26,226.12);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (254.55,210.56) --
	(295.91,221.30) --
	(337.26,233.56);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (254.55,227.18) --
	(295.91,238.03) --
	(337.26,251.65);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt ,line join=round] (254.55,223.69) --
	(295.91,232.98) --
	(337.26,247.85);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt on 4pt off 3pt ,line join=round] (254.55,226.85) --
	(295.91,236.95) --
	(337.26,252.37);
\definecolor{drawColor}{RGB}{27,158,119}
\definecolor{fillColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (254.55,212.10) circle (  1.96);
\definecolor{drawColor}{RGB}{217,95,2}
\definecolor{fillColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (254.55,206.43) circle (  1.96);
\definecolor{drawColor}{RGB}{117,112,179}
\definecolor{fillColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (254.55,210.56) circle (  1.96);
\definecolor{drawColor}{RGB}{231,41,138}
\definecolor{fillColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (254.55,227.18) circle (  1.96);
\definecolor{drawColor}{RGB}{102,166,30}
\definecolor{fillColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (254.55,223.69) circle (  1.96);
\definecolor{drawColor}{RGB}{230,171,2}
\definecolor{fillColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (254.55,226.85) circle (  1.96);
\definecolor{drawColor}{RGB}{27,158,119}
\definecolor{fillColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (295.91,221.05) circle (  1.96);
\definecolor{drawColor}{RGB}{217,95,2}
\definecolor{fillColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (295.91,214.71) circle (  1.96);
\definecolor{drawColor}{RGB}{117,112,179}
\definecolor{fillColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (295.91,221.30) circle (  1.96);
\definecolor{drawColor}{RGB}{231,41,138}
\definecolor{fillColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (295.91,238.03) circle (  1.96);
\definecolor{drawColor}{RGB}{102,166,30}
\definecolor{fillColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (295.91,232.98) circle (  1.96);
\definecolor{drawColor}{RGB}{230,171,2}
\definecolor{fillColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (295.91,236.95) circle (  1.96);
\definecolor{drawColor}{RGB}{27,158,119}
\definecolor{fillColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (337.26,235.32) circle (  1.96);
\definecolor{drawColor}{RGB}{217,95,2}
\definecolor{fillColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (337.26,226.12) circle (  1.96);
\definecolor{drawColor}{RGB}{117,112,179}
\definecolor{fillColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (337.26,233.56) circle (  1.96);
\definecolor{drawColor}{RGB}{231,41,138}
\definecolor{fillColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (337.26,251.65) circle (  1.96);
\definecolor{drawColor}{RGB}{102,166,30}
\definecolor{fillColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (337.26,247.85) circle (  1.96);
\definecolor{drawColor}{RGB}{230,171,2}
\definecolor{fillColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (337.26,252.37) circle (  1.96);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (250.42,204.13) rectangle (341.40,254.67);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (245.47,212.64) {7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (245.47,229.13) {8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (245.47,245.61) {9};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (247.67,215.67) --
	(250.42,215.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (247.67,232.16) --
	(250.42,232.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (247.67,248.64) --
	(250.42,248.64);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (262.82,201.38) --
	(262.82,204.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (279.36,201.38) --
	(279.36,204.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (295.91,201.38) --
	(295.91,204.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (312.45,201.38) --
	(312.45,204.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (328.99,201.38) --
	(328.99,204.13);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (262.82,193.12) {30};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (279.36,193.12) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (295.91,193.12) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (312.45,193.12) {60};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (328.99,193.12) {70};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (295.91,181.09) {Probabilistic (\%)};
\end{scope}
\begin{scope}
\path[clip] ( 12.91, 86.72) rectangle (124.24,173.45);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 12.91, 86.72) rectangle (124.24,173.45);
\end{scope}
\begin{scope}
\path[clip] ( 32.16,117.41) rectangle (118.74,167.95);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 32.16,117.41) rectangle (118.74,167.95);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 32.16,128.06) --
	(118.74,128.06);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 32.16,144.84) --
	(118.74,144.84);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 32.16,161.63) --
	(118.74,161.63);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 45.83,117.41) --
	( 45.83,167.95);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 65.54,117.41) --
	( 65.54,167.95);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 85.24,117.41) --
	( 85.24,167.95);

\path[draw=drawColor,line width= 0.3pt,line join=round] (104.95,117.41) --
	(104.95,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.16,119.67) --
	(118.74,119.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.16,136.45) --
	(118.74,136.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.16,153.24) --
	(118.74,153.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.97,117.41) --
	( 35.97,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 55.68,117.41) --
	( 55.68,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.39,117.41) --
	( 75.39,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 95.10,117.41) --
	( 95.10,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.80,117.41) --
	(114.80,167.95);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 36.09,119.72) --
	( 36.21,119.77) --
	( 36.45,119.88) --
	( 36.76,120.09) --
	( 36.92,120.09) --
	( 37.87,120.46) --
	( 41.95,122.00) --
	( 42.06,121.25) --
	( 42.30,122.28) --
	( 43.86,123.42) --
	( 47.80,124.68) --
	( 47.92,124.73) --
	( 48.15,124.86) --
	( 53.77,127.57) --
	( 54.01,127.77) --
	( 59.62,130.91) --
	( 59.74,129.19) --
	( 59.86,130.97) --
	( 59.98,128.46) --
	( 60.45,134.18) --
	( 65.71,132.99) --
	( 71.57,135.39) --
	( 77.42,136.47) --
	( 83.27,148.25) --
	( 83.51,142.00) --
	( 83.98,141.34) --
	(107.04,151.71) --
	(107.51,150.96) --
	(114.80,160.18);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 36.09,119.71) --
	( 36.21,119.75) --
	( 36.45,119.86) --
	( 36.76,120.06) --
	( 36.92,120.04) --
	( 37.87,120.43) --
	( 41.95,122.21) --
	( 42.06,121.72) --
	( 42.30,121.78) --
	( 43.86,123.23) --
	( 47.80,125.46) --
	( 47.92,124.49) --
	( 48.15,125.11) --
	( 53.77,127.15) --
	( 54.01,127.34) --
	( 59.62,129.94) --
	( 59.74,128.39) --
	( 59.86,128.78) --
	( 59.98,128.97) --
	( 60.45,130.60) --
	( 65.71,132.43) --
	( 71.57,134.46) --
	( 77.42,136.74) --
	( 83.27,148.62) --
	( 83.51,137.63) --
	( 83.98,139.91) --
	(107.04,148.68) --
	(107.51,151.41) --
	(114.80,157.85);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 36.09,119.72) --
	( 36.21,119.80) --
	( 36.45,119.91) --
	( 36.76,120.11) --
	( 36.92,120.11) --
	( 37.87,120.53) --
	( 41.95,121.97) --
	( 42.06,121.57) --
	( 42.30,122.16) --
	( 43.86,123.50) --
	( 47.80,125.79) --
	( 47.92,125.05) --
	( 48.15,124.49) --
	( 53.77,127.44) --
	( 54.01,127.85) --
	( 59.62,128.96) --
	( 59.74,128.40) --
	( 59.86,129.86) --
	( 59.98,128.71) --
	( 60.45,129.25) --
	( 65.71,132.97) --
	( 71.57,136.19) --
	( 77.42,135.39) --
	( 83.27,145.85) --
	( 83.51,139.99) --
	( 83.98,143.21) --
	(107.04,148.76) --
	(107.51,151.40) --
	(114.80,159.74);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 36.09,119.72) --
	( 36.21,119.77) --
	( 36.45,119.90) --
	( 36.76,120.12) --
	( 36.92,120.09) --
	( 37.87,120.59) --
	( 41.95,122.13) --
	( 42.06,121.81) --
	( 42.30,122.38) --
	( 43.86,123.94) --
	( 47.80,127.25) --
	( 47.92,125.25) --
	( 48.15,124.05) --
	( 53.77,127.79) --
	( 54.01,128.92) --
	( 59.62,131.74) --
	( 59.74,131.68) --
	( 59.86,131.74) --
	( 59.98,132.83) --
	( 60.45,132.18) --
	( 65.71,133.74) --
	( 71.57,136.49) --
	( 77.42,138.64) --
	( 83.27,149.54) --
	( 83.51,141.59) --
	( 83.98,137.80) --
	(107.04,155.32) --
	(107.51,155.78) --
	(114.80,165.65);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt ,line join=round] ( 36.09,119.73) --
	( 36.21,119.78) --
	( 36.45,119.90) --
	( 36.76,120.14) --
	( 36.92,120.11) --
	( 37.87,120.59) --
	( 41.95,122.31) --
	( 42.06,122.77) --
	( 42.30,122.88) --
	( 43.86,123.76) --
	( 47.80,125.71) --
	( 47.92,125.16) --
	( 48.15,124.65) --
	( 53.77,128.14) --
	( 54.01,128.34) --
	( 59.62,131.35) --
	( 59.74,129.35) --
	( 59.86,131.05) --
	( 59.98,129.44) --
	( 60.45,132.32) --
	( 65.71,134.56) --
	( 71.57,137.57) --
	( 77.42,139.20) --
	( 83.27,153.56) --
	( 83.51,144.11) --
	( 83.98,143.55) --
	(107.04,152.34) --
	(107.51,152.40) --
	(114.80,164.25);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt on 4pt off 3pt ,line join=round] ( 36.09,119.71) --
	( 36.21,119.79) --
	( 36.45,119.90) --
	( 36.76,120.13) --
	( 36.92,120.12) --
	( 37.87,120.58) --
	( 41.95,122.27) --
	( 42.06,121.03) --
	( 42.30,122.91) --
	( 43.86,123.95) --
	( 47.80,126.79) --
	( 47.92,126.12) --
	( 48.15,124.49) --
	( 53.77,128.53) --
	( 54.01,128.94) --
	( 59.62,131.40) --
	( 59.74,132.84) --
	( 59.86,132.17) --
	( 59.98,133.57) --
	( 60.45,131.59) --
	( 65.71,133.74) --
	( 71.57,137.52) --
	( 77.42,140.05) --
	( 83.27,151.84) --
	( 83.51,142.28) --
	( 83.98,142.84) --
	(107.04,154.31) --
	(107.51,157.04) --
	(114.80,165.44);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 32.16,117.41) rectangle (118.74,167.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 27.21,116.63) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 27.21,133.42) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 27.21,150.21) {20};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 29.41,119.67) --
	( 32.16,119.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 29.41,136.45) --
	( 32.16,136.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 29.41,153.24) --
	( 32.16,153.24);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.97,114.66) --
	( 35.97,117.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 55.68,114.66) --
	( 55.68,117.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.39,114.66) --
	( 75.39,117.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 95.10,114.66) --
	( 95.10,117.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.80,114.66) --
	(114.80,117.41);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.97,106.40) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 55.68,106.40) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 75.39,106.40) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 95.10,106.40) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (114.80,106.40) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 75.45, 94.36) {Facts ($\times 10^3$)};
\end{scope}
\begin{scope}
\path[clip] (124.24, 86.72) rectangle (235.57,173.45);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (124.24, 86.72) rectangle (235.57,173.45);
\end{scope}
\begin{scope}
\path[clip] (139.09,117.41) rectangle (230.07,167.95);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (139.09,117.41) rectangle (230.07,167.95);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (139.09,125.20) --
	(230.07,125.20);

\path[draw=drawColor,line width= 0.3pt,line join=round] (139.09,137.64) --
	(230.07,137.64);

\path[draw=drawColor,line width= 0.3pt,line join=round] (139.09,150.08) --
	(230.07,150.08);

\path[draw=drawColor,line width= 0.3pt,line join=round] (139.09,162.53) --
	(230.07,162.53);

\path[draw=drawColor,line width= 0.3pt,line join=round] (157.99,117.41) --
	(157.99,167.95);

\path[draw=drawColor,line width= 0.3pt,line join=round] (187.53,117.41) --
	(187.53,167.95);

\path[draw=drawColor,line width= 0.3pt,line join=round] (217.07,117.41) --
	(217.07,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.09,118.97) --
	(230.07,118.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.09,131.42) --
	(230.07,131.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.09,143.86) --
	(230.07,143.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.09,156.30) --
	(230.07,156.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (143.22,117.41) --
	(143.22,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (172.76,117.41) --
	(172.76,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (202.30,117.41) --
	(202.30,167.95);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (143.22,139.22) --
	(146.18,155.92) --
	(152.08,149.90) --
	(155.04,144.58) --
	(157.99,119.97) --
	(160.95,148.61) --
	(166.85,150.58) --
	(169.81,134.55) --
	(172.76,120.32) --
	(175.72,119.71) --
	(178.67,145.34) --
	(187.53,156.56) --
	(190.49,148.79) --
	(193.44,152.28) --
	(199.35,145.83) --
	(202.30,137.52) --
	(205.25,160.94) --
	(208.21,153.17) --
	(211.16,158.51) --
	(220.02,147.15) --
	(222.98,150.54) --
	(225.93,157.87);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (143.22,137.59) --
	(146.18,152.49) --
	(152.08,147.86) --
	(155.04,143.63) --
	(157.99,119.89) --
	(160.95,147.13) --
	(166.85,148.48) --
	(169.81,133.92) --
	(172.76,120.21) --
	(175.72,119.77) --
	(178.67,144.01) --
	(187.53,155.28) --
	(190.49,146.35) --
	(193.44,151.46) --
	(199.35,144.05) --
	(202.30,136.79) --
	(205.25,158.32) --
	(208.21,151.95) --
	(211.16,157.12) --
	(220.02,145.33) --
	(222.98,148.60) --
	(225.93,155.10);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (143.22,139.02) --
	(146.18,154.20) --
	(152.08,149.08) --
	(155.04,144.06) --
	(157.99,120.08) --
	(160.95,149.06) --
	(166.85,150.27) --
	(169.81,134.93) --
	(172.76,120.26) --
	(175.72,119.89) --
	(178.67,145.20) --
	(187.53,156.80) --
	(190.49,149.30) --
	(193.44,152.29) --
	(199.35,146.04) --
	(202.30,139.11) --
	(205.25,159.46) --
	(208.21,152.25) --
	(211.16,158.93) --
	(220.02,147.40) --
	(222.98,150.46) --
	(225.93,158.55);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (143.22,141.93) --
	(146.18,159.04) --
	(152.08,155.16) --
	(155.04,148.18) --
	(157.99,120.05) --
	(160.95,153.61) --
	(166.85,156.02) --
	(169.81,137.76) --
	(172.76,120.45) --
	(175.72,119.89) --
	(178.67,149.94) --
	(187.53,162.77) --
	(190.49,151.09) --
	(193.44,158.05) --
	(199.35,149.64) --
	(202.30,139.09) --
	(205.25,165.65) --
	(208.21,157.19) --
	(211.16,164.32) --
	(220.02,148.44) --
	(222.98,153.84) --
	(225.93,161.54);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt ,line join=round] (143.22,140.95) --
	(146.18,158.76) --
	(152.08,151.56) --
	(155.04,147.74) --
	(157.99,120.04) --
	(160.95,151.92) --
	(166.85,153.94) --
	(169.81,136.38) --
	(172.76,120.40) --
	(175.72,119.79) --
	(178.67,148.77) --
	(187.53,160.96) --
	(190.49,151.05) --
	(193.44,156.05) --
	(199.35,148.99) --
	(202.30,140.68) --
	(205.25,164.32) --
	(208.21,155.58) --
	(211.16,162.98) --
	(220.02,149.46) --
	(222.98,152.57) --
	(225.93,162.34);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt on 4pt off 3pt ,line join=round] (143.22,141.69) --
	(146.18,159.20) --
	(152.08,154.84) --
	(155.04,147.27) --
	(157.99,120.09) --
	(160.95,154.14) --
	(166.85,155.58) --
	(169.81,137.53) --
	(172.76,120.51) --
	(175.72,119.88) --
	(178.67,149.27) --
	(187.53,162.44) --
	(190.49,152.59) --
	(193.44,158.05) --
	(199.35,150.06) --
	(202.30,140.19) --
	(205.25,164.92) --
	(208.21,156.18) --
	(211.16,164.59) --
	(220.02,149.49) --
	(222.98,153.89) --
	(225.93,161.55);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (139.09,117.41) rectangle (230.07,167.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (134.14,115.94) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (134.14,128.39) {3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (134.14,140.83) {6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (134.14,153.27) {9};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (136.34,118.97) --
	(139.09,118.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (136.34,131.42) --
	(139.09,131.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (136.34,143.86) --
	(139.09,143.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] (136.34,156.30) --
	(139.09,156.30);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (143.22,114.66) --
	(143.22,117.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (172.76,114.66) --
	(172.76,117.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (202.30,114.66) --
	(202.30,117.41);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (143.22,106.40) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (172.76,106.40) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (202.30,106.40) {20};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (184.58, 94.36) {Independent pairs};
\end{scope}
\begin{scope}
\path[clip] (235.57, 86.72) rectangle (346.90,173.45);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (235.57, 86.72) rectangle (346.90,173.45);
\end{scope}
\begin{scope}
\path[clip] (250.42,117.41) rectangle (341.40,167.95);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (250.42,117.41) rectangle (341.40,167.95);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (250.42,125.38) --
	(341.40,125.38);

\path[draw=drawColor,line width= 0.3pt,line join=round] (250.42,137.76) --
	(341.40,137.76);

\path[draw=drawColor,line width= 0.3pt,line join=round] (250.42,150.15) --
	(341.40,150.15);

\path[draw=drawColor,line width= 0.3pt,line join=round] (250.42,162.54) --
	(341.40,162.54);

\path[draw=drawColor,line width= 0.3pt,line join=round] (264.89,117.41) --
	(264.89,167.95);

\path[draw=drawColor,line width= 0.3pt,line join=round] (285.57,117.41) --
	(285.57,167.95);

\path[draw=drawColor,line width= 0.3pt,line join=round] (306.24,117.41) --
	(306.24,167.95);

\path[draw=drawColor,line width= 0.3pt,line join=round] (326.92,117.41) --
	(326.92,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.42,119.18) --
	(341.40,119.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.42,131.57) --
	(341.40,131.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.42,143.96) --
	(341.40,143.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.42,156.35) --
	(341.40,156.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.55,117.41) --
	(254.55,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (275.23,117.41) --
	(275.23,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (295.91,117.41) --
	(295.91,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (316.58,117.41) --
	(316.58,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (337.26,117.41) --
	(337.26,167.95);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.55,139.33) --
	(257.51,153.19) --
	(263.41,149.97) --
	(266.37,147.54) --
	(268.34,158.45) --
	(269.32,120.56) --
	(272.27,120.30) --
	(278.18,150.65) --
	(281.14,134.69) --
	(284.09,120.52) --
	(287.04,119.91) --
	(290.00,145.43) --
	(298.86,156.60) --
	(301.81,148.87) --
	(304.77,152.34) --
	(309.69,142.15) --
	(310.68,145.92) --
	(313.63,137.65) --
	(316.58,160.96) --
	(319.54,153.23) --
	(322.49,158.54) --
	(323.48,119.82) --
	(331.35,147.23) --
	(334.31,150.61) --
	(337.26,155.49);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (254.55,137.72) --
	(257.51,149.95) --
	(263.41,147.94) --
	(266.37,147.05) --
	(268.34,155.61) --
	(269.32,120.51) --
	(272.27,120.06) --
	(278.18,148.56) --
	(281.14,134.06) --
	(284.09,120.41) --
	(287.04,119.97) --
	(290.00,144.10) --
	(298.86,155.33) --
	(301.81,146.44) --
	(304.77,151.52) --
	(309.69,140.79) --
	(310.68,144.15) --
	(313.63,136.91) --
	(316.58,158.35) --
	(319.54,152.01) --
	(322.49,157.16) --
	(323.48,119.71) --
	(331.35,145.43) --
	(334.31,148.68) --
	(337.26,152.58);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (254.55,139.14) --
	(257.51,151.37) --
	(263.41,149.16) --
	(266.37,147.66) --
	(268.34,157.54) --
	(269.32,120.78) --
	(272.27,120.36) --
	(278.18,150.34) --
	(281.14,135.07) --
	(284.09,120.47) --
	(287.04,120.09) --
	(290.00,145.29) --
	(298.86,156.84) --
	(301.81,149.37) --
	(304.77,152.35) --
	(309.69,141.07) --
	(310.68,146.13) --
	(313.63,139.23) --
	(316.58,159.49) --
	(319.54,152.31) --
	(322.49,158.96) --
	(323.48,119.82) --
	(331.35,147.49) --
	(334.31,150.53) --
	(337.26,154.98);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (254.55,142.04) --
	(257.51,155.46) --
	(263.41,155.20) --
	(266.37,152.47) --
	(268.34,163.20) --
	(269.32,120.74) --
	(272.27,120.28) --
	(278.18,156.06) --
	(281.14,137.88) --
	(284.09,120.65) --
	(287.04,120.09) --
	(290.00,150.01) --
	(298.86,162.78) --
	(301.81,151.15) --
	(304.77,158.08) --
	(309.69,144.54) --
	(310.68,149.72) --
	(313.63,139.21) --
	(316.58,165.65) --
	(319.54,157.22) --
	(322.49,164.33) --
	(323.48,119.80) --
	(331.35,148.52) --
	(334.31,153.89) --
	(337.26,159.35);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt ,line join=round] (254.55,141.06) --
	(257.51,157.32) --
	(263.41,151.62) --
	(266.37,150.39) --
	(268.34,161.36) --
	(269.32,120.67) --
	(272.27,120.32) --
	(278.18,153.99) --
	(281.14,136.51) --
	(284.09,120.60) --
	(287.04,119.99) --
	(290.00,148.85) --
	(298.86,160.98) --
	(301.81,151.12) --
	(304.77,156.09) --
	(309.69,145.54) --
	(310.68,149.06) --
	(313.63,140.79) --
	(316.58,164.32) --
	(319.54,155.62) --
	(322.49,162.99) --
	(323.48,119.83) --
	(331.35,149.53) --
	(334.31,152.63) --
	(337.26,158.52);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt on 4pt off 3pt ,line join=round] (254.55,141.80) --
	(257.51,155.92) --
	(263.41,154.89) --
	(266.37,151.09) --
	(268.34,162.99) --
	(269.32,120.78) --
	(272.27,120.43) --
	(278.18,155.62) --
	(281.14,137.65) --
	(284.09,120.71) --
	(287.04,120.09) --
	(290.00,149.34) --
	(298.86,162.45) --
	(301.81,152.65) --
	(304.77,158.08) --
	(309.69,144.05) --
	(310.68,150.13) --
	(313.63,140.30) --
	(316.58,164.93) --
	(319.54,156.22) --
	(322.49,164.60) --
	(323.48,119.84) --
	(331.35,149.57) --
	(334.31,153.94) --
	(337.26,159.64);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (250.42,117.41) rectangle (341.40,167.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (245.47,116.15) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (245.47,128.54) {3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (245.47,140.93) {6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (245.47,153.32) {9};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (247.67,119.18) --
	(250.42,119.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (247.67,131.57) --
	(250.42,131.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] (247.67,143.96) --
	(250.42,143.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (247.67,156.35) --
	(250.42,156.35);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.55,114.66) --
	(254.55,117.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (275.23,114.66) --
	(275.23,117.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (295.91,114.66) --
	(295.91,117.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (316.58,114.66) --
	(316.58,117.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (337.26,114.66) --
	(337.26,117.41);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (254.55,106.40) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (275.23,106.40) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (295.91,106.40) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (316.58,106.40) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (337.26,106.40) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (295.91, 94.36) {Independent (\%)};
\end{scope}
\begin{scope}
\path[clip] ( 12.91,  0.00) rectangle (124.24, 86.72);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 12.91,  0.00) rectangle (124.24, 86.72);
\end{scope}
\begin{scope}
\path[clip] ( 32.16, 30.69) rectangle (118.74, 81.22);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 32.16, 30.69) rectangle (118.74, 81.22);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 32.16, 40.36) --
	(118.74, 40.36);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 32.16, 56.68) --
	(118.74, 56.68);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 32.16, 73.00) --
	(118.74, 73.00);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.59, 30.69) --
	( 46.59, 81.22);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 67.58, 30.69) --
	( 67.58, 81.22);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 88.57, 30.69) --
	( 88.57, 81.22);

\path[draw=drawColor,line width= 0.3pt,line join=round] (109.56, 30.69) --
	(109.56, 81.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.16, 32.20) --
	(118.74, 32.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.16, 48.52) --
	(118.74, 48.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.16, 64.84) --
	(118.74, 64.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.16, 81.16) --
	(118.74, 81.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 36.09, 30.69) --
	( 36.09, 81.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 57.08, 30.69) --
	( 57.08, 81.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.07, 30.69) --
	( 78.07, 81.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 99.06, 30.69) --
	( 99.06, 81.22);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 36.09, 33.22) --
	( 37.09, 55.98) --
	( 38.09, 61.81) --
	( 39.08, 67.20) --
	( 40.08, 70.71) --
	( 41.07, 72.69) --
	( 42.07, 73.54) --
	( 43.07, 73.63) --
	( 44.06, 73.32) --
	( 45.06, 72.94) --
	( 46.06, 72.56) --
	( 47.05, 72.19) --
	( 48.05, 71.81) --
	( 49.05, 71.44) --
	( 50.04, 71.07) --
	( 51.04, 70.69) --
	( 52.03, 70.32) --
	( 53.03, 69.95) --
	( 54.03, 69.58) --
	( 55.02, 69.21) --
	( 56.02, 68.84) --
	( 57.02, 68.48) --
	( 58.01, 68.11) --
	( 59.01, 67.74) --
	( 60.00, 67.38) --
	( 61.00, 67.01) --
	( 62.00, 66.65) --
	( 62.99, 66.28) --
	( 63.99, 65.92) --
	( 64.99, 65.56) --
	( 65.98, 65.20) --
	( 66.98, 64.84) --
	( 67.98, 64.48) --
	( 68.97, 64.12) --
	( 69.97, 63.76) --
	( 70.96, 63.40) --
	( 71.96, 63.04) --
	( 72.96, 62.69) --
	( 73.95, 62.33) --
	( 74.95, 61.97) --
	( 75.95, 61.62) --
	( 76.94, 61.26) --
	( 77.94, 60.91) --
	( 78.94, 60.56) --
	( 79.93, 60.20) --
	( 80.93, 59.85) --
	( 81.92, 59.50) --
	( 82.92, 59.14) --
	( 83.92, 58.79) --
	( 84.91, 58.44) --
	( 85.91, 58.09) --
	( 86.91, 57.74) --
	( 87.90, 57.39) --
	( 88.90, 57.04) --
	( 89.89, 56.69) --
	( 90.89, 56.34) --
	( 91.89, 55.99) --
	( 92.88, 55.64) --
	( 93.88, 55.29) --
	( 94.88, 54.95) --
	( 95.87, 54.60) --
	( 96.87, 54.25) --
	( 97.87, 53.90) --
	( 98.86, 53.56) --
	( 99.86, 53.21) --
	(100.85, 52.86) --
	(101.85, 52.52) --
	(102.85, 52.17) --
	(103.84, 51.82) --
	(104.84, 51.48) --
	(105.84, 51.13) --
	(106.83, 50.79) --
	(107.83, 50.44) --
	(108.83, 50.09) --
	(109.82, 49.75) --
	(110.82, 49.40) --
	(111.81, 49.06) --
	(112.81, 48.71) --
	(113.81, 48.37) --
	(114.80, 48.02);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] ( 36.09, 32.98) --
	( 37.09, 54.54) --
	( 38.09, 59.85) --
	( 39.08, 64.85) --
	( 40.08, 68.09) --
	( 41.07, 69.93) --
	( 42.07, 70.71) --
	( 43.07, 70.77) --
	( 44.06, 70.48) --
	( 45.06, 70.12) --
	( 46.06, 69.76) --
	( 47.05, 69.40) --
	( 48.05, 69.04) --
	( 49.05, 68.68) --
	( 50.04, 68.33) --
	( 51.04, 67.97) --
	( 52.03, 67.62) --
	( 53.03, 67.27) --
	( 54.03, 66.92) --
	( 55.02, 66.57) --
	( 56.02, 66.22) --
	( 57.02, 65.87) --
	( 58.01, 65.53) --
	( 59.01, 65.18) --
	( 60.00, 64.84) --
	( 61.00, 64.50) --
	( 62.00, 64.15) --
	( 62.99, 63.81) --
	( 63.99, 63.47) --
	( 64.99, 63.13) --
	( 65.98, 62.80) --
	( 66.98, 62.46) --
	( 67.98, 62.12) --
	( 68.97, 61.79) --
	( 69.97, 61.45) --
	( 70.96, 61.12) --
	( 71.96, 60.79) --
	( 72.96, 60.45) --
	( 73.95, 60.12) --
	( 74.95, 59.79) --
	( 75.95, 59.46) --
	( 76.94, 59.13) --
	( 77.94, 58.80) --
	( 78.94, 58.47) --
	( 79.93, 58.15) --
	( 80.93, 57.82) --
	( 81.92, 57.49) --
	( 82.92, 57.17) --
	( 83.92, 56.84) --
	( 84.91, 56.52) --
	( 85.91, 56.20) --
	( 86.91, 55.87) --
	( 87.90, 55.55) --
	( 88.90, 55.23) --
	( 89.89, 54.91) --
	( 90.89, 54.58) --
	( 91.89, 54.26) --
	( 92.88, 53.94) --
	( 93.88, 53.62) --
	( 94.88, 53.30) --
	( 95.87, 52.98) --
	( 96.87, 52.67) --
	( 97.87, 52.35) --
	( 98.86, 52.03) --
	( 99.86, 51.71) --
	(100.85, 51.39) --
	(101.85, 51.08) --
	(102.85, 50.76) --
	(103.84, 50.44) --
	(104.84, 50.12) --
	(105.84, 49.81) --
	(106.83, 49.49) --
	(107.83, 49.18) --
	(108.83, 48.86) --
	(109.82, 48.54) --
	(110.82, 48.23) --
	(111.81, 47.91) --
	(112.81, 47.60) --
	(113.81, 47.28) --
	(114.80, 46.96);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] ( 36.09, 33.06) --
	( 37.09, 55.40) --
	( 38.09, 61.11) --
	( 39.08, 66.53) --
	( 40.08, 70.05) --
	( 41.07, 72.05) --
	( 42.07, 72.90) --
	( 43.07, 72.99) --
	( 44.06, 72.68) --
	( 45.06, 72.31) --
	( 46.06, 71.93) --
	( 47.05, 71.55) --
	( 48.05, 71.18) --
	( 49.05, 70.81) --
	( 50.04, 70.44) --
	( 51.04, 70.07) --
	( 52.03, 69.70) --
	( 53.03, 69.33) --
	( 54.03, 68.96) --
	( 55.02, 68.59) --
	( 56.02, 68.23) --
	( 57.02, 67.86) --
	( 58.01, 67.50) --
	( 59.01, 67.14) --
	( 60.00, 66.77) --
	( 61.00, 66.41) --
	( 62.00, 66.05) --
	( 62.99, 65.69) --
	( 63.99, 65.33) --
	( 64.99, 64.98) --
	( 65.98, 64.62) --
	( 66.98, 64.26) --
	( 67.98, 63.91) --
	( 68.97, 63.55) --
	( 69.97, 63.20) --
	( 70.96, 62.85) --
	( 71.96, 62.49) --
	( 72.96, 62.14) --
	( 73.95, 61.79) --
	( 74.95, 61.44) --
	( 75.95, 61.09) --
	( 76.94, 60.74) --
	( 77.94, 60.39) --
	( 78.94, 60.04) --
	( 79.93, 59.70) --
	( 80.93, 59.35) --
	( 81.92, 59.00) --
	( 82.92, 58.66) --
	( 83.92, 58.31) --
	( 84.91, 57.97) --
	( 85.91, 57.62) --
	( 86.91, 57.28) --
	( 87.90, 56.94) --
	( 88.90, 56.59) --
	( 89.89, 56.25) --
	( 90.89, 55.91) --
	( 91.89, 55.57) --
	( 92.88, 55.23) --
	( 93.88, 54.89) --
	( 94.88, 54.54) --
	( 95.87, 54.20) --
	( 96.87, 53.86) --
	( 97.87, 53.52) --
	( 98.86, 53.19) --
	( 99.86, 52.85) --
	(100.85, 52.51) --
	(101.85, 52.17) --
	(102.85, 51.83) --
	(103.84, 51.49) --
	(104.84, 51.15) --
	(105.84, 50.82) --
	(106.83, 50.48) --
	(107.83, 50.14) --
	(108.83, 49.80) --
	(109.82, 49.46) --
	(110.82, 49.13) --
	(111.81, 48.79) --
	(112.81, 48.45) --
	(113.81, 48.12) --
	(114.80, 47.78);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 36.09, 33.25) --
	( 37.09, 59.80) --
	( 38.09, 66.04) --
	( 39.08, 71.90) --
	( 40.08, 75.67) --
	( 41.07, 77.78) --
	( 42.07, 78.63) --
	( 43.07, 78.64) --
	( 44.06, 78.22) --
	( 45.06, 77.72) --
	( 46.06, 77.23) --
	( 47.05, 76.75) --
	( 48.05, 76.27) --
	( 49.05, 75.79) --
	( 50.04, 75.31) --
	( 51.04, 74.84) --
	( 52.03, 74.38) --
	( 53.03, 73.91) --
	( 54.03, 73.45) --
	( 55.02, 73.00) --
	( 56.02, 72.54) --
	( 57.02, 72.09) --
	( 58.01, 71.65) --
	( 59.01, 71.20) --
	( 60.00, 70.76) --
	( 61.00, 70.32) --
	( 62.00, 69.89) --
	( 62.99, 69.46) --
	( 63.99, 69.03) --
	( 64.99, 68.61) --
	( 65.98, 68.18) --
	( 66.98, 67.76) --
	( 67.98, 67.35) --
	( 68.97, 66.93) --
	( 69.97, 66.52) --
	( 70.96, 66.11) --
	( 71.96, 65.71) --
	( 72.96, 65.30) --
	( 73.95, 64.90) --
	( 74.95, 64.50) --
	( 75.95, 64.10) --
	( 76.94, 63.71) --
	( 77.94, 63.32) --
	( 78.94, 62.93) --
	( 79.93, 62.54) --
	( 80.93, 62.15) --
	( 81.92, 61.77) --
	( 82.92, 61.38) --
	( 83.92, 61.00) --
	( 84.91, 60.62) --
	( 85.91, 60.24) --
	( 86.91, 59.87) --
	( 87.90, 59.49) --
	( 88.90, 59.12) --
	( 89.89, 58.75) --
	( 90.89, 58.38) --
	( 91.89, 58.01) --
	( 92.88, 57.64) --
	( 93.88, 57.28) --
	( 94.88, 56.91) --
	( 95.87, 56.55) --
	( 96.87, 56.19) --
	( 97.87, 55.82) --
	( 98.86, 55.46) --
	( 99.86, 55.10) --
	(100.85, 54.75) --
	(101.85, 54.39) --
	(102.85, 54.03) --
	(103.84, 53.67) --
	(104.84, 53.32) --
	(105.84, 52.96) --
	(106.83, 52.61) --
	(107.83, 52.25) --
	(108.83, 51.90) --
	(109.82, 51.54) --
	(110.82, 51.19) --
	(111.81, 50.84) --
	(112.81, 50.48) --
	(113.81, 50.13) --
	(114.80, 49.78);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] ( 36.09, 33.22) --
	( 37.09, 58.30) --
	( 38.09, 64.82) --
	( 39.08, 70.85) --
	( 40.08, 74.75) --
	( 41.07, 76.94) --
	( 42.07, 77.84) --
	( 43.07, 77.89) --
	( 44.06, 77.49) --
	( 45.06, 77.02) --
	( 46.06, 76.55) --
	( 47.05, 76.08) --
	( 48.05, 75.62) --
	( 49.05, 75.16) --
	( 50.04, 74.70) --
	( 51.04, 74.25) --
	( 52.03, 73.80) --
	( 53.03, 73.35) --
	( 54.03, 72.91) --
	( 55.02, 72.47) --
	( 56.02, 72.03) --
	( 57.02, 71.59) --
	( 58.01, 71.16) --
	( 59.01, 70.72) --
	( 60.00, 70.30) --
	( 61.00, 69.87) --
	( 62.00, 69.45) --
	( 62.99, 69.03) --
	( 63.99, 68.61) --
	( 64.99, 68.19) --
	( 65.98, 67.78) --
	( 66.98, 67.37) --
	( 67.98, 66.96) --
	( 68.97, 66.55) --
	( 69.97, 66.15) --
	( 70.96, 65.74) --
	( 71.96, 65.34) --
	( 72.96, 64.94) --
	( 73.95, 64.55) --
	( 74.95, 64.15) --
	( 75.95, 63.76) --
	( 76.94, 63.37) --
	( 77.94, 62.98) --
	( 78.94, 62.59) --
	( 79.93, 62.21) --
	( 80.93, 61.82) --
	( 81.92, 61.44) --
	( 82.92, 61.06) --
	( 83.92, 60.68) --
	( 84.91, 60.30) --
	( 85.91, 59.92) --
	( 86.91, 59.55) --
	( 87.90, 59.17) --
	( 88.90, 58.80) --
	( 89.89, 58.43) --
	( 90.89, 58.06) --
	( 91.89, 57.69) --
	( 92.88, 57.32) --
	( 93.88, 56.95) --
	( 94.88, 56.59) --
	( 95.87, 56.22) --
	( 96.87, 55.86) --
	( 97.87, 55.49) --
	( 98.86, 55.13) --
	( 99.86, 54.77) --
	(100.85, 54.41) --
	(101.85, 54.05) --
	(102.85, 53.69) --
	(103.84, 53.33) --
	(104.84, 52.97) --
	(105.84, 52.61) --
	(106.83, 52.25) --
	(107.83, 51.89) --
	(108.83, 51.54) --
	(109.82, 51.18) --
	(110.82, 50.82) --
	(111.81, 50.46) --
	(112.81, 50.11) --
	(113.81, 49.75) --
	(114.80, 49.39);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt on 4pt off 3pt ,line join=round] ( 36.09, 33.25) --
	( 37.09, 59.84) --
	( 38.09, 66.25) --
	( 39.08, 72.14) --
	( 40.08, 75.94) --
	( 41.07, 78.06) --
	( 42.07, 78.92) --
	( 43.07, 78.93) --
	( 44.06, 78.51) --
	( 45.06, 78.01) --
	( 46.06, 77.52) --
	( 47.05, 77.03) --
	( 48.05, 76.55) --
	( 49.05, 76.07) --
	( 50.04, 75.59) --
	( 51.04, 75.12) --
	( 52.03, 74.65) --
	( 53.03, 74.18) --
	( 54.03, 73.72) --
	( 55.02, 73.26) --
	( 56.02, 72.81) --
	( 57.02, 72.36) --
	( 58.01, 71.91) --
	( 59.01, 71.46) --
	( 60.00, 71.02) --
	( 61.00, 70.58) --
	( 62.00, 70.15) --
	( 62.99, 69.72) --
	( 63.99, 69.29) --
	( 64.99, 68.86) --
	( 65.98, 68.44) --
	( 66.98, 68.02) --
	( 67.98, 67.60) --
	( 68.97, 67.18) --
	( 69.97, 66.77) --
	( 70.96, 66.36) --
	( 71.96, 65.96) --
	( 72.96, 65.55) --
	( 73.95, 65.15) --
	( 74.95, 64.75) --
	( 75.95, 64.35) --
	( 76.94, 63.96) --
	( 77.94, 63.56) --
	( 78.94, 63.17) --
	( 79.93, 62.78) --
	( 80.93, 62.40) --
	( 81.92, 62.01) --
	( 82.92, 61.63) --
	( 83.92, 61.25) --
	( 84.91, 60.87) --
	( 85.91, 60.49) --
	( 86.91, 60.11) --
	( 87.90, 59.74) --
	( 88.90, 59.37) --
	( 89.89, 58.99) --
	( 90.89, 58.63) --
	( 91.89, 58.26) --
	( 92.88, 57.89) --
	( 93.88, 57.52) --
	( 94.88, 57.16) --
	( 95.87, 56.80) --
	( 96.87, 56.43) --
	( 97.87, 56.07) --
	( 98.86, 55.71) --
	( 99.86, 55.35) --
	(100.85, 54.99) --
	(101.85, 54.64) --
	(102.85, 54.28) --
	(103.84, 53.92) --
	(104.84, 53.57) --
	(105.84, 53.21) --
	(106.83, 52.86) --
	(107.83, 52.50) --
	(108.83, 52.15) --
	(109.82, 51.79) --
	(110.82, 51.44) --
	(111.81, 51.09) --
	(112.81, 50.74) --
	(113.81, 50.38) --
	(114.80, 50.03);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 32.16, 30.69) rectangle (118.74, 81.22);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 27.21, 29.17) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 27.21, 45.49) {5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 27.21, 61.81) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 27.21, 78.13) {15};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 29.41, 32.20) --
	( 32.16, 32.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 29.41, 48.52) --
	( 32.16, 48.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 29.41, 64.84) --
	( 32.16, 64.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 29.41, 81.16) --
	( 32.16, 81.16);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 36.09, 27.94) --
	( 36.09, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 57.08, 27.94) --
	( 57.08, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.07, 27.94) --
	( 78.07, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 99.06, 27.94) --
	( 99.06, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.09, 19.68) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 57.08, 19.68) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 78.07, 19.68) {0.4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 99.06, 19.68) {0.6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 75.45,  7.64) {Proportion listed};
\end{scope}
\begin{scope}
\path[clip] (124.24,  0.00) rectangle (235.57, 86.72);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (124.24,  0.00) rectangle (235.57, 86.72);
\end{scope}
\begin{scope}
\path[clip] (139.09, 32.18) rectangle (230.07, 81.22);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (139.09, 32.18) rectangle (230.07, 81.22);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (139.09, 40.82) --
	(230.07, 40.82);

\path[draw=drawColor,line width= 0.3pt,line join=round] (139.09, 52.61) --
	(230.07, 52.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (139.09, 64.41) --
	(230.07, 64.41);

\path[draw=drawColor,line width= 0.3pt,line join=round] (139.09, 76.20) --
	(230.07, 76.20);

\path[draw=drawColor,line width= 0.3pt,line join=round] (139.34, 32.18) --
	(139.34, 81.22);

\path[draw=drawColor,line width= 0.3pt,line join=round] (145.79, 32.18) --
	(145.79, 81.22);

\path[draw=drawColor,line width= 0.3pt,line join=round] (158.70, 32.18) --
	(158.70, 81.22);

\path[draw=drawColor,line width= 0.3pt,line join=round] (171.60, 32.18) --
	(171.60, 81.22);

\path[draw=drawColor,line width= 0.3pt,line join=round] (184.51, 32.18) --
	(184.51, 81.22);

\path[draw=drawColor,line width= 0.3pt,line join=round] (197.42, 32.18) --
	(197.42, 81.22);

\path[draw=drawColor,line width= 0.3pt,line join=round] (210.32, 32.18) --
	(210.32, 81.22);

\path[draw=drawColor,line width= 0.3pt,line join=round] (223.23, 32.18) --
	(223.23, 81.22);

\path[draw=drawColor,line width= 0.3pt,line join=round] (229.68, 32.18) --
	(229.68, 81.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.09, 34.92) --
	(230.07, 34.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.09, 46.71) --
	(230.07, 46.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.09, 58.51) --
	(230.07, 58.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.09, 70.30) --
	(230.07, 70.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (152.24, 32.18) --
	(152.24, 81.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.15, 32.18) --
	(165.15, 81.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (178.06, 32.18) --
	(178.06, 81.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (190.96, 32.18) --
	(190.96, 81.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (203.87, 32.18) --
	(203.87, 81.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (216.78, 32.18) --
	(216.78, 81.22);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 1.1pt,line join=round] (143.22, 34.40) --
	(144.27, 34.41) --
	(145.32, 34.42) --
	(146.36, 34.43) --
	(147.41, 34.46) --
	(148.46, 34.50) --
	(149.50, 34.56) --
	(150.55, 34.64) --
	(151.60, 34.75) --
	(152.65, 34.89) --
	(153.69, 35.06) --
	(154.74, 35.28) --
	(155.79, 35.53) --
	(156.83, 35.83) --
	(157.88, 36.18) --
	(158.93, 36.58) --
	(159.97, 37.04) --
	(161.02, 37.55) --
	(162.07, 38.14) --
	(163.11, 38.79) --
	(164.16, 39.51) --
	(165.21, 40.31) --
	(166.26, 41.19) --
	(167.30, 42.15) --
	(168.35, 43.20) --
	(169.40, 44.34) --
	(170.44, 45.58) --
	(171.49, 46.91) --
	(172.54, 48.35) --
	(173.58, 49.89) --
	(174.63, 51.54) --
	(175.68, 53.30) --
	(176.73, 55.15) --
	(177.77, 57.04) --
	(178.82, 58.95) --
	(179.87, 60.85) --
	(180.91, 62.68) --
	(181.96, 64.43) --
	(183.01, 66.06) --
	(184.05, 67.53) --
	(185.10, 68.81) --
	(186.15, 69.86) --
	(187.19, 70.65) --
	(188.24, 71.18) --
	(189.29, 71.50) --
	(190.34, 71.67) --
	(191.38, 71.75) --
	(192.43, 71.78) --
	(193.48, 71.84) --
	(194.52, 71.96) --
	(195.57, 72.17) --
	(196.62, 72.40) --
	(197.66, 72.61) --
	(198.71, 72.76) --
	(199.76, 72.80) --
	(200.81, 72.75) --
	(201.85, 72.64) --
	(202.90, 72.51) --
	(203.95, 72.38) --
	(204.99, 72.26) --
	(206.04, 72.14) --
	(207.09, 72.03) --
	(208.13, 71.91) --
	(209.18, 71.80) --
	(210.23, 71.69) --
	(211.27, 71.59) --
	(212.32, 71.53) --
	(213.37, 71.50) --
	(214.42, 71.51) --
	(215.46, 71.59) --
	(216.51, 71.71) --
	(217.56, 71.87) --
	(218.60, 72.06) --
	(219.65, 72.26) --
	(220.70, 72.46) --
	(221.74, 72.66) --
	(222.79, 72.85) --
	(223.84, 73.05) --
	(224.88, 73.24) --
	(225.93, 73.43);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] (143.22, 34.82) --
	(144.27, 34.78) --
	(145.32, 34.75) --
	(146.36, 34.72) --
	(147.41, 34.71) --
	(148.46, 34.71) --
	(149.50, 34.73) --
	(150.55, 34.77) --
	(151.60, 34.84) --
	(152.65, 34.93) --
	(153.69, 35.06) --
	(154.74, 35.23) --
	(155.79, 35.44) --
	(156.83, 35.69) --
	(157.88, 35.99) --
	(158.93, 36.34) --
	(159.97, 36.74) --
	(161.02, 37.21) --
	(162.07, 37.73) --
	(163.11, 38.32) --
	(164.16, 38.99) --
	(165.21, 39.72) --
	(166.26, 40.54) --
	(167.30, 41.43) --
	(168.35, 42.41) --
	(169.40, 43.48) --
	(170.44, 44.64) --
	(171.49, 45.89) --
	(172.54, 47.25) --
	(173.58, 48.70) --
	(174.63, 50.27) --
	(175.68, 51.94) --
	(176.73, 53.69) --
	(177.77, 55.49) --
	(178.82, 57.30) --
	(179.87, 59.10) --
	(180.91, 60.84) --
	(181.96, 62.50) --
	(183.01, 64.03) --
	(184.05, 65.42) --
	(185.10, 66.62) --
	(186.15, 67.59) --
	(187.19, 68.32) --
	(188.24, 68.79) --
	(189.29, 69.06) --
	(190.34, 69.20) --
	(191.38, 69.24) --
	(192.43, 69.27) --
	(193.48, 69.33) --
	(194.52, 69.47) --
	(195.57, 69.71) --
	(196.62, 69.99) --
	(197.66, 70.26) --
	(198.71, 70.46) --
	(199.76, 70.55) --
	(200.81, 70.56) --
	(201.85, 70.52) --
	(202.90, 70.46) --
	(203.95, 70.43) --
	(204.99, 70.40) --
	(206.04, 70.38) --
	(207.09, 70.35) --
	(208.13, 70.30) --
	(209.18, 70.22) --
	(210.23, 70.13) --
	(211.27, 70.02) --
	(212.32, 69.93) --
	(213.37, 69.87) --
	(214.42, 69.85) --
	(215.46, 69.90) --
	(216.51, 69.99) --
	(217.56, 70.12) --
	(218.60, 70.29) --
	(219.65, 70.46) --
	(220.70, 70.64) --
	(221.74, 70.82) --
	(222.79, 70.99) --
	(223.84, 71.16) --
	(224.88, 71.32) --
	(225.93, 71.49);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] (143.22, 34.84) --
	(144.27, 34.81) --
	(145.32, 34.78) --
	(146.36, 34.76) --
	(147.41, 34.76) --
	(148.46, 34.76) --
	(149.50, 34.79) --
	(150.55, 34.84) --
	(151.60, 34.92) --
	(152.65, 35.03) --
	(153.69, 35.18) --
	(154.74, 35.36) --
	(155.79, 35.58) --
	(156.83, 35.86) --
	(157.88, 36.18) --
	(158.93, 36.56) --
	(159.97, 36.99) --
	(161.02, 37.49) --
	(162.07, 38.05) --
	(163.11, 38.68) --
	(164.16, 39.39) --
	(165.21, 40.18) --
	(166.26, 41.04) --
	(167.30, 41.99) --
	(168.35, 43.03) --
	(169.40, 44.17) --
	(170.44, 45.40) --
	(171.49, 46.73) --
	(172.54, 48.16) --
	(173.58, 49.71) --
	(174.63, 51.37) --
	(175.68, 53.13) --
	(176.73, 54.99) --
	(177.77, 56.89) --
	(178.82, 58.81) --
	(179.87, 60.70) --
	(180.91, 62.54) --
	(181.96, 64.28) --
	(183.01, 65.88) --
	(184.05, 67.32) --
	(185.10, 68.56) --
	(186.15, 69.55) --
	(187.19, 70.27) --
	(188.24, 70.72) --
	(189.29, 70.95) --
	(190.34, 71.02) --
	(191.38, 71.00) --
	(192.43, 70.95) --
	(193.48, 70.94) --
	(194.52, 71.01) --
	(195.57, 71.19) --
	(196.62, 71.42) --
	(197.66, 71.66) --
	(198.71, 71.84) --
	(199.76, 71.93) --
	(200.81, 71.95) --
	(201.85, 71.95) --
	(202.90, 71.96) --
	(203.95, 72.01) --
	(204.99, 72.08) --
	(206.04, 72.14) --
	(207.09, 72.17) --
	(208.13, 72.16) --
	(209.18, 72.07) --
	(210.23, 71.91) --
	(211.27, 71.71) --
	(212.32, 71.52) --
	(213.37, 71.35) --
	(214.42, 71.25) --
	(215.46, 71.25) --
	(216.51, 71.34) --
	(217.56, 71.51) --
	(218.60, 71.75) --
	(219.65, 72.05) --
	(220.70, 72.40) --
	(221.74, 72.79) --
	(222.79, 73.21) --
	(223.84, 73.65) --
	(224.88, 74.11) --
	(225.93, 74.58);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] (143.22, 34.59) --
	(144.27, 34.57) --
	(145.32, 34.56) --
	(146.36, 34.55) --
	(147.41, 34.55) --
	(148.46, 34.57) --
	(149.50, 34.62) --
	(150.55, 34.69) --
	(151.60, 34.78) --
	(152.65, 34.92) --
	(153.69, 35.09) --
	(154.74, 35.30) --
	(155.79, 35.56) --
	(156.83, 35.88) --
	(157.88, 36.24) --
	(158.93, 36.67) --
	(159.97, 37.16) --
	(161.02, 37.72) --
	(162.07, 38.35) --
	(163.11, 39.06) --
	(164.16, 39.85) --
	(165.21, 40.73) --
	(166.26, 41.69) --
	(167.30, 42.75) --
	(168.35, 43.91) --
	(169.40, 45.17) --
	(170.44, 46.53) --
	(171.49, 48.01) --
	(172.54, 49.60) --
	(173.58, 51.31) --
	(174.63, 53.15) --
	(175.68, 55.11) --
	(176.73, 57.16) --
	(177.77, 59.27) --
	(178.82, 61.40) --
	(179.87, 63.51) --
	(180.91, 65.56) --
	(181.96, 67.52) --
	(183.01, 69.34) --
	(184.05, 70.99) --
	(185.10, 72.43) --
	(186.15, 73.62) --
	(187.19, 74.53) --
	(188.24, 75.14) --
	(189.29, 75.54) --
	(190.34, 75.78) --
	(191.38, 75.93) --
	(192.43, 76.07) --
	(193.48, 76.25) --
	(194.52, 76.55) --
	(195.57, 76.96) --
	(196.62, 77.42) --
	(197.66, 77.84) --
	(198.71, 78.17) --
	(199.76, 78.34) --
	(200.81, 78.38) --
	(201.85, 78.35) --
	(202.90, 78.29) --
	(203.95, 78.24) --
	(204.99, 78.19) --
	(206.04, 78.13) --
	(207.09, 78.05) --
	(208.13, 77.94) --
	(209.18, 77.79) --
	(210.23, 77.61) --
	(211.27, 77.40) --
	(212.32, 77.21) --
	(213.37, 77.05) --
	(214.42, 76.94) --
	(215.46, 76.91) --
	(216.51, 76.95) --
	(217.56, 77.05) --
	(218.60, 77.19) --
	(219.65, 77.37) --
	(220.70, 77.57) --
	(221.74, 77.78) --
	(222.79, 78.00) --
	(223.84, 78.23) --
	(224.88, 78.47) --
	(225.93, 78.71);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] (143.22, 34.93) --
	(144.27, 34.88) --
	(145.32, 34.83) --
	(146.36, 34.79) --
	(147.41, 34.76) --
	(148.46, 34.75) --
	(149.50, 34.76) --
	(150.55, 34.80) --
	(151.60, 34.87) --
	(152.65, 34.97) --
	(153.69, 35.11) --
	(154.74, 35.29) --
	(155.79, 35.53) --
	(156.83, 35.81) --
	(157.88, 36.15) --
	(158.93, 36.55) --
	(159.97, 37.02) --
	(161.02, 37.55) --
	(162.07, 38.16) --
	(163.11, 38.85) --
	(164.16, 39.62) --
	(165.21, 40.47) --
	(166.26, 41.42) --
	(167.30, 42.46) --
	(168.35, 43.60) --
	(169.40, 44.84) --
	(170.44, 46.20) --
	(171.49, 47.66) --
	(172.54, 49.24) --
	(173.58, 50.95) --
	(174.63, 52.77) --
	(175.68, 54.73) --
	(176.73, 56.78) --
	(177.77, 58.89) --
	(178.82, 61.01) --
	(179.87, 63.12) --
	(180.91, 65.17) --
	(181.96, 67.13) --
	(183.01, 68.94) --
	(184.05, 70.59) --
	(185.10, 72.02) --
	(186.15, 73.20) --
	(187.19, 74.09) --
	(188.24, 74.69) --
	(189.29, 75.06) --
	(190.34, 75.26) --
	(191.38, 75.36) --
	(192.43, 75.42) --
	(193.48, 75.50) --
	(194.52, 75.67) --
	(195.57, 75.93) --
	(196.62, 76.22) --
	(197.66, 76.47) --
	(198.71, 76.62) --
	(199.76, 76.63) --
	(200.81, 76.53) --
	(201.85, 76.38) --
	(202.90, 76.23) --
	(203.95, 76.11) --
	(204.99, 76.03) --
	(206.04, 75.97) --
	(207.09, 75.91) --
	(208.13, 75.83) --
	(209.18, 75.73) --
	(210.23, 75.60) --
	(211.27, 75.46) --
	(212.32, 75.34) --
	(213.37, 75.27) --
	(214.42, 75.25) --
	(215.46, 75.31) --
	(216.51, 75.44) --
	(217.56, 75.62) --
	(218.60, 75.83) --
	(219.65, 76.05) --
	(220.70, 76.28) --
	(221.74, 76.49) --
	(222.79, 76.69) --
	(223.84, 76.88) --
	(224.88, 77.07) --
	(225.93, 77.26);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt on 4pt off 3pt ,line join=round] (143.22, 34.70) --
	(144.27, 34.67) --
	(145.32, 34.64) --
	(146.36, 34.63) --
	(147.41, 34.62) --
	(148.46, 34.63) --
	(149.50, 34.67) --
	(150.55, 34.73) --
	(151.60, 34.82) --
	(152.65, 34.95) --
	(153.69, 35.11) --
	(154.74, 35.32) --
	(155.79, 35.58) --
	(156.83, 35.89) --
	(157.88, 36.26) --
	(158.93, 36.68) --
	(159.97, 37.18) --
	(161.02, 37.74) --
	(162.07, 38.38) --
	(163.11, 39.09) --
	(164.16, 39.89) --
	(165.21, 40.78) --
	(166.26, 41.75) --
	(167.30, 42.83) --
	(168.35, 44.00) --
	(169.40, 45.28) --
	(170.44, 46.66) --
	(171.49, 48.17) --
	(172.54, 49.78) --
	(173.58, 51.52) --
	(174.63, 53.39) --
	(175.68, 55.38) --
	(176.73, 57.47) --
	(177.77, 59.62) --
	(178.82, 61.78) --
	(179.87, 63.92) --
	(180.91, 66.00) --
	(181.96, 67.97) --
	(183.01, 69.80) --
	(184.05, 71.45) --
	(185.10, 72.88) --
	(186.15, 74.04) --
	(187.19, 74.90) --
	(188.24, 75.46) --
	(189.29, 75.79) --
	(190.34, 75.97) --
	(191.38, 76.05) --
	(192.43, 76.14) --
	(193.48, 76.28) --
	(194.52, 76.56) --
	(195.57, 76.97) --
	(196.62, 77.44) --
	(197.66, 77.88) --
	(198.71, 78.21) --
	(199.76, 78.35) --
	(200.81, 78.35) --
	(201.85, 78.26) --
	(202.90, 78.12) --
	(203.95, 77.97) --
	(204.99, 77.81) --
	(206.04, 77.65) --
	(207.09, 77.48) --
	(208.13, 77.29) --
	(209.18, 77.08) --
	(210.23, 76.86) --
	(211.27, 76.64) --
	(212.32, 76.46) --
	(213.37, 76.32) --
	(214.42, 76.25) --
	(215.46, 76.25) --
	(216.51, 76.34) --
	(217.56, 76.49) --
	(218.60, 76.69) --
	(219.65, 76.95) --
	(220.70, 77.23) --
	(221.74, 77.55) --
	(222.79, 77.89) --
	(223.84, 78.25) --
	(224.88, 78.62) --
	(225.93, 78.99);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (139.09, 32.18) rectangle (230.07, 81.22);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (134.14, 31.89) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (134.14, 43.68) {3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (134.14, 55.48) {6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (134.14, 67.27) {9};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (136.34, 34.92) --
	(139.09, 34.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] (136.34, 46.71) --
	(139.09, 46.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] (136.34, 58.51) --
	(139.09, 58.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (136.34, 70.30) --
	(139.09, 70.30);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (152.24, 29.43) --
	(152.24, 32.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.15, 29.43) --
	(165.15, 32.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (178.06, 29.43) --
	(178.06, 32.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (190.96, 29.43) --
	(190.96, 32.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (203.87, 29.43) --
	(203.87, 32.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (216.78, 29.43) --
	(216.78, 32.18);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (146.31, 19.68) {10};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.62] at (155.10, 23.27) {3};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (159.21, 19.68) {10};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.62] at (168.01, 23.27) {4};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (172.12, 19.68) {10};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.62] at (180.92, 23.27) {5};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (185.03, 19.68) {10};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.62] at (193.82, 23.27) {6};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (197.93, 19.68) {10};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.62] at (206.73, 23.27) {7};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (210.84, 19.68) {10};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.62] at (219.64, 23.27) {8};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (184.58,  7.64) {Possible facts};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (233.95,  8.57) rectangle (348.52, 78.15);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (239.45, 64.01) {Algorithm};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (239.45, 42.98) rectangle (253.90, 57.44);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (240.89, 50.21) -- (252.45, 50.21);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{27,158,119}
\definecolor{fillColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (246.67, 50.21) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (239.45, 28.53) rectangle (253.90, 42.98);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (240.89, 35.75) -- (252.45, 35.75);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{217,95,2}
\definecolor{fillColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (246.67, 35.75) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (239.45, 14.07) rectangle (253.90, 28.53);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (240.89, 21.30) -- (252.45, 21.30);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{117,112,179}
\definecolor{fillColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (246.67, 21.30) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (298.38, 42.98) rectangle (312.83, 57.44);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (299.83, 50.21) -- (311.39, 50.21);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{231,41,138}
\definecolor{fillColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (305.61, 50.21) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (298.38, 28.53) rectangle (312.83, 42.98);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt ,line join=round] (299.83, 35.75) -- (311.39, 35.75);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{102,166,30}
\definecolor{fillColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (305.61, 35.75) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (298.38, 14.07) rectangle (312.83, 28.53);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt on 4pt off 3pt ,line join=round] (299.83, 21.30) -- (311.39, 21.30);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{230,171,2}
\definecolor{fillColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (305.61, 21.30) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (259.40, 47.18) {BDD};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (259.40, 32.72) {d-DNNF};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (259.40, 18.27) {K-Best};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.33, 47.18) {NNF};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.33, 32.72) {SDD};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.33, 18.27) {SDDX};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (346.90,346.90);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 13.34,173.45) {Mean inference time (s)};
\end{scope}
\end{tikzpicture}

\end{document}