### 2020-07-22 talk/impact.tex
Adapt plots to slides

% Created by tikzDevice version 0.12.3 on 2020-07-22 15:37:51
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (303.53,130.09);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (303.53,130.09);
\end{scope}
\begin{scope}
\path[clip] ( 40.51, 30.69) rectangle (127.90,124.59);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 40.51, 30.69) rectangle (127.90,124.59);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 40.51, 50.03) --
	(127.90, 50.03);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 40.51, 81.30) --
	(127.90, 81.30);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 40.51,112.57) --
	(127.90,112.57);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 56.44, 30.69) --
	( 56.44,124.59);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 80.36, 30.69) --
	( 80.36,124.59);

\path[draw=drawColor,line width= 0.3pt,line join=round] (104.27, 30.69) --
	(104.27,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.51, 34.40) --
	(127.90, 34.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.51, 65.67) --
	(127.90, 65.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.51, 96.94) --
	(127.90, 96.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.48, 30.69) --
	( 44.48,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.40, 30.69) --
	( 68.40,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 92.31, 30.69) --
	( 92.31,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (116.23, 30.69) --
	(116.23,124.59);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.48, 34.95) --
	( 68.40, 35.86) --
	( 92.31, 39.64) --
	(116.23, 66.05);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.48, 54.81) --
	( 68.40, 57.14) --
	( 82.39, 60.74) --
	( 92.31, 63.15);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.48, 55.77) --
	( 68.40, 58.31) --
	( 92.31, 59.68) --
	(116.23, 62.08);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.48, 57.27) --
	( 68.40, 58.33) --
	( 92.31, 59.95) --
	(116.23, 60.30);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.48, 54.34) --
	( 68.40, 55.03) --
	( 92.31, 58.92) --
	(116.23, 67.56);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.48, 37.33) --
	( 68.40, 37.69) --
	( 92.31, 40.51) --
	(116.23,120.32);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 30.69) rectangle (127.90,124.59);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.56, 31.37) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.56, 62.64) {2000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.56, 93.91) {4000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.76, 34.40) --
	( 40.51, 34.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.76, 65.67) --
	( 40.51, 65.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.76, 96.94) --
	( 40.51, 96.94);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.48, 27.94) --
	( 44.48, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.40, 27.94) --
	( 68.40, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 92.31, 27.94) --
	( 92.31, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (116.23, 27.94) --
	(116.23, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 44.48, 19.68) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 68.40, 19.68) {2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 92.31, 19.68) {4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (116.23, 19.68) {8};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 84.20,  7.64) {Value};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08, 77.64) {Mean number of nodes};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (138.90, 21.17) rectangle (298.03,134.11);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (144.40,119.96) {Variable};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (144.40, 98.94) rectangle (158.85,113.39);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (145.84,106.16) -- (157.41,106.16);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (144.40, 84.48) rectangle (158.85, 98.94);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,line join=round] (145.84, 91.71) -- (157.41, 91.71);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (144.40, 70.03) rectangle (158.85, 84.48);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,line join=round] (145.84, 77.26) -- (157.41, 77.26);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (144.40, 55.57) rectangle (158.85, 70.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,line join=round] (145.84, 62.80) -- (157.41, 62.80);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (144.40, 41.12) rectangle (158.85, 55.57);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 0.6pt,line join=round] (145.84, 48.35) -- (157.41, 48.35);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (144.40, 26.67) rectangle (158.85, 41.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 0.6pt,line join=round] (145.84, 33.89) -- (157.41, 33.89);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (164.35,103.13) {The number of predicates};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (164.35, 88.68) {Maximum arity};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (164.35, 74.23) {The number of variables};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (164.35, 59.77) {The number of constants};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (164.35, 45.32) {The number of additional clauses};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (164.35, 30.86) {The maximum number of nodes};
\end{scope}
\end{tikzpicture}