### 2020-08-13 seminar_talk/phase_transition.tex
More slides

% Created by tikzDevice version 0.12.3 on 2020-08-12 19:57:41
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (303.53,130.09);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (303.53,130.09);
\end{scope}
\begin{scope}
\path[clip] ( 34.79, 30.69) rectangle (298.03,124.59);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 34.79, 30.69) rectangle (298.03,124.59);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.79, 77.01) --
	(298.03, 77.01);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.79,100.92) --
	(298.03,100.92);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.79,107.58) --
	(298.03,107.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.79, 34.95) --
	(298.03, 34.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.79, 91.75) --
	(298.03, 91.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.20, 30.69) --
	( 40.20,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 49.22, 30.69) --
	( 49.22,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 58.23, 30.69) --
	( 58.23,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 67.25, 30.69) --
	( 67.25,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 76.26, 30.69) --
	( 76.26,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 85.28, 30.69) --
	( 85.28,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 94.29, 30.69) --
	( 94.29,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (103.31, 30.69) --
	(103.31,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (112.32, 30.69) --
	(112.32,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (121.34, 30.69) --
	(121.34,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.35, 30.69) --
	(130.35,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.37, 30.69) --
	(139.37,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (148.38, 30.69) --
	(148.38,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (157.40, 30.69) --
	(157.40,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (166.41, 30.69) --
	(166.41,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (175.43, 30.69) --
	(175.43,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (184.44, 30.69) --
	(184.44,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (193.46, 30.69) --
	(193.46,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (202.47, 30.69) --
	(202.47,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (211.49, 30.69) --
	(211.49,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (220.50, 30.69) --
	(220.50,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (229.52, 30.69) --
	(229.52,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (238.53, 30.69) --
	(238.53,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (247.55, 30.69) --
	(247.55,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (256.56, 30.69) --
	(256.56,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (265.58, 30.69) --
	(265.58,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (274.59, 30.69) --
	(274.59,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (283.61, 30.69) --
	(283.61,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (292.62, 30.69) --
	(292.62,124.59);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.20, 75.95) --
	( 40.20, 92.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.20, 52.30) --
	( 40.20, 43.78) --
	( 40.20, 30.65) --
	( 40.20,  0.76);
\definecolor{fillColor}{RGB}{224,236,244}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 36.82, 75.95) --
	( 36.82, 52.30) --
	( 40.20, 52.30) --
	( 43.58, 52.30) --
	( 43.58, 75.95) --
	( 40.20, 75.95) --
	( 36.82, 75.95) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 36.82, 64.33) --
	( 40.20, 64.33) --
	( 43.58, 64.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 49.22, 78.10) --
	( 49.22, 94.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 49.22, 52.30) --
	( 49.22, 39.14) --
	( 49.22,  9.06);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 45.84, 78.10) --
	( 45.84, 52.30) --
	( 49.22, 52.30) --
	( 52.60, 52.30) --
	( 52.60, 78.10) --
	( 49.22, 78.10) --
	( 45.84, 78.10) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 45.84, 64.78) --
	( 49.22, 64.78) --
	( 52.60, 64.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 58.23, 78.40) --
	( 58.23, 95.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 58.23, 51.68) --
	( 58.23, 37.75) --
	( 58.23,  2.66);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 54.85, 78.40) --
	( 54.85, 51.68) --
	( 58.23, 51.68) --
	( 61.61, 51.68) --
	( 61.61, 78.40) --
	( 58.23, 78.40) --
	( 54.85, 78.40) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 54.85, 65.00) --
	( 58.23, 65.00) --
	( 61.61, 65.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 67.25, 79.27) --
	( 67.25, 96.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 67.25, 51.80) --
	( 67.25, 38.19) --
	( 67.25,  5.26);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 63.87, 79.27) --
	( 63.87, 51.80) --
	( 67.25, 51.80) --
	( 70.63, 51.80) --
	( 70.63, 79.27) --
	( 67.25, 79.27) --
	( 63.87, 79.27) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 63.87, 65.00) --
	( 67.25, 65.00) --
	( 70.63, 65.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 76.26, 80.00) --
	( 76.26, 97.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 76.26, 51.80) --
	( 76.26, 38.19) --
	( 76.26,  5.26);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 72.88, 80.00) --
	( 72.88, 51.80) --
	( 76.26, 51.80) --
	( 79.64, 51.80) --
	( 79.64, 80.00) --
	( 76.26, 80.00) --
	( 72.88, 80.00) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 72.88, 65.50) --
	( 76.26, 65.50) --
	( 79.64, 65.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 85.28, 80.28) --
	( 85.28, 97.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 85.28, 51.93) --
	( 85.28, 38.08) --
	( 85.28,  3.55);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 81.90, 80.28) --
	( 81.90, 51.93) --
	( 85.28, 51.93) --
	( 88.66, 51.93) --
	( 88.66, 80.28) --
	( 85.28, 80.28) --
	( 81.90, 80.28) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 81.90, 65.29) --
	( 85.28, 65.29) --
	( 88.66, 65.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 94.29, 80.04) --
	( 94.29, 97.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 94.29, 51.30) --
	( 94.29, 37.75) --
	( 94.29,  5.26);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 90.91, 80.04) --
	( 90.91, 51.30) --
	( 94.29, 51.30) --
	( 97.67, 51.30) --
	( 97.67, 80.04) --
	( 94.29, 80.04) --
	( 90.91, 80.04) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 90.91, 65.00) --
	( 94.29, 65.00) --
	( 97.67, 65.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] (103.31, 81.09) --
	(103.31, 98.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (103.31, 51.30) --
	(103.31, 37.97) --
	(103.31,  6.85);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 99.93, 81.09) --
	( 99.93, 51.30) --
	(103.31, 51.30) --
	(106.69, 51.30) --
	(106.69, 81.09) --
	(103.31, 81.09) --
	( 99.93, 81.09) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 99.93, 65.36) --
	(103.31, 65.36) --
	(106.69, 65.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (112.32, 80.74) --
	(112.32, 98.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (112.32, 51.68) --
	(112.32, 38.62) --
	(112.32,  9.06);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (108.94, 80.74) --
	(108.94, 51.68) --
	(112.32, 51.68) --
	(115.70, 51.68) --
	(115.70, 80.74) --
	(112.32, 80.74) --
	(108.94, 80.74) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (108.94, 65.43) --
	(112.32, 65.43) --
	(115.70, 65.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (121.34, 81.32) --
	(121.34, 99.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (121.34, 51.30) --
	(121.34, 37.64) --
	(121.34,  4.42);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (117.96, 81.32) --
	(117.96, 51.30) --
	(121.34, 51.30) --
	(124.72, 51.30) --
	(124.72, 81.32) --
	(121.34, 81.32) --
	(117.96, 81.32) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (117.96, 65.29) --
	(121.34, 65.29) --
	(124.72, 65.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.35, 81.32) --
	(130.35, 98.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.35, 51.43) --
	(130.35, 42.90) --
	(130.35, 29.76) --
	(130.35,  0.00);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (126.97, 81.32) --
	(126.97, 51.43) --
	(130.35, 51.43) --
	(133.73, 51.43) --
	(133.73, 81.32) --
	(130.35, 81.32) --
	(126.97, 81.32) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (126.97, 65.64) --
	(130.35, 65.64) --
	(133.73, 65.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.37, 81.39) --
	(139.37, 99.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.37, 51.30) --
	(139.37, 37.97) --
	(139.37,  6.85);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (135.99, 81.39) --
	(135.99, 51.30) --
	(139.37, 51.30) --
	(142.75, 51.30) --
	(142.75, 81.39) --
	(139.37, 81.39) --
	(135.99, 81.39) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (135.99, 65.64) --
	(139.37, 65.64) --
	(142.75, 65.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (148.38, 81.77) --
	(148.38, 99.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (148.38, 51.17) --
	(148.38, 37.53) --
	(148.38,  4.42);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (145.00, 81.77) --
	(145.00, 51.17) --
	(148.38, 51.17) --
	(151.76, 51.17) --
	(151.76, 81.77) --
	(148.38, 81.77) --
	(145.00, 81.77) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (145.00, 65.43) --
	(148.38, 65.43) --
	(151.76, 65.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (157.40, 81.73) --
	(157.40, 99.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] (157.40, 51.05) --
	(157.40, 37.64) --
	(157.40,  6.06);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (154.02, 81.73) --
	(154.02, 51.05) --
	(157.40, 51.05) --
	(160.78, 51.05) --
	(160.78, 81.73) --
	(157.40, 81.73) --
	(154.02, 81.73) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (154.02, 65.14) --
	(157.40, 65.14) --
	(160.78, 65.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (166.41, 81.95) --
	(166.41, 99.79);

\path[draw=drawColor,line width= 0.6pt,line join=round] (166.41, 51.17) --
	(166.41, 37.97) --
	(166.41,  7.61);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (163.03, 81.95) --
	(163.03, 51.17) --
	(166.41, 51.17) --
	(169.79, 51.17) --
	(169.79, 81.95) --
	(166.41, 81.95) --
	(163.03, 81.95) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (163.03, 65.43) --
	(166.41, 65.43) --
	(169.79, 65.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (175.43, 81.87) --
	(175.43, 99.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] (175.43, 51.43) --
	(175.43, 38.19) --
	(175.43,  7.61);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (172.05, 81.87) --
	(172.05, 51.43) --
	(175.43, 51.43) --
	(178.81, 51.43) --
	(178.81, 81.87) --
	(175.43, 81.87) --
	(172.05, 81.87) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (172.05, 65.64) --
	(175.43, 65.64) --
	(178.81, 65.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (184.44, 82.43) --
	(184.44,100.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (184.44, 51.05) --
	(184.44, 37.53) --
	(184.44,  5.26);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (181.06, 82.43) --
	(181.06, 51.05) --
	(184.44, 51.05) --
	(187.82, 51.05) --
	(187.82, 82.43) --
	(184.44, 82.43) --
	(181.06, 82.43) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (181.06, 65.68) --
	(184.44, 65.68) --
	(187.82, 65.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] (193.46, 82.36) --
	(193.46,100.38);

\path[draw=drawColor,line width= 0.6pt,line join=round] (193.46, 50.92) --
	(193.46, 37.97) --
	(193.46,  9.06);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (190.08, 82.36) --
	(190.08, 50.92) --
	(193.46, 50.92) --
	(196.84, 50.92) --
	(196.84, 82.36) --
	(193.46, 82.36) --
	(190.08, 82.36) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (190.08, 65.43) --
	(193.46, 65.43) --
	(196.84, 65.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (202.47, 82.57) --
	(202.47,100.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (202.47, 50.66) --
	(202.47, 37.31) --
	(202.47,  6.06);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (199.09, 82.57) --
	(199.09, 50.66) --
	(202.47, 50.66) --
	(205.85, 50.66) --
	(205.85, 82.57) --
	(202.47, 82.57) --
	(199.09, 82.57) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (199.09, 65.36) --
	(202.47, 65.36) --
	(205.85, 65.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (211.49, 83.13) --
	(211.49,101.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (211.49, 50.92) --
	(211.49, 37.53) --
	(211.49,  6.06);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (208.11, 83.13) --
	(208.11, 50.92) --
	(211.49, 50.92) --
	(214.87, 50.92) --
	(214.87, 83.13) --
	(211.49, 83.13) --
	(208.11, 83.13) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (208.11, 65.50) --
	(211.49, 65.50) --
	(214.87, 65.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (220.50, 83.03) --
	(220.50,101.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (220.50, 50.66) --
	(220.50, 36.97) --
	(220.50,  3.55);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (217.12, 83.03) --
	(217.12, 50.66) --
	(220.50, 50.66) --
	(223.88, 50.66) --
	(223.88, 83.03) --
	(220.50, 83.03) --
	(217.12, 83.03) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (217.12, 65.36) --
	(220.50, 65.36) --
	(223.88, 65.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (229.52, 83.55) --
	(229.52,101.89);

\path[draw=drawColor,line width= 0.6pt,line join=round] (229.52, 50.53) --
	(229.52, 36.97) --
	(229.52,  4.42);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (226.14, 83.55) --
	(226.14, 50.53) --
	(229.52, 50.53) --
	(232.90, 50.53) --
	(232.90, 83.55) --
	(229.52, 83.55) --
	(226.14, 83.55) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (226.14, 65.36) --
	(229.52, 65.36) --
	(232.90, 65.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (238.53, 82.85) --
	(238.53,100.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] (238.53, 50.79) --
	(238.53, 37.53) --
	(238.53,  6.85);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (235.15, 82.85) --
	(235.15, 50.79) --
	(238.53, 50.79) --
	(241.91, 50.79) --
	(241.91, 82.85) --
	(238.53, 82.85) --
	(235.15, 82.85) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (235.15, 65.21) --
	(238.53, 65.21) --
	(241.91, 65.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (247.55, 83.58) --
	(247.55,101.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (247.55, 50.39) --
	(247.55, 37.08) --
	(247.55,  6.06);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (244.17, 83.58) --
	(244.17, 50.39) --
	(247.55, 50.39) --
	(250.93, 50.39) --
	(250.93, 83.58) --
	(247.55, 83.58) --
	(244.17, 83.58) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (244.17, 64.78) --
	(247.55, 64.78) --
	(250.93, 64.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (256.56, 83.79) --
	(256.56,102.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (256.56, 50.26) --
	(256.56, 36.97) --
	(256.56,  6.06);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (253.18, 83.79) --
	(253.18, 72.33) --
	(253.18, 50.26) --
	(256.56, 50.26) --
	(259.94, 50.26) --
	(259.94, 72.33) --
	(259.94, 83.79) --
	(256.56, 83.79) --
	(253.18, 83.79) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (253.18, 65.07) --
	(256.56, 65.07) --
	(259.94, 65.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (265.58, 84.22) --
	(265.58,102.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (265.58, 50.00) --
	(265.58, 37.31) --
	(265.58,  9.75);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (262.20, 84.22) --
	(262.20, 72.62) --
	(262.20, 50.00) --
	(265.58, 50.00) --
	(268.96, 50.00) --
	(268.96, 72.62) --
	(268.96, 84.22) --
	(265.58, 84.22) --
	(262.20, 84.22) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (262.20, 64.78) --
	(265.58, 64.78) --
	(268.96, 64.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (274.59, 84.32) --
	(274.59,102.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (274.59, 49.73) --
	(274.59, 36.27) --
	(274.59,  4.42);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (271.21, 84.32) --
	(271.21, 72.65) --
	(271.21, 49.73) --
	(274.59, 49.73) --
	(277.98, 49.73) --
	(277.98, 72.65) --
	(277.98, 84.32) --
	(274.59, 84.32) --
	(271.21, 84.32) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (271.21, 64.63) --
	(274.59, 64.63) --
	(277.98, 64.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] (283.61, 84.89) --
	(283.61,103.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] (283.61, 49.45) --
	(283.61, 36.39) --
	(283.61,  6.85);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (280.23, 84.89) --
	(280.23, 73.06) --
	(280.23, 49.45) --
	(283.61, 49.45) --
	(286.99, 49.45) --
	(286.99, 73.06) --
	(286.99, 84.89) --
	(283.61, 84.89) --
	(280.23, 84.89) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (280.23, 63.57) --
	(283.61, 63.57) --
	(286.99, 63.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] (292.62, 86.34) --
	(292.62,105.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (292.62, 49.04) --
	(292.62, 35.80) --
	(292.62,  5.26);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (289.24, 86.34) --
	(289.24, 74.16) --
	(289.24, 49.04) --
	(292.62, 49.04) --
	(296.01, 49.04) --
	(296.01, 74.16) --
	(296.01, 86.34) --
	(292.62, 86.34) --
	(289.24, 86.34) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (289.24, 62.38) --
	(292.62, 62.38) --
	(296.01, 62.38);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 40.20, 71.66) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 49.22, 79.45) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 58.23, 87.80) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 67.25, 91.37) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 76.26, 92.89) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 85.28, 98.48) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 94.29,100.35) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (103.31,104.38) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (112.32,105.75) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (121.34,110.28) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (130.35,110.49) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (139.37,109.72) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (148.38,114.89) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (157.40,113.61) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (166.41,115.59) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (175.43,116.86) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (184.44,115.92) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (193.46,117.85) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (202.47,116.11) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (211.49,117.57) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (220.50,117.72) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (229.52,118.06) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (238.53,117.37) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (247.55,116.43) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (256.56,113.89) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (265.58,112.76) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (274.59,107.36) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (283.61,102.11) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (292.62, 98.37) circle (  1.96);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 34.79, 30.69) rectangle (298.03,124.59);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96, 31.18) {10};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.62] at ( 26.76, 34.78) {2};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96, 87.98) {10};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.62] at ( 26.76, 91.58) {3};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.04, 34.95) --
	( 34.79, 34.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.04, 91.75) --
	( 34.79, 91.75);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.20, 27.94) --
	( 40.20, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 49.22, 27.94) --
	( 49.22, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 58.23, 27.94) --
	( 58.23, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 67.25, 27.94) --
	( 67.25, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 76.26, 27.94) --
	( 76.26, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 85.28, 27.94) --
	( 85.28, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 94.29, 27.94) --
	( 94.29, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (103.31, 27.94) --
	(103.31, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (112.32, 27.94) --
	(112.32, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (121.34, 27.94) --
	(121.34, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.35, 27.94) --
	(130.35, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.37, 27.94) --
	(139.37, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (148.38, 27.94) --
	(148.38, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (157.40, 27.94) --
	(157.40, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (166.41, 27.94) --
	(166.41, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (175.43, 27.94) --
	(175.43, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (184.44, 27.94) --
	(184.44, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (193.46, 27.94) --
	(193.46, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (202.47, 27.94) --
	(202.47, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (211.49, 27.94) --
	(211.49, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (220.50, 27.94) --
	(220.50, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (229.52, 27.94) --
	(229.52, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (238.53, 27.94) --
	(238.53, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (247.55, 27.94) --
	(247.55, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (256.56, 27.94) --
	(256.56, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (265.58, 27.94) --
	(265.58, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (274.59, 27.94) --
	(274.59, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (283.61, 27.94) --
	(283.61, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (292.62, 27.94) --
	(292.62, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 40.20, 19.68) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 49.22, 19.68) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 58.23, 19.68) {2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 67.25, 19.68) {3};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 76.26, 19.68) {4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 85.28, 19.68) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 94.29, 19.68) {6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (103.31, 19.68) {7};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (112.32, 19.68) {8};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (121.34, 19.68) {9};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (130.35, 19.68) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (139.37, 19.68) {11};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (148.38, 19.68) {12};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (157.40, 19.68) {13};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (166.41, 19.68) {14};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (175.43, 19.68) {15};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (184.44, 19.68) {16};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (193.46, 19.68) {17};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (202.47, 19.68) {18};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (211.49, 19.68) {19};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (220.50, 19.68) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (229.52, 19.68) {21};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (238.53, 19.68) {22};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (247.55, 19.68) {23};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (256.56, 19.68) {24};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (265.58, 19.68) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (274.59, 19.68) {26};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (283.61, 19.68) {27};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (292.62, 19.68) {28};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (166.41,  7.64) {Independent pairs of predicates};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08, 77.64) {Search nodes};
\end{scope}
\end{tikzpicture}

\end{document}