### 2020-10-19 talks/fata_seminar/proportion.tex

% Created by tikzDevice version 0.12.3 on 2020-08-12 18:22:39
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (303.53,130.09);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (303.53,130.09);
\end{scope}
\begin{scope}
\path[clip] ( 31.71, 30.69) rectangle (222.60,124.59);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.71, 30.69) rectangle (222.60,124.59);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.71, 48.66) --
	(222.60, 48.66);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.71, 78.98) --
	(222.60, 78.98);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.71,109.30) --
	(222.60,109.30);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 63.53, 30.69) --
	( 63.53,124.59);

\path[draw=drawColor,line width= 0.3pt,line join=round] (109.80, 30.69) --
	(109.80,124.59);

\path[draw=drawColor,line width= 0.3pt,line join=round] (156.08, 30.69) --
	(156.08,124.59);

\path[draw=drawColor,line width= 0.3pt,line join=round] (202.35, 30.69) --
	(202.35,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.71, 33.50) --
	(222.60, 33.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.71, 63.82) --
	(222.60, 63.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.71, 94.14) --
	(222.60, 94.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.71,124.46) --
	(222.60,124.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.39, 30.69) --
	( 40.39,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 86.66, 30.69) --
	( 86.66,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (132.94, 30.69) --
	(132.94,124.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (179.22, 30.69) --
	(179.22,124.59);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 40.39, 35.39) --
	( 42.59, 77.68) --
	( 44.78, 88.52) --
	( 46.98, 98.54) --
	( 49.18,105.04) --
	( 51.37,108.74) --
	( 53.57,110.31) --
	( 55.77,110.47) --
	( 57.96,109.90) --
	( 60.16,109.20) --
	( 62.36,108.50) --
	( 64.55,107.80) --
	( 66.75,107.10) --
	( 68.95,106.41) --
	( 71.14,105.71) --
	( 73.34,105.02) --
	( 75.53,104.33) --
	( 77.73,103.64) --
	( 79.93,102.95) --
	( 82.12,102.27) --
	( 84.32,101.58) --
	( 86.52,100.90) --
	( 88.71,100.22) --
	( 90.91, 99.54) --
	( 93.11, 98.86) --
	( 95.30, 98.18) --
	( 97.50, 97.50) --
	( 99.70, 96.83) --
	(101.89, 96.16) --
	(104.09, 95.48) --
	(106.29, 94.81) --
	(108.48, 94.14) --
	(110.68, 93.47) --
	(112.88, 92.80) --
	(115.07, 92.14) --
	(117.27, 91.47) --
	(119.47, 90.81) --
	(121.66, 90.14) --
	(123.86, 89.48) --
	(126.06, 88.82) --
	(128.25, 88.16) --
	(130.45, 87.50) --
	(132.65, 86.84) --
	(134.84, 86.18) --
	(137.04, 85.53) --
	(139.24, 84.87) --
	(141.43, 84.21) --
	(143.63, 83.56) --
	(145.83, 82.91) --
	(148.02, 82.25) --
	(150.22, 81.60) --
	(152.42, 80.95) --
	(154.61, 80.30) --
	(156.81, 79.65) --
	(159.01, 79.00) --
	(161.20, 78.35) --
	(163.40, 77.70) --
	(165.60, 77.05) --
	(167.79, 76.41) --
	(169.99, 75.76) --
	(172.19, 75.11) --
	(174.38, 74.47) --
	(176.58, 73.82) --
	(178.78, 73.18) --
	(180.97, 72.53) --
	(183.17, 71.89) --
	(185.37, 71.25) --
	(187.56, 70.60) --
	(189.76, 69.96) --
	(191.96, 69.32) --
	(194.15, 68.67) --
	(196.35, 68.03) --
	(198.55, 67.39) --
	(200.74, 66.75) --
	(202.94, 66.10) --
	(205.14, 65.46) --
	(207.33, 64.82) --
	(209.53, 64.18) --
	(211.73, 63.54) --
	(213.92, 62.90);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] ( 40.39, 34.95) --
	( 42.59, 75.01) --
	( 44.78, 84.87) --
	( 46.98, 94.16) --
	( 49.18,100.19) --
	( 51.37,103.60) --
	( 53.57,105.04) --
	( 55.77,105.17) --
	( 57.96,104.62) --
	( 60.16,103.95) --
	( 62.36,103.28) --
	( 64.55,102.61) --
	( 66.75,101.95) --
	( 68.95,101.28) --
	( 71.14,100.63) --
	( 73.34, 99.97) --
	( 75.53, 99.31) --
	( 77.73, 98.66) --
	( 79.93, 98.01) --
	( 82.12, 97.36) --
	( 84.32, 96.71) --
	( 86.52, 96.07) --
	( 88.71, 95.42) --
	( 90.91, 94.78) --
	( 93.11, 94.14) --
	( 95.30, 93.51) --
	( 97.50, 92.87) --
	( 99.70, 92.24) --
	(101.89, 91.61) --
	(104.09, 90.98) --
	(106.29, 90.35) --
	(108.48, 89.72) --
	(110.68, 89.10) --
	(112.88, 88.47) --
	(115.07, 87.85) --
	(117.27, 87.23) --
	(119.47, 86.61) --
	(121.66, 85.99) --
	(123.86, 85.38) --
	(126.06, 84.76) --
	(128.25, 84.15) --
	(130.45, 83.54) --
	(132.65, 82.93) --
	(134.84, 82.32) --
	(137.04, 81.71) --
	(139.24, 81.10) --
	(141.43, 80.50) --
	(143.63, 79.89) --
	(145.83, 79.29) --
	(148.02, 78.69) --
	(150.22, 78.08) --
	(152.42, 77.48) --
	(154.61, 76.88) --
	(156.81, 76.28) --
	(159.01, 75.69) --
	(161.20, 75.09) --
	(163.40, 74.49) --
	(165.60, 73.90) --
	(167.79, 73.30) --
	(169.99, 72.71) --
	(172.19, 72.12) --
	(174.38, 71.52) --
	(176.58, 70.93) --
	(178.78, 70.34) --
	(180.97, 69.75) --
	(183.17, 69.16) --
	(185.37, 68.57) --
	(187.56, 67.98) --
	(189.76, 67.39) --
	(191.96, 66.80) --
	(194.15, 66.22) --
	(196.35, 65.63) --
	(198.55, 65.04) --
	(200.74, 64.45) --
	(202.94, 63.86) --
	(205.14, 63.28) --
	(207.33, 62.69) --
	(209.53, 62.10) --
	(211.73, 61.52) --
	(213.92, 60.93);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] ( 40.39, 35.09) --
	( 42.59, 76.60) --
	( 44.78, 87.21) --
	( 46.98, 97.28) --
	( 49.18,103.82) --
	( 51.37,107.54) --
	( 53.57,109.13) --
	( 55.77,109.29) --
	( 57.96,108.72) --
	( 60.16,108.02) --
	( 62.36,107.32) --
	( 64.55,106.62) --
	( 66.75,105.93) --
	( 68.95,105.23) --
	( 71.14,104.54) --
	( 73.34,103.85) --
	( 75.53,103.17) --
	( 77.73,102.48) --
	( 79.93,101.80) --
	( 82.12,101.12) --
	( 84.32,100.44) --
	( 86.52, 99.76) --
	( 88.71, 99.08) --
	( 90.91, 98.41) --
	( 93.11, 97.74) --
	( 95.30, 97.07) --
	( 97.50, 96.40) --
	( 99.70, 95.73) --
	(101.89, 95.06) --
	(104.09, 94.40) --
	(106.29, 93.73) --
	(108.48, 93.07) --
	(110.68, 92.41) --
	(112.88, 91.75) --
	(115.07, 91.10) --
	(117.27, 90.44) --
	(119.47, 89.78) --
	(121.66, 89.13) --
	(123.86, 88.48) --
	(126.06, 87.83) --
	(128.25, 87.18) --
	(130.45, 86.53) --
	(132.65, 85.88) --
	(134.84, 85.23) --
	(137.04, 84.59) --
	(139.24, 83.94) --
	(141.43, 83.30) --
	(143.63, 82.66) --
	(145.83, 82.02) --
	(148.02, 81.38) --
	(150.22, 80.74) --
	(152.42, 80.10) --
	(154.61, 79.46) --
	(156.81, 78.82) --
	(159.01, 78.19) --
	(161.20, 77.55) --
	(163.40, 76.92) --
	(165.60, 76.28) --
	(167.79, 75.65) --
	(169.99, 75.02) --
	(172.19, 74.38) --
	(174.38, 73.75) --
	(176.58, 73.12) --
	(178.78, 72.49) --
	(180.97, 71.86) --
	(183.17, 71.23) --
	(185.37, 70.60) --
	(187.56, 69.97) --
	(189.76, 69.34) --
	(191.96, 68.71) --
	(194.15, 68.09) --
	(196.35, 67.46) --
	(198.55, 66.83) --
	(200.74, 66.20) --
	(202.94, 65.58) --
	(205.14, 64.95) --
	(207.33, 64.32) --
	(209.53, 63.70) --
	(211.73, 63.07) --
	(213.92, 62.44);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 40.39, 35.45) --
	( 42.59, 84.78) --
	( 44.78, 96.38) --
	( 46.98,107.26) --
	( 49.18,114.27) --
	( 51.37,118.18) --
	( 53.57,119.76) --
	( 55.77,119.78) --
	( 57.96,119.00) --
	( 60.16,118.08) --
	( 62.36,117.17) --
	( 64.55,116.27) --
	( 66.75,115.37) --
	( 68.95,114.49) --
	( 71.14,113.60) --
	( 73.34,112.73) --
	( 75.53,111.86) --
	( 77.73,111.00) --
	( 79.93,110.15) --
	( 82.12,109.30) --
	( 84.32,108.45) --
	( 86.52,107.62) --
	( 88.71,106.79) --
	( 90.91,105.97) --
	( 93.11,105.15) --
	( 95.30,104.34) --
	( 97.50,103.53) --
	( 99.70,102.73) --
	(101.89,101.93) --
	(104.09,101.14) --
	(106.29,100.36) --
	(108.48, 99.58) --
	(110.68, 98.80) --
	(112.88, 98.03) --
	(115.07, 97.27) --
	(117.27, 96.51) --
	(119.47, 95.75) --
	(121.66, 95.00) --
	(123.86, 94.26) --
	(126.06, 93.51) --
	(128.25, 92.78) --
	(130.45, 92.04) --
	(132.65, 91.31) --
	(134.84, 90.59) --
	(137.04, 89.86) --
	(139.24, 89.15) --
	(141.43, 88.43) --
	(143.63, 87.72) --
	(145.83, 87.01) --
	(148.02, 86.31) --
	(150.22, 85.61) --
	(152.42, 84.91) --
	(154.61, 84.21) --
	(156.81, 83.52) --
	(159.01, 82.83) --
	(161.20, 82.14) --
	(163.40, 81.46) --
	(165.60, 80.77) --
	(167.79, 80.09) --
	(169.99, 79.42) --
	(172.19, 78.74) --
	(174.38, 78.07) --
	(176.58, 77.39) --
	(178.78, 76.72) --
	(180.97, 76.06) --
	(183.17, 75.39) --
	(185.37, 74.72) --
	(187.56, 74.06) --
	(189.76, 73.40) --
	(191.96, 72.73) --
	(194.15, 72.07) --
	(196.35, 71.41) --
	(198.55, 70.76) --
	(200.74, 70.10) --
	(202.94, 69.44) --
	(205.14, 68.78) --
	(207.33, 68.13) --
	(209.53, 67.47) --
	(211.73, 66.82) --
	(213.92, 66.16);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] ( 40.39, 35.39) --
	( 42.59, 82.00) --
	( 44.78, 94.11) --
	( 46.98,105.31) --
	( 49.18,112.55) --
	( 51.37,116.62) --
	( 53.57,118.30) --
	( 55.77,118.38) --
	( 57.96,117.65) --
	( 60.16,116.77) --
	( 62.36,115.90) --
	( 64.55,115.03) --
	( 66.75,114.17) --
	( 68.95,113.32) --
	( 71.14,112.47) --
	( 73.34,111.63) --
	( 75.53,110.79) --
	( 77.73,109.96) --
	( 79.93,109.13) --
	( 82.12,108.31) --
	( 84.32,107.50) --
	( 86.52,106.69) --
	( 88.71,105.88) --
	( 90.91,105.08) --
	( 93.11,104.28) --
	( 95.30,103.49) --
	( 97.50,102.71) --
	( 99.70,101.92) --
	(101.89,101.15) --
	(104.09,100.37) --
	(106.29, 99.60) --
	(108.48, 98.84) --
	(110.68, 98.08) --
	(112.88, 97.32) --
	(115.07, 96.57) --
	(117.27, 95.82) --
	(119.47, 95.08) --
	(121.66, 94.34) --
	(123.86, 93.60) --
	(126.06, 92.87) --
	(128.25, 92.14) --
	(130.45, 91.41) --
	(132.65, 90.69) --
	(134.84, 89.97) --
	(137.04, 89.25) --
	(139.24, 88.54) --
	(141.43, 87.83) --
	(143.63, 87.12) --
	(145.83, 86.41) --
	(148.02, 85.71) --
	(150.22, 85.01) --
	(152.42, 84.31) --
	(154.61, 83.62) --
	(156.81, 82.92) --
	(159.01, 82.23) --
	(161.20, 81.54) --
	(163.40, 80.86) --
	(165.60, 80.17) --
	(167.79, 79.49) --
	(169.99, 78.81) --
	(172.19, 78.13) --
	(174.38, 77.45) --
	(176.58, 76.78) --
	(178.78, 76.10) --
	(180.97, 75.43) --
	(183.17, 74.76) --
	(185.37, 74.09) --
	(187.56, 73.42) --
	(189.76, 72.75) --
	(191.96, 72.09) --
	(194.15, 71.42) --
	(196.35, 70.75) --
	(198.55, 70.09) --
	(200.74, 69.42) --
	(202.94, 68.76) --
	(205.14, 68.10) --
	(207.33, 67.44) --
	(209.53, 66.77) --
	(211.73, 66.11) --
	(213.92, 65.45);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt on 4pt off 3pt ,line join=round] ( 40.39, 35.44) --
	( 42.59, 84.85) --
	( 44.78, 96.76) --
	( 46.98,107.71) --
	( 49.18,114.77) --
	( 51.37,118.71) --
	( 53.57,120.30) --
	( 55.77,120.32) --
	( 57.96,119.54) --
	( 60.16,118.61) --
	( 62.36,117.70) --
	( 64.55,116.79) --
	( 66.75,115.89) --
	( 68.95,115.00) --
	( 71.14,114.12) --
	( 73.34,113.24) --
	( 75.53,112.37) --
	( 77.73,111.50) --
	( 79.93,110.64) --
	( 82.12,109.79) --
	( 84.32,108.95) --
	( 86.52,108.11) --
	( 88.71,107.28) --
	( 90.91,106.45) --
	( 93.11,105.63) --
	( 95.30,104.82) --
	( 97.50,104.01) --
	( 99.70,103.21) --
	(101.89,102.41) --
	(104.09,101.62) --
	(106.29,100.83) --
	(108.48,100.05) --
	(110.68, 99.27) --
	(112.88, 98.50) --
	(115.07, 97.74) --
	(117.27, 96.97) --
	(119.47, 96.22) --
	(121.66, 95.47) --
	(123.86, 94.72) --
	(126.06, 93.97) --
	(128.25, 93.24) --
	(130.45, 92.50) --
	(132.65, 91.77) --
	(134.84, 91.04) --
	(137.04, 90.32) --
	(139.24, 89.60) --
	(141.43, 88.89) --
	(143.63, 88.18) --
	(145.83, 87.47) --
	(148.02, 86.76) --
	(150.22, 86.06) --
	(152.42, 85.36) --
	(154.61, 84.67) --
	(156.81, 83.97) --
	(159.01, 83.28) --
	(161.20, 82.60) --
	(163.40, 81.91) --
	(165.60, 81.23) --
	(167.79, 80.55) --
	(169.99, 79.87) --
	(172.19, 79.20) --
	(174.38, 78.52) --
	(176.58, 77.85) --
	(178.78, 77.18) --
	(180.97, 76.51) --
	(183.17, 75.85) --
	(185.37, 75.18) --
	(187.56, 74.52) --
	(189.76, 73.86) --
	(191.96, 73.20) --
	(194.15, 72.54) --
	(196.35, 71.88) --
	(198.55, 71.22) --
	(200.74, 70.56) --
	(202.94, 69.91) --
	(205.14, 69.25) --
	(207.33, 68.59) --
	(209.53, 67.94) --
	(211.73, 67.28) --
	(213.92, 66.63);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 31.71, 30.69) rectangle (222.60,124.59);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 26.76, 30.47) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 26.76, 60.79) {5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 26.76, 91.11) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 26.76,121.43) {15};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96, 33.50) --
	( 31.71, 33.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96, 63.82) --
	( 31.71, 63.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96, 94.14) --
	( 31.71, 94.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,124.46) --
	( 31.71,124.46);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.39, 27.94) --
	( 40.39, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 86.66, 27.94) --
	( 86.66, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (132.94, 27.94) --
	(132.94, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (179.22, 27.94) --
	(179.22, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 40.39, 19.68) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 86.66, 19.68) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (132.94, 19.68) {0.4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (179.22, 19.68) {0.6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (127.16,  7.64) {Proportion of listed facts};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08, 77.64) {Inference time (s)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (233.60, 21.17) rectangle (298.03,134.11);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (239.10,119.96) {Algorithm};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (239.10, 98.94) rectangle (253.55,113.39);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 1.1pt,line join=round] (240.54,106.16) -- (252.11,106.16);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (239.10, 84.48) rectangle (253.55, 98.94);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] (240.54, 91.71) -- (252.11, 91.71);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (239.10, 70.03) rectangle (253.55, 84.48);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] (240.54, 77.26) -- (252.11, 77.26);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (239.10, 55.57) rectangle (253.55, 70.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] (240.54, 62.80) -- (252.11, 62.80);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (239.10, 41.12) rectangle (253.55, 55.57);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] (240.54, 48.35) -- (252.11, 48.35);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (239.10, 26.67) rectangle (253.55, 41.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt on 4pt off 3pt ,line join=round] (240.54, 33.89) -- (252.11, 33.89);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (259.05,103.13) {BDD};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (259.05, 88.68) {d-DNNF};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (259.05, 74.23) {K-Best};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (259.05, 59.77) {NNF};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (259.05, 45.32) {SDD};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (259.05, 30.86) {SDDX};
\end{scope}
\end{tikzpicture}

\end{document}